---
title: "Counts Analysis"
author: "Rachael Murray"
date: "2024-12-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(stringr)
theme_set(theme_cowplot(font_size = 7) + theme(strip.background = element_blank()))
```
# Summary

This file analyses the counts per gene across samples.

* reads in the sample sheet `input/experiment/CryptoRNAseqFLC1SampleKey.csv`
* reads in raw count data for all samples from file `results/counts/counts.txt`, output by `featureCounts` running in the nextflow pipeline, which has gene (Geneid) as rows and counts as columns
* normalizes to counts per million for each sample using function `calculate_cpm`
* makes a "tidy" or long-format dataset with counts and CPM spread out by Geneid and Sampleid and sample information
* uses tidy data to plot expression against time for individual genes

This analysis has largely been superseded by rlog (regularized logarithm) normalisation.


### Define functions

Define some functions that will be useful for the analysis

```{r count_functions}
calculate_cpm <- function(x, round_digits = NA) {
    cpm <- x / sum(x, na.rm=TRUE) * 1e6
    if( !is.na(round_digits) ) {
        cpm <- round(cpm, digits = round_digits)
    }
    cpm
}
```


## Load counts and sample sheet

```{r load_counts}
sample_sheet <- 
  here::here("input", "experiment","CryptoRNAseqFLC1SampleKey.csv") %>%
  read_csv() %>%
  select(!Strain) %>%
  mutate(FLC1 = factor(FLC1, 
                       levels = c("WT", "delta")),
         Condition = paste(Media, "_", Temperature, sep = ""),
         BioRep = as.character(BioRep)
  ) %>% 
  mutate(Condition = factor(Condition, 
                            levels = c("YPD_30", "YPD_37", "YPD_CFW_37", "YPD_EGTA_37", "YPD_CFW_EGTA_37"),
                            labels = c("30°C", "37°C", "37°C\n + CFW", "37°C\n + EGTA", "37°C\n + CFW\n + EGTA"))
  )


counts_bygene_wide <- 
  here::here("results","counts", "counts.txt") %>%
  read_tsv(comment = "#")
counts_bygene_wide
```


## Make tidy (long-format) counts for plotting

```{r make_tidy_counts, dependson = "load_counts"}
counts_bygene_long <- counts_bygene_wide %>%
    select(-Chr,-Start,-End,-Strand,-Length) %>%
    pivot_longer(ends_with(".bam"), names_to = "SampleID", values_to = "Count") %>%
    mutate(SampleID = as.numeric(word(SampleID, 2, sep = "_"))) %>%
    group_by(SampleID) %>%
    mutate(TPM = calculate_cpm(Count, round_digits = 2)) %>%
    ungroup() %>%
    left_join(sample_sheet, by = "SampleID") %>%
    select(SampleID, SampleName, FLC1, Condition, Media, Temperature, BioRep, Geneid, Count, TPM) %>%
    arrange(Geneid, FLC1, Condition, Media, Temperature, BioRep)

counts_bygene_long

```
### Write out to file `counts_quantseqfwd_CryptoRNAseqFLC1_tidy.txt`

```{r write_tidy_counts, dependson = "make_tidy_counts"}
file_counts_long <- here::here("results",
                               "counts_quantseqfwd_CryptoRNAseqFLC1_tidy.txt"
                               )

write_lines(c("# counts_quantseqfwd_CryptoRNAseqFLC1_tidy.txt",
              "# Tidy/long-format counts from QuantSeq FLC1∆ RNAseq in Cryptococcus neoformans",
              paste("# ", date()),
              "# Columns:",
              "#   SampleName: encodes Strain, Tepmerature, Media, Replicate",
              "#   FLC1: genotype for FLC1, either WT (present) or delta (deletion/disruption)",
              "#   Media: growth media  YPD, YPD_CFW (50uM CFW), YPD_EGTA (10mM EGTA), YPD_CFW_EGTA (50uM CFW + 10mM EGTA)",
              "#   Temperature: 30°C or 37°C",
              "#   BioRep: biological replicate (1, 2 or 3)",
              "#   Geneid: Geneid with systematic Cryptococcus H99 gene name",
              "#   Count: QuantSeq 3'-end counts assigned to gene",
              "#   TPM: Transcripts per million (normalized Count for each sample)",
              "# "),
            file = file_counts_long)
#write_tsv(counts_bygene_long %>% select(-Condition), file = file_counts_long, col_names = TRUE, append = TRUE)
```

## Plot some individual genes

```{r plot_my_gene, dependson = "make_tidy_counts"}
plot_my_genes_TPM <- function(mygenes = "CNAG_06125",
                              mydata = counts_bygene_long,
                              mymedia = c("YPD", "YPD_EGTA", "YPD_CFW"),
                              mytemperature = c(30, 37)){
    ggplot(data = filter(mydata, 
                         Geneid %in% mygenes,
                         Media %in% mymedia,
                         Temperature %in% mytemperature),
                         aes(x = Condition, y = TPM, colour = FLC1)) +
        geom_point(aes(shape = BioRep), size = 1, alpha = 0.7) +
        stat_summary(aes(group=FLC1), fun.min="mean",fun.max="mean",
                     geom="errorbar",linewidth=1,width=0.4) +  
        guides(shape = guide_legend(override.aes = list(size = 1, alpha = 1)),
               color = guide_legend(override.aes = list(linetype = 1, shape = NA))) +
        #scale_y_log10() +
        theme(axis.title.x=element_blank()) + 
        facet_grid(.~Geneid) 
}


cowplot::plot_grid(
plot_my_genes_TPM("CNAG_04283") + labs(title = "FLC1"),
plot_my_genes_TPM("CNAG_01208") + labs(title = "CCH1"),
plot_my_genes_TPM("CNAG_00036") + labs(title = "SEC14"),
plot_my_genes_TPM("CNAG_01230") + labs(title = "chitin deacetylase 2"),
plot_my_genes_TPM("CNAG_01562") + labs(title = "pr4/barwin domain protein"),
plot_my_genes_TPM("CNAG_06501") + labs(title = "GAS1"),
plot_my_genes_TPM("CNAG_05818") + labs(title = "CHS5"),
plot_my_genes_TPM("CNAG_01232") + labs(title = "PMC1"),
plot_my_genes_TPM("CNAG_02714") + labs(title = "EFB1"),
rows = 3)


cowplot::plot_grid(
  plot_my_genes_TPM("CNAG_00327") + labs(title = ""),
  plot_my_genes_TPM("CNAG_01656") + labs(title = "")
)

plot_my_genes_TPM("CNAG_01653") + labs(title = "CIG1")
plot_my_genes_TPM("CNAG_00522") + labs(title = "TRICALBIN")
plot_my_genes_TPM("CNAG_02045") + labs(title = "")
plot_my_genes_TPM("CNAG_04105") + labs(title = "")
plot_my_genes_TPM("CNAG_05785") + labs(title = "")
plot_my_genes_TPM("CNAG_06241") + labs(title = "")


plot_my_genes_TPM("CNAG_01824") + labs(title = "TRICALBIN")
plot_my_genes_TPM("CNAG_00522") + labs(title = "TRICALBIN")
plot_my_genes_TPM("CNAG_04200") + labs(title = "TRICALBIN")





plot_grid(
  plot_my_genes_TPM("CNAG_06060") + labs(title = ""),
  plot_my_genes_TPM("CNAG_03597") + labs(title = ""),
  plot_my_genes_TPM("CNAG_00923") + labs(title = ""),
  plot_my_genes_TPM("CNAG_06381") + labs(title = ""),
  plot_my_genes_TPM("CNAG_03595") + labs(title = "")
)


plot_grid(
  plot_my_genes_TPM("CNAG_01446") + labs(title = ""),
  plot_my_genes_TPM("CNAG_05666") + labs(title = ""),
  plot_my_genes_TPM("CNAG_04553") + labs(title = ""),
  plot_my_genes_TPM("CNAG_00981") + labs(title = ""),
  plot_my_genes_TPM("CNAG_02877") + labs(title = "")
)



plot_grid(
  plot_my_genes_TPM("CNAG_01954") + labs(title = ""),
  plot_my_genes_TPM("CNAG_02526") + labs(title = ""),
  plot_my_genes_TPM("CNAG_00587") + labs(title = ""),
  plot_my_genes_TPM("CNAG_04142") + labs(title = ""),
  plot_my_genes_TPM("CNAG_01223") + labs(title = "")
)

plot_grid(
  plot_my_genes_TPM("CNAG_00162") + labs(title = ""),
  plot_my_genes_TPM("CNAG_06297") + labs(title = ""),
  plot_my_genes_TPM("CNAG_06102") + labs(title = ""),
  plot_my_genes_TPM("CNAG_00443") + labs(title = ""),
  plot_my_genes_TPM("CNAG_06438") + labs(title = ""),
  plot_my_genes_TPM("CNAG_01539") + labs(title = ""),
  plot_my_genes_TPM("CNAG_00828") + labs(title = ""),
  plot_my_genes_TPM("CNAG_06578") + labs(title = ""),
  plot_my_genes_TPM("CNAG_01621") + labs(title = "")
)



CNAG_03597

CNAG_02892

CNAG_05785

CNAG_02045
CNAG_04837
CNAG_00522
CNAG_01653


plot_grid(
  plot_my_genes_TPM("CNAG_01446") + labs(title = ""),
  plot_my_genes_TPM("CNAG_05256") + labs(title = ""),
  plot_my_genes_TPM("CNAG_06291") + labs(title = ""),
  plot_my_genes_TPM("CNAG_06698") + labs(title = ""),
  plot_my_genes_TPM("CNAG_06438") + labs(title = ""),
  plot_my_genes_TPM("CNAG_05258") + labs(title = ""),
  plot_my_genes_TPM("CNAG_06576") + labs(title = ""),
  plot_my_genes_TPM("CNAG_02877") + labs(title = ""),
  plot_my_genes_TPM("CNAG_03465") + labs(title = "")
)



plot_grid(
plot_my_genes_TPM("CNAG_00036") + labs(title = ""),		
plot_my_genes_TPM("CNAG_01585") + labs(title = ""),		
plot_my_genes_TPM("CNAG_02718") + labs(title = ""),		
plot_my_genes_TPM("CNAG_03495") + labs(title = ""),		
plot_my_genes_TPM("CNAG_03572") + labs(title = ""),		
plot_my_genes_TPM("CNAG_03667") + labs(title = ""),		
plot_my_genes_TPM("CNAG_03857") + labs(title = ""),		
plot_my_genes_TPM("CNAG_04142") + labs(title = ""),		
plot_my_genes_TPM("CNAG_04157") + labs(title = ""),		
plot_my_genes_TPM("CNAG_04183") + labs(title = ""),
plot_my_genes_TPM("CNAG_06297") + labs(title = "")
)


plot_my_genes_TPM("CNAG_02008") + labs(title = "OVA1")
plot_my_genes_TPM("CNAG_00036") + labs(title = "SEC14")
plot_my_genes_TPM("CNAG_01653") + labs(title = "CIG1")






# Zinc transporter
plot_my_genes_TPM("CNAG_00895") + labs(title = "ZIP1")
plot_my_genes_TPM("CNAG_03398") + labs(title = "ZIP2")

# Iron uptake 


plot_grid(
plot_my_genes_TPM("CNAG_00979") + labs(title = "CTR4"),
plot_my_genes_TPM("CNAG_07701") + labs(title = "CTR1"),
plot_my_genes_TPM("CNAG_02775") + labs(title = "BIM1")
)

plot_my_genes_TPM("CNAG_03315") + labs(title = "")
CNAG_00110
CNAG_00979
CNAG_02864


plot_grid(
plot_my_genes_TPM("CNAG_05135") + labs(title = "PMR1-like"),
plot_my_genes_TPM("CNAG_05592") + labs(title = "PMR1-like")
)


plot_my_genes_TPM("CNAG_00066") + labs(title = "FLC type II")
plot_my_genes_TPM("CNAG_03751") + labs(title = "MID1")

```

# Glucan metabolic genes upregulated by EGTA in the delta mutant

```{r}
cowplot::plot_grid(
  plot_my_genes_TPM("CNAG_06931", mymedia = c("YPD","YPD_EGTA"), mytemperature = 37) + labs(title = "Beta-glucosidase"),
  plot_my_genes_TPM("CNAG_00393", mymedia = c("YPD","YPD_EGTA"), mytemperature = 37) + labs(title = "1,4-alpha-glucan-branching enzyme"),
  plot_my_genes_TPM("CNAG_00897", mymedia = c("YPD","YPD_EGTA"), mytemperature = 37) + labs(title = "putative glucosidase"),
  plot_my_genes_TPM("CNAG_04879", mymedia = c("YPD","YPD_EGTA"), mytemperature = 37) + labs(title = "Glycogen debranching enzyme")
)
```

# Cell wall gene down in WT vs FLC1∆ at 37°C 

```{r}

cowplot::plot_grid(
  plot_my_genes_TPM("CNAG_06508", mymedia = c("YPD","YPD_EGTA"), mytemperature = c(37, 30)) + labs(title = "FKS1"),
  plot_my_genes_TPM("CNAG_06835", mymedia = c("YPD","YPD_EGTA"), mytemperature = c(37, 30)) + labs(title = "KRE61"),
  plot_my_genes_TPM("CNAG_06031", mymedia = c("YPD","YPD_EGTA"), mytemperature = c(37, 30)) + labs(title = "KRE63"),
  plot_my_genes_TPM("CNAG_05818", mymedia = c("YPD","YPD_EGTA"), mytemperature = c(37, 30)) + labs(title = "CHS5"),
  plot_my_genes_TPM("CNAG_02217", mymedia = c("YPD","YPD_EGTA"), mytemperature = c(37, 30)) + labs(title = "CHS7")
)
```










# Hypothetic proteins upregulated by EGTA in the delta mutant

```{r}
cowplot::plot_grid(
  plot_my_genes_TPM("CNAG_04877", mymedia = c("YPD","YPD_EGTA"), mytemperature = 37) + labs(title = ""),
  plot_my_genes_TPM("CNAG_00091", mymedia = c("YPD","YPD_EGTA"), mytemperature = 37) + labs(title = ""),
  plot_my_genes_TPM("CNAG_07912", mymedia = c("YPD","YPD_EGTA"), mytemperature = 37) + labs(title = ""),
  plot_my_genes_TPM("CNAG_06109", mymedia = c("YPD","YPD_EGTA"), mytemperature = 37) + labs(title = ""),
  plot_my_genes_TPM("CNAG_00984", mymedia = c("YPD","YPD_EGTA"), mytemperature = 37) + labs(title = "glucose and ribitol dehydrogenase"))

```


```{r}
cowplot::plot_grid(
  plot_my_genes_TPM("CNAG_00979", mymedia = c("YPD", "YPD_EGTA"), mytemperature = 37) + labs(title = "CTR4"),
  plot_my_genes_TPM("CNAG_07701", mymedia = c("YPD", "YPD_EGTA"), mytemperature = 37) + labs(title = "CTR1"),
  plot_my_genes_TPM("CNAG_02775", mymedia = c("YPD", "YPD_EGTA"), mytemperature = 37) + labs(title = "BIM1"),
  plot_my_genes_TPM("CNAG_02217", mymedia = c("YPD", "YPD_EGTA"), mytemperature = 37) + labs(title = "CHS7"),
  plot_my_genes_TPM("CNAG_06487", mymedia = c("YPD", "YPD_EGTA"), mytemperature = 37) + labs(title = "CHS6"),
  plot_my_genes_TPM("CNAG_00876", mymedia = c("YPD", "YPD_EGTA"), mytemperature = 37) + labs(title = "FRE7"),
  plot_my_genes_TPM("CNAG_07334", mymedia = c("YPD", "YPD_EGTA"), mytemperature = 37) + labs(title = "FRE4"),
  plot_my_genes_TPM("CNAG_03498", mymedia = c("YPD", "YPD_EGTA"), mytemperature = 37) + labs(title = "FRE201"),
  plot_my_genes_TPM("CNAG_01653", mymedia = c("YPD", "YPD_EGTA"), mytemperature = 37) + labs(title = "CIG2"),
  plot_my_genes_TPM("CNAG_03398", mymedia = c("YPD", "YPD_EGTA"), mytemperature = 37) + labs(title = "ZIP2"),
  plot_my_genes_TPM("CNAG_06576", mymedia = c("YPD", "YPD_EGTA"), mytemperature = 37) + labs(title = "CAR1"),
  plot_my_genes_TPM("CNAG_06291", mymedia = c("YPD", "YPD_EGTA"), mytemperature = 37) + labs(title = "FPD1"),
  plot_my_genes_TPM("CNAG_01341", mymedia = c("YPD", "YPD_EGTA"), mytemperature = 37) + labs(title = "ISP3"),
  plot_my_genes_TPM("CNAG_00091", mymedia = c("YPD", "YPD_EGTA"), mytemperature = 37) + labs(title = "CNAG_00091")
  )
```

```{r}
plot_my_genes_TPM("CNAG_03385")
plot_my_genes_TPM("CNAG_03385")
plot_my_genes_TPM("CNAG_01446")
plot_my_genes_TPM("CNAG_05387")

```

```{r}
TFs <- read_csv(here("TFlist.csv")) %>% 
  mutate(gene = paste("CNAG_", gene, sep = "")) %>% 
  mutate(Geneid = gene)

TFs %>% 
  left_join(counts_bygene_long)
```
# Differential expression of transciption factors in WT vs FLC1 at 30°C and/or 37°C

```{r}
cowplot::plot_grid(
  plot_my_genes_TPM("CNAG_04837", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "MLN1"),
  plot_my_genes_TPM("CNAG_04457", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "FZC30"),
  plot_my_genes_TPM("CNAG_00559", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "BZP3"),
  plot_my_genes_TPM("CNAG_01841", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "GLN3"),
  plot_my_genes_TPM("CNAG_05431", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "RIM101"),
  plot_my_genes_TPM("CNAG_02877", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "FZC51"),
  plot_my_genes_TPM("CNAG_05861", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "FKH101"),
  plot_my_genes_TPM("CNAG_04804", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "SRE1"),
  plot_my_genes_TPM("CNAG_06276", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "CEP3"),
  plot_my_genes_TPM("CNAG_02516", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "HLH5"),
  plot_my_genes_TPM("CNAG_00239", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "YAP1"),
  plot_my_genes_TPM("CNAG_00871", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "CLR3"),
  plot_my_genes_TPM("CNAG_03998", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "RLM1"),
  plot_my_genes_TPM("CNAG_00828", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "SIP401"),
  plot_my_genes_TPM("CNAG_03346", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "BZP4"),
  plot_my_genes_TPM("CNAG_00791", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "HLH1"),
  plot_my_genes_TPM("CNAG_02476", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "YRM101"),
  plot_my_genes_TPM("CNAG_02774", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "MAL13"),
  plot_my_genes_TPM("CNAG_01551", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "GAT201"),
  plot_my_genes_TPM("CNAG_00156", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37)) + labs(title = "CRZ1")) 


ggsave(here("TPM_plots_TFs_30C_37C_YPD.png"), width = 15, height = 10, bg = "white")


```



```{r}
plot_my_genes_TPM("CNAG_00091", mymedia = c("YPD", "YPD_EGTA"), mytemperature = c(30,37))
plot_my_genes_TPM("CNAG_04891", mymedia = c("YPD", "YPD_EGTA", "YPD_CFW"), mytemperature = c(30,37))

cowplot::plot_grid(
plot_my_genes_TPM("CNAG_01954", mymedia = c("YPD"), mytemperature = c(30,37)),
plot_my_genes_TPM("CNAG_02526", mymedia = c("YPD"), mytemperature = c(30,37)),
plot_my_genes_TPM("CNAG_00587", mymedia = c("YPD"), mytemperature = c(30,37)),
plot_my_genes_TPM("CNAG_04142", mymedia = c("YPD"), mytemperature = c(30,37)),
plot_my_genes_TPM("CNAG_01223", mymedia = c("YPD"), mytemperature = c(30,37)),
plot_my_genes_TPM("CNAG_00162", mymedia = c("YPD"), mytemperature = c(30,37)),
plot_my_genes_TPM("CNAG_05667", mymedia = c("YPD"), mytemperature = c(30,37)),
plot_my_genes_TPM("CNAG_02526", mymedia = c("YPD"), mytemperature = c(30,37))
)

CNAG_02526
CNAG_01954
CNAG_01653
```



