---
title: "Differential Gene Expression of CFW vs YPD in FLC1∆ at 37°C"
author: "Rachael Murray"
date: "2024-12-19"
output: html_document
---

# Summary

Differential Gene Expression +CFW vs -CFW in FLC1∆ at 37°C samples, using DESeq2

This analysis produces files:

- `deseq_df_CFWvsYPD_delta_37C.txt`
- `deglist_CFWvsYPD_delta_37C_up2x_FDR5.txt`
- `deglist_CFWvsYPD_delta_37C_down2x_FDR5.txt`



## Load count data from shared script load_count_data.Rmd

```{r call_load_count_data, child = "load_count_data.Rmd"}
```


# DESeq on a subset of data, only WT in CFW and YPD at 37C

```{r subset_CFWvsYPD_delta_37C, fig.height=8, fig.width=8}
sample_sheet_CFWvsYPD_delta_37C <-
  dplyr::filter(sample_sheet,
                FLC1 == "delta", 
                Media %in% c("YPD", "YPD_CFW"), 
                Temperature == "37"
  )

dds_CFWvsYPD_delta_37C <- 
  DESeqDataSetFromMatrix(
    countData = counts_all %>%
      dplyr::select(sample_sheet_CFWvsYPD_delta_37C$SampleName) %>%
      magrittr::set_rownames(counts_all$Gene),
    colData = sample_sheet_CFWvsYPD_delta_37C,
    design = ~ Media) %>%
  DESeq()

dds_CFWvsYPD_delta_37C
```


```{r results_CFWvsYPD_delta_37C}
results(dds_CFWvsYPD_delta_37C)

resultsNames(dds_CFWvsYPD_delta_37C)
```

```{r deseq_df_CFWvsYPD_delta_37C}
deseq_df_CFWvsYPD_delta_37C <- 
  biobroom::tidy.DESeqDataSet(dds_CFWvsYPD_delta_37C) %>%
  dplyr::filter(term == "Media_YPD_CFW_vs_YPD") %>%
  dplyr::left_join(gene_info_df,
                   by = "gene") %>% 
  dplyr::select(gene, 
                geneName, 
                baseMean, 
                log2FC = estimate, 
                stderror,
                padj = p.adjusted)

```



## List DEGs most UP in CFW vs YPD in delta at 37°C

We cut off at 2x changes with 5% FDR here.

```{r DEGs_CFW_up, fig.width = 6, fig.height = 3}
# Make data frames of DEGs up and down, arranged by log2FC
DEGdf_CFW_up2x_FDR5 <- 
  dplyr::filter(deseq_df_CFWvsYPD_delta_37C,
                log2FC > 1, padj < 0.05) %>%
  dplyr::arrange(desc(log2FC))
print(DEGdf_CFW_up2x_FDR5, n = 20)
```

Print the top 20 *named* DEGs

```{r DEGs_CFW_up_named}
DEGdf_CFW_up2x_FDR5 %>%
  dplyr::filter(!is.na(geneName)) %>%
  print(n = 20)
```



## List DEGs most DOWN in CFW vs YPD in WT at 37°C


```{r DEGs_CFW_down, fig.width = 6, fig.height = 3}
# Make data frames of DEGs up and down, arranged by log2FC
DEGdf_CFW_down2x_FDR5 <- 
  dplyr::filter(deseq_df_CFWvsYPD_delta_37C,
                log2FC < -1, padj < 0.05) %>%
  dplyr::arrange(log2FC)
print(DEGdf_CFW_down2x_FDR5, n = 20)
```

Print the top 20 *named* DEGs

```{r DEGs_30C_down_named}
DEGdf_CFW_down2x_FDR5 %>%
  dplyr::filter(!is.na(geneName)) %>%
  print(n = 20)
```


## Volcano plot of DEGs

```{r volcano_CFWvsYPD_delta_37C, fig.width = 6, fig.height = 3}
volcano_CFWvsYPD_delta_37C <- 
  ggplot(data = deseq_df_CFWvsYPD_delta_37C,
         aes(x = log2FC, y = -log10(padj)) ) +
  geom_point(size = 0.5, colour = "grey50") +
  geom_point(data = DEGdf_CFW_up2x_FDR5,
             size = 1, colour = "darkblue") +
  geom_point(data = DEGdf_CFW_down2x_FDR5,
             size = 1, colour = "darkred") +
  scale_y_continuous("-log10(p)",
                     limits = c(0,20), expand = c(0,0),
                     oob=scales::squish) +
  scale_x_continuous("log2 fold-change, YPD + CFW vs YPD, FLC1∆, 37°C",
                     breaks = -5:5,
                     limits = c(-4,4), expand = c(0.01,0.01),
                     oob = scales::squish) +
  coord_cartesian(clip = "off") 
volcano_CFWvsYPD_delta_37C
```



## Write outputs to files

```{r write_DEGs}
deseq_df_CFWvsYPD_delta_37C %>%
  dplyr::mutate(
    across(c(baseMean, log2FC, stderror), round, digits = 3)
  ) %>%
  dplyr::mutate(padj = signif(padj, digits = 4)) %>%
  write_tsv(file = here::here("results", "DEGs",
                              "deseq_df_CFWvsYPD_delta_37C.txt"))

DEGdf_CFW_up2x_FDR5 %>%
  dplyr::pull(gene) %>%
  write_lines(file = here::here("results", "DEGs",
                            "deglist_CFWvsYPD_delta_37C_up2x_FDR5.txt"))

DEGdf_CFW_down2x_FDR5 %>%
  dplyr::pull(gene) %>%
  write_lines(file = here::here("results", "DEGs",
                            "deglist_CFWvsYPD_delta_37C_down2x_FDR5.txt"))
```


