---
title: "DGE_EGTAvsYPD_37C_WT_delta_thesis"
author: "Rachael Murray"
date: "2025-03-04"
output: html_document
---

# Summary

Differential Gene Expression +EGTA vs -EGTA in WT and +EGTA vs -EGTA in flc1∆ at 37°C samples, using DESeq2

This analysis produces files:

- `deseq_df_EGTAvsYPD_WT_37C.txt`
- `deglist_EGTAvsYPD_WT_37C_up2x_FDR5.txt`
- `deglist_EGTAvsYPD_WT_37C_down2x_FDR5.txt`
- `deseq_df_EGTAvsYPD_delta_37C.txt`
- `deglist_EGTAvsYPD_WT_delta_up2x_FDR5.txt`
- `deglist_EGTAvsYPD_WT_delta_down2x_FDR5.txt`

## EGTA vs YPD in WT at 37°C

### Load count data from shared script load_count_data.Rmd

```{r call_load_count_data, child = "load_count_data.Rmd"}
```

### DESeq on a subset of data, only WT in YPD

```{r subset_EGTAvsYPD_WT_37C, fig.height=8, fig.width=8}
sample_sheet_EGTAvsYPD_WT_37C <-
  dplyr::filter(sample_sheet,
                FLC1 == "WT", 
                Media %in% c("YPD", "YPD_EGTA"), 
                Temperature == "37"
  )

dds_EGTAvsYPD_WT_37C <- 
  DESeqDataSetFromMatrix(
    countData = counts_all %>%
      dplyr::select(sample_sheet_EGTAvsYPD_WT_37C$SampleName) %>%
      magrittr::set_rownames(counts_all$Gene),
    colData = sample_sheet_EGTAvsYPD_WT_37C,
    design = ~ Media) %>%
  DESeq()

dds_EGTAvsYPD_WT_37C
```


```{r results_EGTAvsYPD_WT_37C}
results(dds_EGTAvsYPD_WT_37C)

resultsNames(dds_EGTAvsYPD_WT_37C)
```

```{r deseq_df_EGTAvsYPD_WT_37C}
deseq_df_EGTAvsYPD_WT_37C <- 
  biobroom::tidy.DESeqDataSet(dds_EGTAvsYPD_WT_37C) %>%
  dplyr::filter(term == "Media_YPD_EGTA_vs_YPD") %>%
  dplyr::left_join(gene_info_df,
                   by = "gene") %>% 
  dplyr::select(gene, 
                geneName, 
                baseMean, 
                log2FC = estimate, 
                stderror,
                padj = p.adjusted)

```



### List DEGs most UP in EGTA vs YPD in WT at 37°C

We cut off at 2x changes with 5% FDR here.

```{r DEGs_EGTA_up_WT, fig.width = 6, fig.height = 3}
# Make data frames of DEGs up and down, arranged by log2FC
DEGdf_EGTA_WT_up2x_FDR5 <- 
  dplyr::filter(deseq_df_EGTAvsYPD_WT_37C,
                log2FC > 1, padj < 0.05) %>%
  dplyr::arrange(desc(log2FC))
print(DEGdf_EGTA_WT_up2x_FDR5, n = 20)
```

Print the top 20 *named* DEGs

```{r DEGs_EGTA_up_named_WT}
DEGdf_EGTA_WT_up2x_FDR5 %>%
  dplyr::filter(!is.na(geneName)) %>%
  print(n = 20)
```


### List DEGs most DOWN in EGTA vs YPD in WT at 37°C

```{r DEGs_EGTA_down_WT, fig.width = 6, fig.height = 3}
# Make data frames of DEGs up and down, arranged by log2FC
DEGdf_EGTA_WT_down2x_FDR5 <- 
  dplyr::filter(deseq_df_EGTAvsYPD_WT_37C,
                log2FC < -1, padj < 0.05) %>%
  dplyr::arrange(log2FC)
print(DEGdf_EGTA_WT_down2x_FDR5, n = 20)
```

Print the top 20 *named* DEGs

```{r DEGs_30C_down_named_WT}
DEGdf_EGTA_WT_down2x_FDR5 %>%
  dplyr::filter(!is.na(geneName)) %>%
  print(n = 20)
```

### Volcano plot of DEGs

```{r volcano_EGTAvsYPD_WT_37C, fig.width = 6, fig.height = 3}
volcano_EGTAvsYPD_WT_37C <- 
  ggplot(data = deseq_df_EGTAvsYPD_WT_37C,
         aes(x = log2FC, y = -log10(padj)) ) +
  geom_point(size = 0.3, colour = "grey70") +
  geom_point(data = DEGdf_EGTA_WT_up2x_FDR5,
             size = 0.3, colour = "darkblue") +
  geom_point(data = DEGdf_EGTA_WT_down2x_FDR5,
             size = 0.3, colour = "darkred") +
  scale_y_continuous("-log10(p)",
                     limits = c(0,50), expand = c(0,0),
                     oob=scales::squish) +
  scale_x_continuous("log2 fold-change\n← up in YPD       up in YPD + EGTA →",
                     breaks = -4:4,
                     limits = c(-4,4), expand = c(0.01,0.01),
                     oob = scales::squish) +
  coord_cartesian(clip = "off") 
volcano_EGTAvsYPD_WT_37C

ggsave(here("EGTAvsYPD_WT_37C_volcano.png"), volcano_EGTAvsYPD_WT_37C, width = 4, height = 3)
```


### Write outputs to files

```{r write_DEGs_WT}
deseq_df_EGTAvsYPD_WT_37C %>%
  dplyr::mutate(
    across(c(baseMean, log2FC, stderror), round, digits = 3)
  ) %>%
  dplyr::mutate(padj = signif(padj, digits = 4)) %>%
  write_tsv(file = here::here("results", "DEGs",
                              "deseq_df_EGTAvsYPD_WT_37C.txt"))

DEGdf_EGTA_WT_up2x_FDR5 %>%
  dplyr::pull(gene) %>%
  write_lines(file = here::here("results", "DEGs",
                            "deglist_EGTAvsYPD_WT_37C_up2x_FDR5.txt"))

DEGdf_EGTA_WT_down2x_FDR5 %>%
  dplyr::pull(gene) %>%
  write_lines(file = here::here("results", "DEGs",
                            "deglist_EGTAvsYPD_WT_37C_down2x_FDR5.txt"))
```


## EGTA vs YPD in flc1∆ at 37°C

### DESeq on a subset of data, only delta in YPD

```{r subset_EGTAvsYPD_delta_37C, fig.height=8, fig.width=8}
sample_sheet_EGTAvsYPD_delta_37C <-
  dplyr::filter(sample_sheet,
                FLC1 == "delta", 
                Media %in% c("YPD", "YPD_EGTA"), 
                Temperature == "37"
  )

dds_EGTAvsYPD_delta_37C <- 
  DESeqDataSetFromMatrix(
    countData = counts_all %>%
      dplyr::select(sample_sheet_EGTAvsYPD_delta_37C$SampleName) %>%
      magrittr::set_rownames(counts_all$Gene),
    colData = sample_sheet_EGTAvsYPD_delta_37C,
    design = ~ Media) %>%
  DESeq()

dds_EGTAvsYPD_delta_37C
```



```{r results_EGTAvsYPD_delta_37C}
results(dds_EGTAvsYPD_delta_37C)

resultsNames(dds_EGTAvsYPD_delta_37C)
```

```{r deseq_df_EGTAvsYPD_delta_37C}
deseq_df_EGTAvsYPD_delta_37C <- 
  biobroom::tidy.DESeqDataSet(dds_EGTAvsYPD_delta_37C) %>%
  dplyr::filter(term == "Media_YPD_EGTA_vs_YPD") %>%
  dplyr::left_join(gene_info_df,
                   by = "gene") %>% 
  dplyr::select(gene, 
                geneName, 
                baseMean, 
                log2FC = estimate, 
                stderror,
                padj = p.adjusted)

```



### List DEGs most UP in EGTA vs YPD in delta at 37°C

We cut off at 2x changes with 5% FDR here.

```{r DEGs_EGTA_up_delta, fig.width = 6, fig.height = 3}
# Make data frames of DEGs up and down, arranged by log2FC
DEGdf_EGTA_delta_up2x_FDR5 <- 
  dplyr::filter(deseq_df_EGTAvsYPD_delta_37C,
                log2FC > 1, padj < 0.05) %>%
  dplyr::arrange(desc(log2FC))
print(DEGdf_EGTA_delta_up2x_FDR5, n = 20)
```

Print the top 20 *named* DEGs

```{r DEGs_EGTA_up_named_delta}
DEGdf_EGTA_delta_up2x_FDR5 %>%
  dplyr::filter(!is.na(geneName)) %>%
  print(n = 20)
```


### List DEGs most DOWN in EGTA vs YPD in delta at 37°C


```{r DEGs_EGTA_down_delta, fig.width = 6, fig.height = 3}
# Make data frames of DEGs up and down, arranged by log2FC
DEGdf_EGTA_delta_down2x_FDR5 <- 
  dplyr::filter(deseq_df_EGTAvsYPD_delta_37C,
                log2FC < -1, padj < 0.05) %>%
  dplyr::arrange(log2FC)
print(DEGdf_EGTA_delta_down2x_FDR5, n = 20)
```

Print the top 20 *named* DEGs

```{r DEGs_30C_down_named_delta}
DEGdf_EGTA_delta_down2x_FDR5 %>%
  dplyr::filter(!is.na(geneName)) %>%
  print(n = 20)
```

### Volcano plot of DEGs

```{r volcano_EGTAvsYPD_delta_37C, fig.width = 6, fig.height = 3}
volcano_EGTAvsYPD_delta_37C <- 
  ggplot(data = deseq_df_EGTAvsYPD_delta_37C,
         aes(x = log2FC, y = -log10(padj)) ) +
  geom_point(size = 0.3, colour = "grey70") +
  geom_point(data = DEGdf_EGTA_delta_up2x_FDR5,
             size = 0.3, colour = "darkblue") +
  geom_point(data = DEGdf_EGTA_delta_down2x_FDR5,
             size = 0.3, colour = "darkred") +
  scale_y_continuous("-log10(p)",
                     limits = c(0,50), expand = c(0,0),
                     oob=scales::squish) +
  scale_x_continuous("log2 fold-change\n← up in YPD       up in YPD + EGTA →",
                     breaks = -4:4,
                     limits = c(-4,4), expand = c(0.01,0.01),
                     oob = scales::squish) +
  coord_cartesian(clip = "off") 
volcano_EGTAvsYPD_delta_37C

ggsave(here("EGTAvsYPD_delta_37C_volcano.png"), volcano_EGTAvsYPD_delta_37C, width = 4, height = 3)
```


### Write outputs to files

```{r write_DEGs_delta}
deseq_df_EGTAvsYPD_delta_37C %>%
  dplyr::mutate(
    across(c(baseMean, log2FC, stderror), round, digits = 3)
  ) %>%
  dplyr::mutate(padj = signif(padj, digits = 4)) %>%
  write_tsv(file = here::here("results", "DEGs",
                              "deseq_df_EGTAvsYPD_delta_37C.txt"))

DEGdf_EGTA_delta_up2x_FDR5 %>%
  dplyr::pull(gene) %>%
  write_lines(file = here::here("results", "DEGs",
                            "deglist_EGTAvsYPD_delta_37C_up2x_FDR5.txt"))

DEGdf_EGTA_delta_down2x_FDR5 %>%
  dplyr::pull(gene) %>%
  write_lines(file = here::here("results", "DEGs",
                            "deglist_EGTAvsYPD_delta_37C_down2x_FDR5.txt"))
```



```{r}
# Turn into lists of genes
EGTA_up_WT <- DEGdf_EGTA_WT_up2x_FDR5$gene
EGTA_up_delta <- DEGdf_EGTA_delta_up2x_FDR5$gene

# Add to a list object with names to display on plot
euler_list_EGTA_up <- list( 
  'FLC1' = EGTA_up_WT , 
  'flc1∆' = EGTA_up_delta)

# Create the euler plot
eulerr_plot_EGTA_up <- plot(
  euler(euler_list_EGTA_up), 
  quantities = TRUE, 
  legend = TRUE,
  fills = c("lightblue2", "#F9E9C8"),
  fontsize = 20
  ) 

# Display eulerr 
eulerr_plot_EGTA_up

png(file=here("eulerr_plot_EGTA_up.png"))
eulerr_plot_EGTA_up
dev.off()
```

```{r}
# Turn into lists of genes
EGTA_down_WT <- DEGdf_EGTA_WT_down2x_FDR5$gene
EGTA_down_delta <- DEGdf_EGTA_delta_down2x_FDR5$gene

# Add to a list object with names to display on plot
euler_list_EGTA_down <- list( 
  'FLC1' = EGTA_down_WT , 
  'flc1∆' = EGTA_down_delta)

# Create the euler plot
eulerr_plot_EGTA_down <- plot(
  euler(euler_list_EGTA_down), 
  quantities = TRUE, 
  legend = TRUE,
  fills = c("grey95", "#F9E9C8"),
  fontsize = 20
  ) 

# Display eulerr 
eulerr_plot_EGTA_down

png(file=here("eulerr_plot_EGTA_down.png"))
eulerr_plot_EGTA_down
dev.off()
```


