---
title: "Differential Gene Expression of FLC1 and EGTA effects with interactions"
author: "Edward Wallace"
date: "2024-12-06"
output: 
  html_document:
    toc: true
---


# Summary

Differential Gene Expression of WT vs delta and EGTA vs none with interaction, at 37°C in YPD samples, using DESeq2,

This analysis produces files focused on the interaction term:

- `deseq_df_FLC1EGTAinteraction.txt`
- `deglist_FLC1EGTAinteraction_up2x_FDR5.txt`
- `deglist_FLC1EGTAinteraction_down2x_FDR5.txt`

This analysis confirms that there is a significant interaction between the strain and the effect of EGTA. Many genes are differently effect by EGTA in the presence or absence of FLC1. There are noticeable signals of:

- genes involved in divalent cations (CTR4, FRE, etc.)
- chitin synthesis (CHS6, CHS7)
- some secreted or cell surface proteins (CIG1, etc)
- transporters (HXS1, etc)


## Load count data from shared script load_count_data.Rmd

```{r call_load_count_data, child = "load_count_data.Rmd"}
```

# DESeq on a subset of data, only 37C in YPD

```{r subset_FLC1EGTAinteraction, fig.height=8, fig.width=8}
sample_sheet_FLC1EGTAinteraction <-
  dplyr::filter(sample_sheet,
                Temperature == "37", Media %in% c("YPD", "YPD_EGTA")) %>%
  dplyr::mutate(EGTA = factor(Media, 
                              levels = c("YPD","YPD_EGTA"),
                              labels = c("No","Yes")))

dds_FLC1EGTAinteraction <- 
  DESeqDataSetFromMatrix(
    countData = counts_all %>%
      dplyr::select(sample_sheet_FLC1EGTAinteraction$SampleName) %>%
      magrittr::set_rownames(counts_all$Gene),
    colData = sample_sheet_FLC1EGTAinteraction,
    design = ~ FLC1 + EGTA + FLC1:EGTA) %>%
  DESeq()

dds_FLC1EGTAinteraction
```
```{r results_FLC1EGTAinteraction}
results(dds_FLC1EGTAinteraction)

resultsNames(dds_FLC1EGTAinteraction)
```

```{r deseq_df_FLC1EGTAinteraction}
deseq_df_FLC1EGTAinteraction <- 
  biobroom::tidy.DESeqDataSet(dds_FLC1EGTAinteraction) %>%
  dplyr::filter(term == "FLC1WT.EGTAYes") %>%
  dplyr::left_join(gene_info_df,
                   by = "gene") %>% 
  dplyr::select(gene, 
                geneName, 
                baseMean, 
                log2FC = estimate, 
                stderror,
                padj = p.adjusted)
```

## List DEGs most UP in WT vs delta

We cut off at 2x changes with 5% FDR here.


```{r DEGs_WT_up, fig.width = 6, fig.height = 3}
# Make data frames of DEGs up and down, arranged by log2FC
DEGdf_interact_up2x_FDR5 <- 
  dplyr::filter(deseq_df_FLC1EGTAinteraction,
                log2FC > 1, padj < 0.05) %>%
  dplyr::arrange(desc(log2FC))
print(DEGdf_interact_up2x_FDR5, n = 20)
```

Print the top 20 *named* DEGs

```{r DEGs_WT_up_named}
DEGdf_interact_up2x_FDR5 %>%
  dplyr::filter(!is.na(geneName)) %>%
  print(n = 20)
```


## List DEGs most DOWN in WT vs delta


```{r DEGs_WT_down, fig.width = 6, fig.height = 3}
# Make data frames of DEGs up and down, arranged by log2FC
DEGdf_interact_down2x_FDR5 <- 
  dplyr::filter(deseq_df_FLC1EGTAinteraction,
                log2FC < -1, padj < 0.05) %>%
  dplyr::arrange(log2FC)
print(DEGdf_interact_down2x_FDR5, n = 20)
```

Print the top 20 *named* DEGs

```{r DEGs_37C_down_named}
DEGdf_interact_down2x_FDR5 %>%
  dplyr::filter(!is.na(geneName)) %>%
  print(n = 20)
```

## Volcano plot of DEGs

ADD volcano plots for the other terms as sanity check?

```{r volcano_FLC1EGTAinteraction, fig.width = 6, fig.height = 3}
volcano_FLC1EGTAinteraction <- 
  ggplot(data = deseq_df_FLC1EGTAinteraction,
         aes(x = log2FC, y = -log10(padj)) ) +
  geom_point(size = 0.5, colour = "grey50") +
  geom_point(data = DEGdf_interact_up2x_FDR5,
             size = 1, colour = "darkblue") +
  geom_point(data = DEGdf_interact_down2x_FDR5,
             size = 1, colour = "darkred") +
  scale_y_continuous("-log10(p)",
                     limits = c(0,20), expand = c(0,0),
                     oob=scales::squish) +
  scale_x_continuous("log2 fold-change,\nInteraction term FLC1 and EGTA",
                     breaks = -5:5,
                     limits = c(-4,4), expand = c(0.01,0.01),
                     oob = scales::squish) +
  coord_cartesian(clip = "off") 
volcano_FLC1EGTAinteraction
```

The volcano plot shows extensive differential effects of EGTA in flc1∆ compared to wild-type cells.



## Write outputs to files

```{r write_DEGs}
deseq_df_FLC1EGTAinteraction %>%
  dplyr::mutate(
    across(c(baseMean, log2FC, stderror), round, digits = 3)
  ) %>%
  dplyr::mutate(padj = signif(padj, digits = 4)) %>%
  write_tsv(file = here::here("results", "DEGs",
                              "deseq_df_FLC1EGTAinteraction.txt"))

DEGdf_interact_up2x_FDR5 %>%
  dplyr::pull(gene) %>%
  write_lines(file = here::here("results", "DEGs",
                            "deglist_FLC1EGTAinteraction_up2x_FDR5.txt"))

DEGdf_interact_down2x_FDR5 %>%
  dplyr::pull(gene) %>%
  write_lines(file = here::here("results", "DEGs",
                            "deglist_FLC1EGTAinteraction_down2x_FDR5.txt"))
```

# Session Info

Package versions, etc., that ran this file.

```{r session_info,results="show",message=TRUE}
sessionInfo()
```