---
title: "load_count_data"
author: "Edward Wallace"
date: "2024-12-04"
output: html_document
---

# Load and organise count data

`load_count_data.Rmd` is just to load and organise the raw count data in R, as well as loading libraries for data analysis.
Other scripts can just run this one instead of duplicating code, especially scripts for differential gene expression analysis.

This runs *after* `QC_PCA.Rmd`, which does the organisation.
This just loads the organised counts from the file `results/summaries/counts_bysamplecode.txt`.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(GGally)
library(here)
library(DESeq2)
library(biobroom)

theme_set(theme_cowplot(font_size = 12) +
    theme(# panel.grid.major = element_line(colour = "grey80", size = 0.5),
          strip.background = element_blank())
    )
```


## Load the organised counts

```{r load_counts}
counts_all <- 
  here::here("results", "summaries", "counts_bysamplecode.txt") %>%
  readr::read_tsv(comment = "#",
                  col_types = "ciiiiiiiiiiiiiiiiiiiiiiiiiiiiiii")   %>%
  magrittr::set_rownames(.$Gene)
counts_all
```

## Load the sample sheet

```{r load_samplesheet}
sample_sheet <- 
  here::here("input", "experiment",
             "CryptoRNAseqFLC1SampleKey.csv") %>%
  readr::read_csv(comment = "#", col_types = "ccccccc") %>%
  dplyr::mutate(Condition = stringr::str_sub(SampleName,start = 1L, end = -3L)) %>%
  dplyr::transmute(SampleID = as_factor(SampleID), 
                   SampleName = as_factor(SampleName),
                   Condition = as_factor(Condition),
                   FLC1 = factor(FLC1, levels = c("delta", "WT")),
                   Temperature = factor(Temperature, levels = c("30", "37")),
                   Media = as_factor(Media),
                   BioRep = as_factor(BioRep))

sample_sheet 
```

## Load H99 Gene information

```{r load_gene_info}
gene_info_df <- 
  here::here("input", "annotation",
             "H99_geneinfo_FungiDB59.txt") %>%
  readr::read_tsv(comment = "#")  %>% 
  dplyr::rename(gene = GeneID, geneName = GeneName)
```