---
title: "Quality control and PCA for FLC1 RNA-seq"
author: "Edward Wallace"
date: "2024-12-04"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(GGally)
library(here)
library(DESeq2)
library(biobroom)

theme_set(theme_cowplot(font_size = 12) +
    theme(# panel.grid.major = element_line(colour = "grey80", size = 0.5),
          strip.background = element_blank())
    )

# Count header columns from featurecounts
count_header_columns <- c("Geneid", "Chr", "Start",
                          "End", "Strand", "Length")

# unwanted string in filename; may need to be edited
unwanted_filename_string1 <- "11308_"
unwanted_filename_string2 <- "_S{[0-9]}+_R2_001_aln.bam"

```

# Summary

Quality control of RNA-seq data after read processing.

This analysis does:

- Scatter plot of counts showing log-scale correlations
- PCA plot showing how samples are similar to one another

It concludes that:

- Replicates group together in PCA space, so the experiment is reproducible
- PC1 contrasts WT with FLC1delta, capturing about 65% of variance
- PC2 contrasts temperature (for WT) and EGTA treatment (more for FLC1delta)
- There is a lot more to be understood here.

# Load and organise raw count data

## Load the raw counts

```{r load_counts}
counts_raw <- 
  here::here("results", "counts", "counts.txt") %>%
  readr::read_tsv(comment = "#")
counts_raw
```

## Load the sample sheet

```{r load_samplesheet}
sample_sheet <- 
  here::here("input", "experiment",
             "CryptoRNAseqFLC1SampleKey.csv") %>%
  readr::read_csv(comment = "#", col_types = "ccccccc") %>%
  dplyr::mutate(Condition = stringr::str_sub(SampleName,start = 1L, end = -3L)) %>% 
  dplyr::mutate(Media = factor(Media,
                               levels = c("YPD","YPD_CFW", "YPD_EGTA", "YPD_CFW_EGTA"),
                               labels = c("YPD", "YPD + CFW", "YPD + EGTA", "YPD + EGTA + CFW")
                                ),
                Temperature = factor(Temperature,
                                     levels = c("30", "37"),
                                     labels = c("30°C", "37°C")))

sample_sheet 
```

## Organise the counts using the sample sheet

```{r organise_counts}
counts_all <- 
  counts_raw %>%
  dplyr::rename_with(stringr::word,
                     start = 2L, sep = fixed("_"), .cols = ends_with(".bam")) %>%
  dplyr::select(sample_sheet$SampleID)  %>%
  magrittr::set_colnames(sample_sheet$SampleName)  %>%
  magrittr::set_rownames(counts_raw$Geneid)

counts_all
```

## Write out counts with sample codes as column names

```{r write_counts_samplenames}
bind_cols(
  counts_raw %>% 
    dplyr::select(Gene = Geneid, Length = Length),
  counts_all 
) %>%
  readr::write_tsv(
    file = here::here("results", "summaries", "counts_bysamplecode.txt"))
```


# Check reproducibility by paired scatter plot in log10-plus-1-space

_Note: This is set to `eval=FALSE` as it can run very slowly for large datasets._ 

```{r paired_scatter, fig.height=20, fig.width=20, eval=FALSE}
#counts_all %>%
 # magrittr::add(1) %>%
  #log10() %>%
  #GGally::ggpairs()

#here::here("results", "figures","paired_scatter_log10countsplusone_all_samples.png") %>%
#ggsave(width = 32, height = 32)
```

# Normalize counts by regularized log (rlog) with DESeq2

The [rlog function from DESeq2](https://rdrr.io/bioc/DESeq2/man/rlog.html) "transforms the count data to the log2 scale in a way which minimizes differences between samples for rows with small counts, and which normalizes with respect to library size."

## Construct DESeq dataset object

EDIT: the `design` formula may need to be edited.

```{r dds_counts_all}
dds_all <- DESeqDataSetFromMatrix(countData = counts_all,
                                  colData = sample_sheet,
                                  design = ~ SampleName)
```

# Do the rlog transformation


```{r rlog_dds_all}
rlog_all <- rlog(dds_all)
head(assay(rlog_all))
```

# Check reproducibility by principal component analysis (PCA) of rlog

## Calculate PCs from the rlog output

Principal component analysis focused on comparing samples to other samples, hence take the transpose.

```{r calculate_pca_rlog}
# calculate principal components of rlog, after extracting from the dataset
pca_rlog <- rlog_all %>%
  assay() %>%
  t() %>%
  prcomp()

# convert principal components to data frame
pcdf_rlog_df <- bind_cols(
  as_tibble(colData(rlog_all)),
  as_tibble(pca_rlog$x)
) %>% 
  mutate(FLC1 = factor(FLC1, levels = c("WT", "delta"), labels = c("*FLC1*", "*flc1*∆")))

pcdf_rlog_df 
```

```{r}
pc_loadings <- pca_rlog$rotation

pc_loadings <- pc_loadings %>% 
  as_tibble(rownames = "gene")

# print the result
pc_loadings


top_genes <- pc_loadings %>% 
  # select only the PCs we are interested in
  select(gene, PC1, PC2, PC3) %>%
  # convert to a "long" format
  pivot_longer(matches("PC"), names_to = "PC", values_to = "loading") %>% 
  # for each PC
  group_by(PC) %>% 
  # arrange by descending order of loading
  arrange(desc(abs(loading))) %>% 
  # take the 10 top rows
  slice_head(n = 10) #%>% 
  # pull the gene column as a vector
  #pull(gene) %>% 
  # ensure only unique genes are retained
 # unique()

top_loadings <- pc_loadings %>% 
  filter(gene %in% top_genes$gene)

loadings_plot_base <- ggplot(data = top_loadings) + 
  geom_hline(yintercept = 0, size = 0.5, colour = "grey90") +
  geom_vline(xintercept = 0, size = 0.5, colour = "grey90") 

loadings_plot_PC1_PC2 <- loadings_plot_base +
  geom_segment(aes(x = 0, y = 0, xend = PC1, yend = PC2), 
               arrow = arrow(length = unit(0.1, "in")),
               colour = "grey70") +
  geom_text_repel(aes(x = PC1, y = PC2, label = gene),
            nudge_y = 0.005, size = 3, colour = "blue", max.overlaps = 20) +
  scale_x_continuous(expand = c(0.02, 0.02)) + xlab("PC1") + ylab("PC2") 

loadings_plot_PC1_PC3 <- loadings_plot_base +
  geom_segment(aes(x = 0, y = 0, xend = PC1, yend = PC3), 
               arrow = arrow(length = unit(0.1, "in")),
               colour = "grey70") +
  geom_text_repel(aes(x = PC1, y = PC3, label = gene),
            nudge_y = 0.005, size = 3, colour = "blue", , max.overlaps = 20) +
  scale_x_continuous(expand = c(0.02, 0.02)) + xlab("PC1") + ylab("PC3")

loadings_plot_PC2_PC3 <- loadings_plot_base +
  geom_segment(aes(x = 0, y = 0, xend = PC2, yend = PC3), 
               arrow = arrow(length = unit(0.1, "in")),
               colour = "grey70") +
  geom_text_repel(aes(x = PC2, y = PC3, label = gene),
            nudge_y = 0.005, size = 3, colour = "blue", , max.overlaps = 20) +
  scale_x_continuous(expand = c(0.02, 0.02)) + xlab("PC2") + ylab("PC3") 

plot_grid(loadings_plot_PC1_PC2, "",
          loadings_plot_PC1_PC3, loadings_plot_PC2_PC3,
          axis = "tblr", align = "hv") 

#ggsave(
 # here("results/figures/PC1vsPC2vsPC3_loadings_plot.png"), width = 10, height = 9, dpi = 1000, bg = "white"
#)
```



## Calculate proportion of variance explained by each PC


```{r calculate_propvar}
propvar_rlog_df <- tibble(
  PC = seq.int(1L, ncol(pca_rlog$x) ),
  prop_var = pca_rlog$sdev^2/sum(pca_rlog$sdev^2)
)
```

## Plot PCA percentage variance explained to find how many PCs to pay attention to.

```{r plot_percentvar, fig.width = 3, fig.height = 2}
plot_percentvar_rlog <- 
  ggplot(data = propvar_rlog_df, 
         aes(x = PC, y = prop_var)) +
  geom_col(fill = "grey20") +
  scale_x_continuous("principal component",
                     limits = c(0.4,10.6), 
                     # hack limits to include the bars and a little space.
                     breaks = 1L:10L,
                     expand = c(0,0)) + 
  scale_y_continuous("prop. of variance", expand = c(0,0))
plot_percentvar_rlog

ggsave(
  here("results/figures/percentvar_rlog.png"), width = 3, height = 2, dpi = 1000
)
```

## Plot PC1 vs PC2 with default values

EDIT: the colour and other aesthetics may need to be edited to highlight features of the data

```{r plot_PC12_default, fig.height=5, fig.width=6, eval = FALSE}
ggplot(data = pcdf_rlog_df,
       aes(colour = interaction(Temperature,Media), 
           shape = FLC1)
       ) +
  scale_shape_manual(values = c(20, 2)) +
  scale_color_manual("Temp., Media", values = c("darkblue","cornflowerblue", "red3", "orange", "darkolivegreen")) +
  #scale_colour_brewer("Temp., Media", 
   #                   type = "qual", palette = 7) + 
  geom_point(aes(x = PC1, y = PC2), size = 3) + 
  xlab("PC1, 64.8% var.") + 
  ylab("PC2, 11.6% var.")

#ggsave(
 # here("results/figures/PC1vsPC2.png"), width = 5.5, height = 2.5, dpi = 1000
  #)
```

## Plot PC1 vs PC2, 3, 4, also with default values

```{r plot_PC1234_text, fig.height=10, fig.width=8}

Design_tbl <- tibble(
  FLCstatus = rep(c("*FLC1*", "*flc1*∆"), each = 5),
  Condition = rep(c("30°C\nYPD\n \n ", "37°C\nYPD\n \n ", "37°C\nYPD\n+ CFW\n ", "37°C\nYPD\n+ EGTA\n ", "37°C\nYPD\n+ CFW\n+ EGTA"), times = 2)
) %>% 
  mutate(Condition = factor(Condition, levels = c("30°C\nYPD\n \n ", "37°C\nYPD\n \n ", "37°C\nYPD\n+ CFW\n ", "37°C\nYPD\n+ EGTA\n ", "37°C\nYPD\n+ CFW\n+ EGTA")),
         FLCstatus = factor(FLCstatus, levels = rev(c("*FLC1*", "*flc1*∆"))))


Design_plot <- ggplot(data = Design_tbl, aes(y = FLCstatus, x = Condition)) +
  scale_x_discrete(position = 'top') +
  #ggtitle(expression(underline('     3 hr, 180 rpm    '))) +
  theme(axis.text.y = element_markdown(),
        plot.title = element_text(hjust = 0.5, face = "plain", size = 12),
        axis.text.x = element_text(hjust=1),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(), 
        legend.position = "none")

# This could be wraped in a function
pc_baseplot <- 
  ggplot(data = pcdf_rlog_df,
       aes(colour = interaction(Temperature,Media), 
           shape = FLC1)
       ) +
  scale_shape_manual(values = c(20, 2)) +
  scale_color_manual("Temp., Media", values = c("darkblue","cornflowerblue", "red3", "orange", "darkolivegreen")) +
  #scale_colour_brewer("Temp., Media", 
   #                   type = "qual", palette = 2) + 
  geom_hline(yintercept = 0, size = 0.5, colour = "grey80") +
  geom_vline(xintercept = 0, size = 0.5, colour = "grey80") +
  theme(legend.position = "none")

pcplot_12 <- pc_baseplot + geom_point(aes(x = PC1, y = PC2), size = 3) + xlab("PC1, 64.8% var.") + ylab("PC2, 11.6% var.")
pcplot_13 <- pc_baseplot + geom_point(aes(x = PC1, y = PC3), size = 3) + xlab("PC1, 64.8% var.") + ylab("PC3, 6.3% var.")
#pcplot_14 <- pc_baseplot + geom_point(aes(x = PC1, y = PC4))
pcplot_23 <- pc_baseplot + geom_point(aes(x = PC2, y = PC3), size = 3) + xlab("PC2, 11.6% var.") + ylab("PC3, 6.3% var.")
#pcplot_24 <- pc_baseplot + geom_point(aes(x = PC2, y = PC4))
#pcplot_34 <- pc_baseplot + geom_point(aes(x = PC3, y = PC4))

pcplot_legend <- get_legend(
  pcplot_12 + 
    theme(legend.position = "right", legend.margin = margin(t = 0, r = 0, b = 20, l = 40, unit = "pt"),
          legend.spacing.y = unit(0.0, 'cm'),
          legend.text = element_markdown()) + 
  guides(shape = guide_legend("FLC1 gene", order=2),
         colour = guide_legend(order=1))
)


grid <-plot_grid(Design_plot, plot_percentvar_rlog, rel_widths = c(1, 0.6), axis = 'b', align = "h")

grid2 <- plot_grid(
          pcplot_12, pcplot_legend,
          pcplot_13, pcplot_23)

grid3 <- plot_grid(
  grid,
  grid2,
  rel_heights = c(1, 3),
  axis = "lr",
  ncol = 1
)

plot_grid(grid, NULL, 
          pcplot_12, pcplot_legend,
          pcplot_13, pcplot_23,
          axis = "tblr", align = "hv",
          rel_heights = c(1.5, 2, 2, 1.5, 2, 2),
          ncol = 2)

ggsave(
  here("results/figures/PC1vsPC2vsPC3.png"), width = 8, height = 8, dpi = 1000
)
```

# Session Info

Package versions, etc., that ran this file.

```{r session_info,results="show",message=TRUE}
sessionInfo()
```
