---
title: "RM_thesis_WTvsdelta"
author: "Rachael Murray"
date: "2025-04-22"
output: html_document
---
# Summary

In this document I explore differences in the transciptome of KN99alpha cells (FLC1) and flc1∆ cells grown in YPD at 30°C, YPD at 37°C and YPD + EGTA at 37°C. 
Differential gene expression analysis was previously performed (see Rmd/DGE_Wtvsdelta_30C_YPD.rmd, Rmd/DGE_Wtvsdelta_37C_YPD.rmd, Rmd/DGE_Wtvsdelta_37C_EGTA.rmd) and GO term enrichment analysis was carried out at fungidb.com. Here I make nice figures for my PhD thesis. 

Input files: 
- results/DEGs/deseq_df_WTvsdelta_30C_YPD.txt
- results/DEGs/deseq_df_WTvsdelta_37C_YPD.txt
- results/DEGs/deseq_df_WTvsdelta_37C_EGTA.txt
- results/GOanalysis/GO_WTvsdelta_30C_YPD/
- results/GOanalysis/GO_WTvsdelta_37C_YPD/
- results/GOanalysis/GO_WTvsdelta_37C_EGTA/
- results/gene_info/genes_to_highlight.csv

# Load count data from shared script load_count_data.Rmd

```{r call_load_count_data, child = "load_count_data.Rmd"}
```


# FLC1 vs flc1∆ at 30°C and 37°C 

## Volcano plots

Open results from differential gene expression analysis and create volcano plot of FLC1 vs flc1∆ at 30°C and FLC1 vs flc1∆ at 37°C. Differentially expressed gene cutoff = log2-fold change > 1 and pvalue < 0.05. 

Open results from differential gene expression analysis. 
```{r Open DEseq results from FLC1 vs flc1∆ at 30°C and 37°C}
# FLC1 vs flc1∆ in YPD at 30°C 
deseq_df_WTvsdelta_30C_YPD <- read_tsv(
 file = here::here("results", "DEGs",
                              "deseq_df_WTvsdelta_30C_YPD.txt"))
# Number up in flc1∆ at 30°C 
DEGs_down_30 <- deseq_df_WTvsdelta_30C_YPD %>% 
                     dplyr::filter(log2FC < -1, padj < 0.05) %>% 
                     arrange(log2FC)
# Number up in FLC1 at 30°C %>% 
DEGs_up_30 <- deseq_df_WTvsdelta_30C_YPD %>% 
                     dplyr::filter(log2FC > 1, padj < 0.05) %>% 
                     arrange(log2FC)
  
# FLC1 vs flc1∆ in YPD at 37°C 
deseq_df_WTvsdelta_37C_YPD <- read_tsv(
 file = here::here("results", "DEGs",
                              "deseq_df_WTvsdelta_37C_YPD.txt"))

# Number up in flc1∆ at 37°C 
DEGs_down_37 <- deseq_df_WTvsdelta_37C_YPD %>% 
                     dplyr::filter(log2FC < -1, padj < 0.05) %>% 
                     arrange(log2FC)

# Number up in FLC1 at 37°C %>% 
DEGs_up_37 <- deseq_df_WTvsdelta_37C_YPD %>% 
                     dplyr::filter(log2FC > 1, padj < 0.05) %>% 
                     arrange(log2FC)

# FLC1 vs flc1∆ in YPD + EGTA at 37°C 
deseq_df_WTvsdelta_37C_EGTA <- read_tsv(
 file = here::here("results", "DEGs",
                              "deseq_df_WTvsdelta_37C_EGTA.txt"))

# Number up in flc1∆ YPD + EGTA at 37°C 
DEGs_down_37_EGTA <- deseq_df_WTvsdelta_37C_EGTA %>% 
                     dplyr::filter(log2FC < -1, padj < 0.05) %>% 
                     arrange(log2FC)

# Number up in FLC1 YPD + EGTA at 37°C %>% 
DEGs_up_37_EGTA <- deseq_df_WTvsdelta_37C_EGTA %>% 
                     dplyr::filter(log2FC > 1, padj < 0.05) %>% 
                     arrange(log2FC)
```
Take a quick look at most up in flc1∆ across all conditions 
```{r}
DEGs_down_37_new_col_name <- DEGs_down_37 %>% 
  select(1, 2, 4, 6) %>% 
  dplyr::rename(log2FC_37 = log2FC,
                padj_37 = padj) %>% 
  filter(gene %in% DEGs_down_30$gene) %>% 
  filter(gene %in% DEGs_down_37_EGTA$gene)

DEGs_down_30_new_col_name <- DEGs_down_30 %>% 
  select(1, 2, 4, 6) %>% 
  dplyr::rename(log2FC_30 = log2FC,
                padj_30 = padj) %>% 
  filter(gene %in% DEGs_down_37$gene) %>% 
  filter(gene %in% DEGs_down_37_EGTA$gene)


DEGs_down_EGTA_new_col_name <- DEGs_down_37_EGTA %>% 
  select(1, 2, 4, 6) %>% 
  dplyr::rename(log2FC_EGTA = log2FC,
                padj_EGTA = padj) %>% 
  filter(gene %in% DEGs_down_30$gene) %>% 
  filter(gene %in% DEGs_down_37$gene)

left_join(DEGs_down_30_new_col_name, DEGs_down_37_new_col_name) %>% 
  left_join(DEGs_down_EGTA_new_col_name) %>% 
  mutate(total = log2FC_30 + log2FC_37 + log2FC_EGTA) %>% 
  arrange(total) %>% 
  write_tsv(here("results", "DEGs_fungiDB_info", "Most_down_WTvsdelta.tsv"))


A <- DEGs_down_30 %>% slice_head(n = 20) 
B <- DEGs_down_37 %>% slice_head(n = 20) 
C <- DEGs_down_37_EGTA %>% slice_head(n = 20)

TOP_genes <- bind_rows(A, B, C) %>% 
  select(1, 2, 4, 6) 

X <- deseq_df_WTvsdelta_30C_YPD %>% 
  select(1, 2, 4, 6) %>% 
  dplyr::rename(log2FC_30 = log2FC,
                padj_30 = padj)

Y <- deseq_df_WTvsdelta_37C_YPD %>% 
  select(1, 2, 4, 6) %>% 
  dplyr::rename(log2FC_37 = log2FC,
                padj_37 = padj)

Z <- deseq_df_WTvsdelta_37C_EGTA %>% 
  select(1, 2, 4, 6) %>% 
  dplyr::rename(log2FC_EGTA  = log2FC,
                padj_EGTA  = padj)

left_join(X, Y) %>% 
  left_join(Z) %>% 
  filter(gene %in% TOP_genes$gene) %>% 
  arrange(log2FC_30) %>% 
  mutate(sig_30 = case_when((log2FC_30 < 1 & padj_30 < 0.05) ~ 'yes'),
         sig_37 = case_when((log2FC_37 < 1 & padj_37 < 0.05) ~ 'yes'),
         sig_EGTA = case_when((log2FC_EGTA < 1 & padj_EGTA < 0.05) ~ 'yes'),
         ) %>%
  write_tsv(here("results", "DEGs_fungiDB_info", "Most_down_WTvsdelta_each_condition_sig.tsv"))
  
```



Make volcano plots for each growth temperature. 
```{r Volcano plots for each comparison, fig.width = 4, fig.height=3}
geom_point_size <- 0.3
limits_x <- c(-6, 6)
colour_positive <- "blue3"
colour_negative <- "firebrick"
colour_neutral <- "grey70"

# Volcano YPD 30°C
volcano_WTvsdelta_30C_YPD <- 
 ggplot(data = deseq_df_WTvsdelta_30C_YPD,
         aes(x = log2FC, y = -log10(padj)) ) +
  geom_point(size = geom_point_size - 0.1, colour = colour_neutral) +
  geom_point(data = deseq_df_WTvsdelta_30C_YPD %>% 
               dplyr::filter(
                 log2FC  > 1, padj < 0.05),
             size = geom_point_size, colour = colour_positive) +
  geom_point(data = deseq_df_WTvsdelta_30C_YPD %>% 
               dplyr::filter(
                 log2FC < -1, padj < 0.05),
             size = geom_point_size, colour = colour_negative) +
  scale_y_continuous("-log10(p)",
                     limits = c(0,50), expand = c(0,0),
                     oob=scales::squish) +
  scale_x_continuous("log2 fold-change\n← up in flc1∆       up in FLC1 →",
                     breaks = -6:6,
                     limits = limits_x, expand = c(0.01,0.01),
                     oob = scales::squish) +
  coord_cartesian(clip = "off") + 
  ggtitle("*FLC1* vs *flc1*∆<br>YPD, 30°C") +
  theme(plot.title = element_markdown(colour = "black", size = 18, face = "plain"),
        axis.text.x = element_text(size = 14, colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black"),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.title.y = element_text(size = 14, colour = "black")) 

volcano_WTvsdelta_30C_YPD

# Volcano YPD 37°C
volcano_WTvsdelta_37C_YPD <- 
 ggplot(data = deseq_df_WTvsdelta_37C_YPD,
         aes(x = log2FC, y = -log10(padj)) ) +
  geom_point(size = geom_point_size - 0.1, colour = colour_neutral) +
  geom_point(data = deseq_df_WTvsdelta_37C_YPD %>% 
               dplyr::filter(
                 log2FC  > 1, padj < 0.05),
             size = geom_point_size, colour = colour_positive) +
  geom_point(data = deseq_df_WTvsdelta_37C_YPD %>% 
               dplyr::filter(
                 log2FC < -1, padj < 0.05),
             size = geom_point_size, colour = colour_negative) +
  scale_y_continuous("-log10(p)",
                     limits = c(0,50), expand = c(0,0),
                     oob=scales::squish) +
  scale_x_continuous("log2 fold-change\n← up in flc1∆       up in FLC1 →",
                     breaks = -6:6,
                     limits = limits_x, expand = c(0.01,0.01),
                     oob = scales::squish) +
  coord_cartesian(clip = "off") + 
  ggtitle("*FLC1* vs *flc1*∆<br>YPD, 37°C") +
  theme(plot.title = element_markdown(colour = "black", size = 18, face = "plain"),
        axis.text.x = element_text(size = 14, colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black"),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.title.y = element_text(size = 14, colour = "black")) 

volcano_WTvsdelta_37C_YPD

# Volcano YPD 37°C
volcano_WTvsdelta_37C_EGTA <- 
 ggplot(data = deseq_df_WTvsdelta_37C_EGTA,
         aes(x = log2FC, y = -log10(padj)) ) +
  geom_point(size = geom_point_size - 0.1, colour = colour_neutral) +
  geom_point(data = deseq_df_WTvsdelta_37C_EGTA %>% 
               dplyr::filter(
                 log2FC  > 1, padj < 0.05),
             size = geom_point_size, colour = colour_positive) +
  geom_point(data = deseq_df_WTvsdelta_37C_EGTA %>% 
               dplyr::filter(
                 log2FC < -1, padj < 0.05),
             size = geom_point_size, colour = colour_negative) +
  scale_y_continuous("-log10(p)",
                     limits = c(0,50), expand = c(0,0),
                     oob=scales::squish) +
 scale_x_continuous("log2 fold-change\n← up in flc1∆       up in FLC1 →",
                     breaks = -6:6,
                     limits = limits_x, expand = c(0.01,0.01),
                     oob = scales::squish) +
  coord_cartesian(clip = "off") + 
  ggtitle("*FLC1* vs *flc1*∆<br>YPD + EGTA, 37°C") +
  theme(plot.title = element_markdown(colour = "black", size = 18, face = "plain"),
        axis.text.x = element_text(size = 14, colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black"),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.title.y = element_text(size = 14, colour = "black")) 

volcano_WTvsdelta_37C_EGTA
```

## Overlap of differentially expressed genes at 30°C and 37°C in FLC1 vs flc1∆

I wanted to know if the same set of genes are differentially expressed genes in flc1∆ at 30°C and 37°C. I first made some venn diagrams but found that a stacked barplot looks much neater. 

Get genes list of up in flc1∆ at 30°C, up in flc1∆ at 37°C, up in FLC1 at 30°C and up in FLC1 at 37°C. 
```{r fig.width=10, fig.height=7}
# Get list of down genes (up in flc1∆) for each growth condition
DGE_30C_YPD_down <- deseq_df_WTvsdelta_30C_YPD %>% 
                     dplyr::filter(log2FC < -1, padj < 0.05) %>% 
                     .$gene 

DGE_37C_YPD_down <- deseq_df_WTvsdelta_37C_YPD %>% 
                     dplyr::filter(log2FC < -1, padj < 0.05) %>% 
                     .$gene

# Get list of up genes (up in FLC1) for each growth condition
DGE_30C_YPD_up <- deseq_df_WTvsdelta_30C_YPD %>% 
                     dplyr::filter(log2FC > 1, padj < 0.05) %>% 
                     .$gene 

DGE_37C_YPD_up <- deseq_df_WTvsdelta_37C_YPD %>% 
                     dplyr::filter(log2FC > 1, padj < 0.05) %>% 
                     .$gene
```

Plot Venn diagrams
```{r Venn diagram, fig.width=5, fig.height=10}
eulerr_list_down <- list( 
  'YPD, 30°C' = DEGs_down_30 %>% .$gene,
  'YPD, 37°C' = DEGs_down_37 %>% .$gene, 
  'YPD + EGTA, 37°C' = DEGs_down_37_EGTA %>% .$gene)

eulerr_list_up <- list( 
  'YPD, 30°C' = DEGs_up_30 %>% .$gene,
  'YPD, 37°C' = DEGs_up_37 %>% .$gene, 
  'YPD + EGTA, 37°C' = DEGs_up_37_EGTA %>% .$gene)

eulerr_plot_down <- plot(
  euler(eulerr_list_down), 
  quantities = list(col = "firebrick"), 
  fills = list(fill = c("grey90", "grey40", "orange"), alpha = 0.4),
  edges = F,
  legend = F,
  fontsize = 20, 
  labels = list(col = "firebrick", font = 3, cex = 1)
  ) 
eulerr_plot_down

eulerr_plot_up <- plot(
  euler(eulerr_list_up), 
  quantities = list(col = "blue3"), 
  fills = list(fill = c("grey90", "grey40", "orange"), alpha = 0.4),
  edges = F,
  legend = F,
  fontsize = 20, 
  labels = list(col = "blue3", font = 3, cex = 1)
  ) 
eulerr_plot_up

eulerr_plots <- plot_grid(
  eulerr_plot_down, 
  eulerr_plot_up,
  ncol = 2
)


ggvenn_plot_down <- ggvenn(
  eulerr_list_down,
  show_percentage = F,
  fill_color = c("grey95","grey95","grey95"),
  fill_alpha = 0.4,
  set_name_size = 5,
  text_color = "firebrick",
  text_size = 5,
  stroke_color = "firebrick",
  stroke_size = 0.5
) + ggtitle("Up in *flc1*∆") + 
  theme(plot.title = element_markdown(hjust = 0.5, face = 'plain', size = 18))
ggvenn_plot_down

ggvenn_plot_up <- ggvenn(
  eulerr_list_up,
  show_percentage = F,
  fill_color = c("grey95","grey95","grey95"),
  fill_alpha = 0.4,
  set_name_size = 5,
  text_color = "blue3",
  text_size = 5,
  stroke_color = "blue3",
  stroke_size = 0.5
) + ggtitle("Up in *FLC1*") + 
  theme(plot.title = element_markdown(hjust = 0.5, face = 'plain', size = 18))

ggvenn_plot_up

cowplot::plot_grid(
  ggvenn_plot_down,
  ggvenn_plot_up,
  ncol = 1,
  scale = 0.8
)

```

8338 total transcripts 

((733+201)/8338)*100
((601+40)/8338)*100
((909+139)/8338)*100

```{r Percentage of genes DE in mutant in three condition, fig.width = 5, fig.height=6}
### Table showing the number of differentially expressed genes and 
DEGs_and_EGs <- tibble(
  condition = rep(c("YPD\n30°C", "YPD\n37°C", "YPD +\nEGTA\n37°C"), times = 3), 
  direction = rep(c("Up in *flc1*∆", "Up in *FLC1*", "Not differentially<br>expressed"), each = 3),
  gene_count = c(733, 610, 909, 201, 40, 139, 7404, 7688, 7290)
) %>% 
  mutate(direction = factor(direction, levels = c("Up in *flc1*∆", "Up in *FLC1*", "Not differentially<br>expressed")))


DEGs_and_EGs_barplot <- ggplot(DEGs_and_EGs, aes(y = gene_count, x = condition, fill = direction)) + 
  geom_bar(stat="identity", position ="fill") + 
  scale_fill_manual(values = c("firebrick", "blue3", "grey90")) + 
  ylab("Proportion of Genes") + 
  xlab("") +
  scale_y_continuous(expand = c(0.001,0.001)) + 
  theme(axis.text.y = element_markdown(size = 16),
        legend.title = element_blank(),
        axis.text = element_text(size = 14, colour = "black"),
        axis.title.y = element_text(size = 14, colour = "black"),
        legend.text = element_markdown(size = 12, colour = "black"),
        legend.spacing.y = unit(0.3, 'cm'),
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm'),
        plot.margin = margin(b = -0.4, unit = 'cm'),
        legend.justification = "top")

DEGs_and_EGs_barplot


DEGs_and_EGs_table <- DEGs_and_EGs %>% 
  filter(direction != "Not differentially<br>expressed") %>%
  mutate(direction = factor(direction, 
                            levels = c("Up in *FLC1*", "Up in *flc1*∆"), 
                            labels = c("DEGs up in *FLC1*", "DEGs Up in *flc1*∆"))
  )


DEGs_and_EGs_table_plot <- 
ggplot(data = DEGs_and_EGs_table,
       aes(x=condition, y=direction, label=gene_count)) +
  geom_tile(fill = "white", colour = "black", linewidth = 0.5) + 
  geom_text(aes(colour = direction), size = 4) + 
  scale_colour_manual(values = c("blue3", "firebrick")) + 
xlab("") +
ylab("No. DEGs") +
theme(legend.position="none",
panel.grid.major.x = element_blank(),
panel.grid.major.y = element_blank(),
panel.grid.minor.x = element_line(),
panel.grid.minor.y = element_line(),
axis.title.y = element_text(size = 14),
axis.line = element_blank(),
axis.ticks = element_blank(),
axis.text.y=element_blank(),
axis.text.x=element_blank(),
panel.ontop = TRUE,
axis.ticks.x=element_blank(),
panel.background = element_rect(fill = "transparent",colour = NA)
)


DEGs_and_EGs_grid <- plot_grid(
  DEGs_and_EGs_barplot, 
  DEGs_and_EGs_table_plot,
  ncol =1,
  align = "v", rel_heights = c(4, 1)
)
```



Stakced barplot 
```{r Barplot, fig.width = 4, fig.height=5}
### Barplot Venn alternative 
DEG_counts <- tibble(
  condition = rep(c("YPD, 30°C only", "YPD, 37°C only", "YPD, 30°C & 37°C"), times = 2), 
  direction = rep(c("up in flc1∆", "up in FLC1"), each = 3),
  count = c(276, 153, 457, 172, 11, 29)
) %>% 
  mutate(
    condition = factor(condition, levels = c("YPD, 30°C only", "YPD, 37°C only", "YPD, 30°C & 37°C")),
    direction = factor(direction, levels = c("up in flc1∆", "up in FLC1"))
  ) 

# Horizonatl bars 
DEG_barplot_30C_37C <- ggplot(DEG_counts, aes(y = direction, x = count, fill = condition)) + 
  geom_bar(stat="identity", colour = "black") + 
  scale_fill_manual(values = c("grey95", "grey65", "grey30")) + 
  xlab("DEGs") + 
  ylab("") +
  scale_x_continuous(limits = c(0,900), expand = c(0.01,0.01)) + 
  theme(axis.text.y = element_markdown(size = 16, colour = c("blue3", "firebrick")),
        legend.title = element_blank(),
        axis.text.x = element_text(size = 12, colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black"),
        legend.text = element_text(size = 14, colour = "black"),
        legend.spacing.y = unit(0.3, 'cm'),
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm'),
        legend.justification = "top")

#DEG_barplot_30C_37C

# Vertical bars
DEG_barplot_30C_37C <- ggplot(DEG_counts, aes(x = direction, y = count, fill = condition)) + 
  geom_bar(stat="identity", width = 0.7, colour = "black") + 
  scale_fill_manual(values = c("grey98", "grey65", "grey30")) + 
  ylab("Number of DEGs") + 
  xlab("") +
  scale_y_continuous(limits = c(0,900), expand = c(0.01,0.01)) + 
  theme(axis.text.x = element_markdown(size = 14, colour = c("firebrick", "blue3"), angle = 45, hjust =1),
        legend.title = element_blank(),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.title.y = element_text(size = 14, colour = "black"),
        legend.text = element_text(size = 14, colour = "black"),
        legend.key.size = unit(0.5,"line"),
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm'),
        legend.justification = "top") 

DEG_barplot_30C_37C

# Percentages useful for adding to chapter results
# Total up in flc1∆ = 886 
# Both = 51.6%
# 30°C = 31.1%
# 37°C = 17.3%
# Total up in FLC1 = 212
# Both = 29/212 = 13.7%
# 30°C = 81.1% 
# 37°C = 29/212 = 5.2%
```
## GO term analysis
GO term analysis FLC1 vs flc1∆ in YPD, 30°C and FLC1 vs flc1∆ in YPD, 37°C. 

```{r Open GO analysis results}
# Open GO analysis for down regulated gene in WT vs flc1∆ at 30°C
MF_slim_30_down <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_30C_YPD/GO_WTvsdelta_30C_YPD_down2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_30C_YPD",
    GO_type = "MF",
    DEG = "down"
  )

BP_slim_30_down <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_30C_YPD/GO_WTvsdelta_30C_YPD_down2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_30C_YPD",
    GO_type = "BP",
    DEG = "down"
  )

CC_slim_30_down <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_30C_YPD/GO_WTvsdelta_30C_YPD_down2x_FDR5_cellular_component_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_30C_YPD",
    GO_type = "CC",
    DEG = "down"
  )

GO_slim_30_down <- bind_rows(
  MF_slim_30_down,
  BP_slim_30_down,
  CC_slim_30_down) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)

GO_slim_30_down$name <- factor(GO_slim_30_down$name, levels = rev(GO_slim_30_down$name))
  

# Open GO analysis for downregulated gene in WT vs flc1∆ at 37°C
MF_slim_37_down <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_YPD/GO_WTvsdelta_37C_YPD_down2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_YPD",
    GO_type = "MF",
    DEG = "down"
  )

BP_slim_37_down <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_YPD/GO_WTvsdelta_37C_YPD_down2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_YPD",
    GO_type = "BP",
    DEG = "down"
  )

CC_slim_37_down <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_YPD/GO_WTvsdelta_37C_YPD_down2x_FDR5_cellular_component_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_YPD",
    GO_type = "CC",
    DEG = "down"
  )

GO_slim_37_down <- bind_rows(
  MF_slim_37_down,
  BP_slim_37_down,
  CC_slim_37_down) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)

GO_slim_37_down$name <- factor(GO_slim_37_down$name, levels = rev(GO_slim_37_down$name))

# Open GO analysis for downregulated gene in WT vs flc1∆ in EGTA at 37°C
MF_slim_EGTA_37_down <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_EGTA/GO_WTvsdelta_37C_EGTA_down2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_EGTA",
    GO_type = "MF",
    DEG = "down"
  )

BP_slim_EGTA_37_down <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_EGTA/GO_WTvsdelta_37C_EGTA_down2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_EGTA",
    GO_type = "BP",
    DEG = "down"
  )

CC_slim_EGTA_37_down <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_EGTA/GO_WTvsdelta_37C_EGTA_down2x_FDR5_cellular_component_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_EGTA",
    GO_type = "CC",
    DEG = "down"
  )

GO_slim_EGTA_37_down <- bind_rows(
  MF_slim_EGTA_37_down,
  BP_slim_EGTA_37_down,
  CC_slim_EGTA_37_down) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)

GO_slim_EGTA_37_down$name <- factor(GO_slim_EGTA_37_down$name, levels = rev(GO_slim_EGTA_37_down$name))


# Open GO analysis upregulated gene in WT vs flc1∆ at 30°C
MF_slim_30_up <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_30C_YPD/GO_WTvsdelta_30C_YPD_up2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_30C_YPD",
    GO_type = "MF",
    DEG = "up"
  )

BP_slim_30_up <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_30C_YPD/GO_WTvsdelta_30C_YPD_up2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_30C_YPD",
    GO_type = "BP", ,
    DEG = "up"
  )

CC_slim_30_up <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_30C_YPD/GO_WTvsdelta_30C_YPD_up2x_FDR5_cellular_component_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_30C_YPD",
    GO_type = "CC", ,
    DEG = "up"
  )

GO_slim_30_up <- bind_rows(
  MF_slim_30_up,
  BP_slim_30_up,
  CC_slim_30_up) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)

GO_slim_30_up$name <- factor(GO_slim_30_up$name, levels = rev(GO_slim_30_up$name))
  

# Open GO analysis for upregulated gene in WT vs flc1∆ at 37°C
MF_slim_37_up <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_YPD/GO_WTvsdelta_37C_YPD_up2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_YPD",
    GO_type = "MF",
    DEG = "up"
  )

BP_slim_37_up <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_YPD/GO_WTvsdelta_37C_YPD_up2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_YPD",
    GO_type = "BP",
    DEG = "up"
  )

CC_slim_37_up <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_YPD/GO_WTvsdelta_37C_YPD_up2x_FDR5_cellular_component_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_YPD",
    GO_type = "CC",
    DEG = "up"
  )

GO_slim_37_up <- bind_rows(
  MF_slim_37_up,
  BP_slim_37_up,
  CC_slim_37_up) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)

GO_slim_37_up$name <- factor(GO_slim_37_up$name, levels = rev(GO_slim_37_up$name))

# Open GO analysis for upregulated gene in WT vs flc1∆ in EGTA at 37°C
MF_slim_EGTA_37_up <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_EGTA/GO_WTvsdelta_37C_EGTA_up2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_EGTA",
    GO_type = "MF",
    DEG = "up"
  )

BP_slim_EGTA_37_up <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_EGTA/GO_WTvsdelta_37C_EGTA_up2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_EGTA",
    GO_type = "BP",
    DEG = "up"
  )

CC_slim_EGTA_37_up <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_EGTA/GO_WTvsdelta_37C_EGTA_up2x_FDR5_cellular_component_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_EGTA",
    GO_type = "CC",
    DEG = "up"
  )

GO_slim_EGTA_37_up <- bind_rows(
  MF_slim_EGTA_37_up,
  BP_slim_EGTA_37_up,
  CC_slim_EGTA_37_up) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)

GO_slim_EGTA_37_up$name <- factor(GO_slim_EGTA_37_up$name, levels = rev(GO_slim_EGTA_37_up$name))


#Combine into a single dataframe!
GOslim <- bind_rows(GO_slim_EGTA_37_up, GO_slim_37_up, GO_slim_30_up, GO_slim_EGTA_37_down, GO_slim_37_down, GO_slim_30_down) %>% 
  mutate(comparison = factor(comparison, 
                             levels = c("WTvsdelta_30C_YPD", "WTvsdelta_37C_YPD", "WTvsdelta_37C_EGTA"),
                             labels = c("FLC1 vs flc1∆ in YPD, 30°C", "FLC1 vs flc1∆ in YPD, 37°C", "FLC1 vs flc1∆ in YPD+EGTA, 37°C")),
         deg = factor(deg, levels = c("down", "up"), labels = c("up in flc1∆", "up in FLC1"))) %>% 
  ungroup() %>%
  group_by(go_type, comparison) %>% 
  arrange(desc(fold_enrichment), .by_group = T) %>% 
  filter(!name %in% c("cellular component", "molecular function", "biological process", "organelle", "intracellular anatomical structure")) 

GOslim$name <- factor(GOslim$name, levels = rev(unique(GOslim$name)))
```
Plot GO term results with fold-enrichment on x-axis, GOterms on y-axis. Size of point = number of genes found for each term, linetype indicates GO term category. Facet by temperature. 
```{r Plot GO terms for 30°C and 37°C, fig.width = 10, fig.height=7}
GOslim_WTvsdelat_YPD_30 <- ggplot(
  data = GOslim %>% filter(comparison == "FLC1 vs flc1∆ in YPD, 30°C")) + 
  geom_segment(aes(x = 0, xend = fold_enrichment, y = name, linetype = go_type)) +
  scale_linetype_manual(values = c("solid", "twodash", "dotted")) + 
  ggnewscale::new_scale_color() +  
  geom_point(aes(x = fold_enrichment, y = name, size = result_count, colour = deg)) +
  scale_size_continuous(limits = c(1,85), breaks = c(20, 40, 60, 80)) + 
  scale_colour_manual(values = c("firebrick", "blue3")) + 
    labs(size="Number of Genes", linetype="GO term", colour = "Gene Expression") + 
    xlab("Fold enrichment") + 
    ylab("") + 
    scale_x_continuous(breaks = c(0,2,4,6,8,10,12,14), limits = c(0,14), expand = c(0.01,0.01)) +
    facet_grid2(deg~., scales = "free_y", axes = "all", space = "free") +
    #ggtitle("FLC1 vs flc1∆ in YPD, 30°C") +
    theme(strip.text = element_text(colour = "white"),
          axis.text.y = element_text(size = 14, colour = "black"),
          axis.text.x = element_text(size = 14, colour = "black"),
          axis.title.x = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.justification = "top",
          legend.spacing.y = unit(0.3, 'cm'),
          legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm')
          ) + 
  guides(linetype = guide_legend(order=1),
         colour = guide_legend(order=2),
         size = guide_legend(order=3)) 

GOslim_WTvsdelat_YPD_30

GOslim_WTvsdelat_YPD_37 <- ggplot(
  data = GOslim %>% filter(comparison == "FLC1 vs flc1∆ in YPD, 37°C")) + 
  geom_segment(aes(x = 0, xend = fold_enrichment, y = name, linetype = go_type)) +
  scale_linetype_manual(values = c("solid", "twodash", "dotted")) + 
  ggnewscale::new_scale_color() +  
  geom_point(aes(x = fold_enrichment, y = name, size = result_count, colour = deg)) +
  scale_size_continuous(limits = c(1,85), breaks = c(20, 40, 60, 80)) + 
  scale_colour_manual(values = c("firebrick", "blue3")) + 
    labs(size="Number of Genes", linetype="GO term", colour = "Gene Expression") + 
    xlab("Fold enrichment") + 
    ylab("") + 
    scale_x_continuous(breaks = c(0,2,4,6,8,10,12,14), limits = c(0,14), expand = c(0.01,0.01)) +
    facet_grid2(deg~., scales = "free_y", axes = "all", space = "free") +
    #ggtitle("FLC1 vs flc1∆ in YPD, 37°C") +
    theme(strip.text = element_text(colour = "white"),
          axis.text.y = element_text(size = 14, colour = "black"),
          axis.text.x = element_text(size = 14, colour = "black"),
          axis.title.x = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.justification = "top",
          legend.spacing.y = unit(0.3, 'cm'),
          legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm')
          ) + 
  guides(linetype = guide_legend(order=1),
         colour = guide_legend(order=2),
         size = guide_legend(order=3))

GOslim_WTvsdelat_YPD_37

GOslim_WTvsdelat_EGTA_37 <- ggplot(
  data = GOslim %>% filter(comparison == "FLC1 vs flc1∆ in YPD+EGTA, 37°C")) + 
  geom_segment(aes(x = 0, xend = fold_enrichment, y = name, linetype = go_type)) +
  scale_linetype_manual(values = c("solid", "twodash", "dotted")) + 
  ggnewscale::new_scale_color() +  
  geom_point(aes(x = fold_enrichment, y = name, size = result_count, colour = deg)) +
  scale_size_continuous(limits = c(1,85), breaks = c(20, 40, 60, 80)) + 
  scale_colour_manual(values = c("firebrick", "blue3")) + 
    labs(size="Number of Genes", linetype="GO term", colour = "Gene Expression") + 
    xlab("Fold enrichment") + 
    ylab("") + 
    scale_x_continuous(breaks = c(0,2,4,6,8,10,12,14), limits = c(0,14), expand = c(0.01,0.01)) +
    facet_grid2(deg~., scales = "free_y", axes = "all", space = "free") +
    #ggtitle("FLC1 vs flc1∆ in YPD, 37°C") +
    theme(strip.text = element_text(colour = "white"),
          axis.text.y = element_text(size = 14, colour = "black"),
          axis.text.x = element_text(size = 14, colour = "black"),
          axis.title.x = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.justification = "top",
          legend.spacing.y = unit(0.3, 'cm'),
          legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm')
          ) + 
  guides(linetype = guide_legend(order=1),
         colour = guide_legend(order=2),
         size = guide_legend(order=3))

GOslim_WTvsdelat_EGTA_37

```


```{r}
GOslim_up_delta <- ggplot(
  data= GOslim %>% 
    filter(deg == 'up in flc1∆') %>%
    mutate(comparison = factor(comparison, 
                              levels = c("FLC1 vs flc1∆ in YPD, 30°C", "FLC1 vs flc1∆ in YPD, 37°C", "FLC1 vs flc1∆ in YPD+EGTA, 37°C"),
                              labels = c("YPD\n30°C", "YPD\n37°C", "YPD +\nEGTA\n37°C")),
            go_type = factor(go_type, 
                            levels = c("Biological process", "Molecular function", "Cellular component"),
                            labels = c("Biological\nprocess", "Molecular\nfunction", "Cellular\ncomponent"))
           ),
  aes(y = name, x = comparison, size = result_count, colour = fold_enrichment)
) + 
  geom_point() +
  scale_size_continuous(limits = c(1,85), breaks = c(20, 40, 60, 80)) + 
  scale_colour_gradient2(low = "grey95", high = "firebrick", limits =  c(1,12), breaks = c(2, 4, 6, 8, 10, 12)) + 
  geom_point(shape = 1, colour = "black") + 
  labs(size="Number of\nGenes", colour = "Fold\nEnrichment") + 
  guides(colour = guide_legend(order=2),
         size = guide_legend(order=1, override.aes=list(shape = 1))) + 
  facet_grid2(go_type~., scales = 'free_y', space = "free")  + 
  ggtitle('Up in *flc1*∆') +
  theme(plot.title = element_markdown(face = "plain", size = 18),
        strip.text.y = element_text(size = 14, colour = "black", angle = 0, hjust = 0),
        panel.background = element_rect(fill = "grey98"),
        strip.background = element_rect(fill = "grey90"),
        axis.text.y = element_text(size = 12, colour = "black"),
        axis.text.x = element_text(size = 14, colour = "black"),
        axis.title = element_blank(),
        legend.title = element_text(size = 16, colour = "black"),
        legend.text = element_text(size = 14, colour = "black"),
        legend.justification = "top",
        legend.spacing.y = unit(0.3, 'cm'),
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm')
    
  )
  
GOslim_up_delta


GOslim_up_WT <- ggplot(
  data= GOslim %>% 
    filter(deg == 'up in FLC1') %>%
    mutate(comparison = factor(comparison, 
                              levels = c("FLC1 vs flc1∆ in YPD, 30°C", "FLC1 vs flc1∆ in YPD, 37°C", "FLC1 vs flc1∆ in YPD+EGTA, 37°C"),
                              labels = c("YPD\n30°C", "YPD\n37°C", "YPD +\nEGTA\n37°C")),
           go_type = factor(go_type, 
                            levels = c("Biological process", "Molecular function", "Cellular component"),
                            labels = c("Biological\nprocess", "Molecular\nfunction", "Cellular\ncomponent"))
           ),
  aes(y = name, x = comparison, size = result_count, colour = fold_enrichment)
) + 
  geom_point() +
  scale_size_continuous(limits = c(1,85), breaks = c(20, 40, 60, 80)) + 
  scale_colour_gradient2(low = "grey95", high = "blue3", limits =  c(1,14), breaks = c(2, 4, 6, 8, 10, 12)) + 
  geom_point(shape = 1, colour = "black") + 
  labs(size="Number of\nGenes", colour = "Fold\nEnrichment") + 
  guides(colour = guide_legend(order=2),
         size = guide_legend(order=1, override.aes=list(shape = 1))) + 
  facet_grid2(go_type~., scales = 'free_y', space = "free") +
  ggtitle('Up in *FLC1*') +
  theme(plot.title = element_markdown(face = "plain", size = 18),
        strip.text.y = element_text(size = 14, colour = "black", angle = 0, hjust = 0),
        panel.background = element_rect(fill = "grey98"),
        strip.background = element_rect(fill = "grey90"),
        axis.text.y = element_text(size = 12, colour = "black"),
        axis.text.x = element_text(size = 14, colour = "black"),
        axis.title = element_blank(),
        legend.title = element_text(size = 16, colour = "black"),
        legend.text = element_text(size = 14, colour = "black"),
        legend.justification = "top",
        legend.spacing.y = unit(0.3, 'cm'),
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm')
    
  )  
GOslim_up_WT


GOslim_grid <- plot_grid(
  GOslim_up_delta,
  GOslim_up_WT,
  ncol = 1,
  axis = "lr",
  align = "v",
  rel_heights = c(12, 25),
  scale = 0.95
)
```


## Muti-panel figure DGE in FLC1vsflc1∆ at 30°C and 37°C

```{r Mulitpannel FLC1vsflc1∆ fig.width=30, fig.height=40}
Volcano <- cowplot::plot_grid(
  volcano_WTvsdelta_30C_YPD,
  volcano_WTvsdelta_37C_YPD,
  volcano_WTvsdelta_37C_EGTA, 
  eulerr_plots,
  axis = "l",
  align = "v",
  ncol =1, 
  rel_heights = c(1,1,1,1),
  rel_widths = c(1, 1, 1, 2),
  scale = 0.9
)

Volcano_grid <- cowplot::plot_grid(
  volcano_WTvsdelta_30C_YPD,
  volcano_WTvsdelta_37C_YPD,
  volcano_WTvsdelta_37C_EGTA, 
  axis = "l",
  align = "v",
  ncol =1, 
  rel_heights = c(1,1,1),
  rel_widths = c(1, 1, 1),
  scale = 0.95
)


DEG_GO <- cowplot::plot_grid(
  GOslim_WTvsdelat_YPD_30,
  GOslim_WTvsdelat_YPD_37 + theme(legend.position = "none"),
  GOslim_WTvsdelat_EGTA_37 + theme(legend.position = "none"),
  axis = "lr",
  align = "v", 
  ncol =1, 
  rel_heights = c(23,15, 28),
  scale = 0.9)

plot_grid(
  Volcano_grid,
  GOslim_grid,
  axis = "tb",
  align = "h", 
  rel_widths = c(1,2)
)

#ggsave(here("results", "figures", "WTvsDelta_YPD30C_YPD37C_EGTA37C_pannel_noVenn.png"), width = 16, height = 13, dpi = 1000)
```


```{r fig.width=16, fig.height=19}
Volcano_grid <- cowplot::plot_grid(
  volcano_WTvsdelta_30C_YPD,
  volcano_WTvsdelta_37C_YPD,
  volcano_WTvsdelta_37C_EGTA, 
  axis = "l",
  align = "v",
  ncol =3, 
  rel_heights = c(1,1,1),
  rel_widths = c(1, 1, 1),
  scale = 0.95
)

Venn_grid <- cowplot::plot_grid(
  DEGs_and_EGs_grid, 
  ggvenn_plot_down,
  ggvenn_plot_up,
  ncol = 1,
  scale = 0.9)


GOterm_grid <- cowplot::plot_grid(
  Venn_grid, 
  GOslim_grid, 
  axis = "t",
  align = "h",
  ncol = 2,
  rel_widths = c(1, 2)
)



plot_grid(
  Volcano_grid,
  GOterm_grid,
  axis = "lr",
  align = "h", 
  ncol =1,
  rel_heights = c(1,3),
  scale = 0.95
)

ggsave(here("results", "figures", "WTvsDelta_YPD30C_YPD37C_EGTA37C_multipannel.png"), width = 16, height = 20, dpi = 1000)
```

