---
title: "RM_thesis_growth_conditions"
author: "Rachael Murray"
date: "2025-04-24"
output: html_document
---

In this document I explore how temperature and EGTA affect the transciptome of Cryptocococcus neoformans KN99alpha (FLC1) cells and flc1∆ cells. I previously performed the following pairwise differential gene expression analyses: 

- YPD, 37°C vs YPD, 30°C - *FLC1* 
- YPD, 37°C vs YPD, 30°C - *flc1*∆ 
- YPD + EGTA, 37°C vs YPD, 37°C - *FLC1*
- YPD + EGTA, 37°C vs YPD, 37°C - *flc1*∆

This document includes code the reproduce multi-pannel figures for my PhD thesis chapter.  

Input files: 
- results/DEGs/deseq_df_37Cvs30C_WT_YPD.txt
- results/DEGs/deseq_df_37Cvs30C_delta_YPD.txt
- results/DEGs/deseq_df_EGTAvsYPD_WT_37C.txt
- results/DEGs/deseq_df_EGTAvsYPD_delta_37C.txt
- results/GOanalysis/GO_37Cvs30C_WT/
- results/GOanalysis/GO_37Cvs30C_delta/
- results/GOanalysis/GO_EGTAvsYPD_WT/
- results/GOanalysis/GO_EGTAvsYPD_delta/
- results/gene_info/genes_to_highlight.csv

Key takeaways: 

- 

 
```{r call_load_count_data, child = "load_count_data.Rmd"}
```
Open results from differential gene expression analysis. 
```{r Open DEseq results from FLC1 vs flc1∆ at 30°C and 37°C}
# 37C vs 30C in FLC1
deseq_df_37Cvs30C_WT <- read_tsv(
 file = here::here("results", "DEGs",
                              "deseq_df_37Cvs30C_WT_YPD.txt"))
# Number up in 30°C in FLC1
DEGs_down_37Cvs30C_WT <- deseq_df_37Cvs30C_WT %>% 
                     dplyr::filter(log2FC < -1, padj < 0.05) %>% 
                     arrange(log2FC)
# Number up in 37°C in FLC1
DEGs_up_37Cvs30C_WT <- deseq_df_37Cvs30C_WT %>% 
                     dplyr::filter(log2FC > 1, padj < 0.05) %>% 
                     arrange(log2FC)

# 37C vs 30C in flc1∆
deseq_df_37Cvs30C_delta <- read_tsv(
 file = here::here("results", "DEGs",
                              "deseq_df_37Cvs30C_delta_YPD.txt"))
# Number up in 30°C in flc1∆
DEGs_down_37Cvs30C_delta <- deseq_df_37Cvs30C_delta %>% 
                     dplyr::filter(log2FC < -1, padj < 0.05) %>% 
                     arrange(log2FC)
# Number up in 37°C in flc1∆
DEGs_up_37Cvs30C_delta <- deseq_df_37Cvs30C_delta %>% 
                     dplyr::filter(log2FC > 1, padj < 0.05) %>% 
                     arrange(log2FC)
  
# EGTA vs YPD at 37C in FLC1
deseq_df_EGTAvsYPD_37C_WT <- read_tsv(
 file = here::here("results", "DEGs",
                              "deseq_df_EGTAvsYPD_WT_37C.txt"))
# Number up in YPD in FLC1
DEGs_down_EGTAvsYPD_37C_WT <- deseq_df_EGTAvsYPD_37C_WT  %>% 
                     dplyr::filter(log2FC < -1, padj < 0.05) %>% 
                     arrange(log2FC)
# Number up in EGTA in FLC1
DEGs_up_EGTAvsYPD_37C_WT<- deseq_df_EGTAvsYPD_37C_WT  %>% 
                     dplyr::filter(log2FC > 1, padj < 0.05) %>% 
                     arrange(desc(log2FC))

# EGTA vs YPD at 37C in flc1∆
deseq_df_EGTAvsYPD_37C_delta <- read_tsv(
 file = here::here("results", "DEGs",
                              "deseq_df_EGTAvsYPD_delta_37C.txt"))
# Number up in YPD in flc1∆
DEGs_down_EGTAvsYPD_37C_delta <- deseq_df_EGTAvsYPD_37C_delta  %>% 
                     dplyr::filter(log2FC < -1, padj < 0.05) %>% 
                     arrange(log2FC)
# Number up in EGTA in flc1∆
DEGs_up_EGTAvsYPD_37C_delta <- deseq_df_EGTAvsYPD_37C_delta  %>% 
                     dplyr::filter(log2FC > 1, padj < 0.05) %>% 
                     arrange(desc(log2FC))

```

```{r}
A <- DEGs_up_37Cvs30C_WT %>% arrange(desc(log2FC)) %>% 
  slice_head(n=40)

B <- DEGs_up_37Cvs30C_delta %>% arrange(desc(log2FC)) %>% 
  slice_head(n=40) %>% 
  filter(gene %in% A$gene)

DEGs_up_37Cvs30C_delta %>% arrange(desc(log2FC)) %>% 
  write_tsv(here("results/DEGs_fungiDB_info/37vs30_delta_up.tsv"))


```




```{r fig.width = 4, fig.height=4}




log2FC_WT_sig <- deseq_df_37Cvs30C_WT %>% 
  filter(padj < 0.1) %>% 
  select(gene, log2FC) %>% 
  dplyr::rename(log2FC_WT = log2FC) 

log2FC_delta_sig <- deseq_df_37Cvs30C_delta %>% 
  filter(padj < 0.1) %>% 
  select(gene, log2FC) %>% 
  dplyr::rename(log2FC_delta = log2FC) 


X <- inner_join(log2FC_WT_sig, log2FC_delta_sig) 

Y <- deseq_df_37Cvs30C_WT %>% 
  select(gene, log2FC) %>% 
  dplyr::rename(log2FC_WT = log2FC) %>% 
  left_join(deseq_df_37Cvs30C_delta) %>% 
  select(gene, log2FC, log2FC_WT) %>% 
  dplyr::rename(log2FC_delta = log2FC) 


ggplot(data = Y, aes(x = log2FC_delta, y = log2FC_WT)) + 
  geom_hline(yintercept = 1, size = 0.5, colour = "grey60") +
  geom_hline(yintercept = -1, size = 0.5, colour = "grey60") +
  geom_vline(xintercept = 1, size = 0.5, colour = "grey60") +
  geom_vline(xintercept = -1, size = 0.5, colour = "grey60") +
  geom_point(colour = "grey95", size = 0.5) + 
  geom_point(data = Y %>% filter(gene %in% log2FC_WT_sig$gene),
             aes(x = log2FC_delta, y = log2FC_WT), 
             colour = 'red', 
             alpha = 0.2,
             size = 0.5) +
  geom_point(data = Y %>% filter(gene %in% log2FC_delta_sig$gene),
             aes(x = log2FC_delta, y = log2FC_WT), 
             colour = 'blue3', 
             alpha = 0.2,
             size = 0.5) +
  xlim(-5, 7) + 
  ylim(-5, 7) 


X %>% 
  filter(log2FC_WT > 0, log2FC_delta < 0)


X %>% 
  filter(log2FC_WT < 0, log2FC_delta > 0)


log2FC_delta_EGTA <- deseq_df_EGTAvsYPD_37C_delta %>% 
  select(gene, log2FC) %>% 
  dplyr::rename(log2FC_delta = log2FC) 

log2FC_WT_EGTA <- deseq_df_EGTAvsYPD_37C_WT %>% 
  select(gene, log2FC) %>% 
  dplyr::rename(log2FC_WT = log2FC) 

EGTA <- left_join(log2FC_delta_EGTA, log2FC_WT_EGTA)

EGTA_delta_sig <- deseq_df_EGTAvsYPD_37C_delta %>% 
  filter(padj < 0.1) %>% 
  select(gene, log2FC) %>% 
  dplyr::rename(log2FC_delta = log2FC)

EGTA_WT_sig <- deseq_df_EGTAvsYPD_37C_delta %>% 
  filter(padj < 0.1) %>% 
  select(gene, log2FC) %>% 
  dplyr::rename(log2FC_WT = log2FC)
  


X <- inner_join(log2FC_WT_sig, log2FC_delta_sig) 

Y <- deseq_df_37Cvs30C_WT %>% 
  select(gene, log2FC) %>% 
  dplyr::rename(log2FC_WT = log2FC) %>% 
  left_join(deseq_df_37Cvs30C_delta) %>% 
  select(gene, log2FC, log2FC_WT) %>% 
  dplyr::rename(log2FC_delta = log2FC) 

ggplot(data = Y, aes(x = log2FC_delta, y = log2FC_WT)) + 
  geom_point() + 
  ylim(-5, 5) + 
  xlim(-5, 5)




ggplot(data = EGTA, aes(x = log2FC_delta, y = log2FC_WT)) +
  geom_hline(yintercept = 1, size = 0.5, colour = "grey60") +
  geom_hline(yintercept = -1, size = 0.5, colour = "grey60") +
  geom_vline(xintercept = 1, size = 0.5, colour = "grey60") +
  geom_vline(xintercept = -1, size = 0.5, colour = "grey60") +
  geom_point(colour = "grey40", size = 0.5) + 
  xlim(-5, 9) +
  ylim(-5,9)
  geom_point(data = EGTA %>% 
               filter(gene %in% EGTA_WT_sig$gene, gene %in% EGTA_delta_sig$gene),
             aes(x = log2FC_delta, y = log2FC_WT), 
             colour = "red", size = 0.6) + 
  geom_text_repel(aes(label = gene), size = 1)

EGTA %>% 
  arrange(desc(log2FC_delta))



EGTA <- deseq_df_EGTAvsYPD_37C_WT %>% 
  select(gene, log2FC) %>% 
  dplyr::rename(log2FC_WT_EGTA = log2FC) %>% 
  left_join(deseq_df_EGTAvsYPD_37C_delta) %>% 
  select(gene, log2FC, log2FC_WT_EGTA) %>% 
  dplyr::rename(log2FC_delta_EGTA = log2FC)  

EGTA_labs <- EGTA %>% 
  filter(log2FC_delta_EGTA > 2.5 | log2FC_WT_EGTA > 2.5 |
         log2FC_delta_EGTA < -2 | log2FC_WT_EGTA < -2)

ggplot(data = EGTA, aes(x = log2FC_delta_EGTA, y = log2FC_WT_EGTA)) + 
  geom_point(size = 0.5) + 
  geom_text_repel(data = EGTA_labs, aes(x = log2FC_delta_EGTA, y = log2FC_WT_EGTA, label = gene), size = 1)

library(ggrepel)

```







Make volcano plots for each growth temperature. 
```{r Volcano plots for each comparison, fig.width = 4, fig.height=3}
geom_point_size <- 0.3
limits_x <- c(-6, 6)
colour_positive <- "blue3"
colour_negative <- "firebrick"
colour_neutral <- "grey70"

# Volcano 37C vs 30C in WT
volcano_37Cvs30C_WT <- 
 ggplot(data = deseq_df_37Cvs30C_WT,
         aes(x = log2FC, y = -log10(padj)) ) +
  geom_point(size = geom_point_size - 0.1, colour = colour_neutral) +
  geom_point(data = deseq_df_37Cvs30C_WT %>% 
               dplyr::filter(
                 log2FC  > 1, padj < 0.05),
             size = geom_point_size, colour = colour_positive) +
  geom_point(data = deseq_df_37Cvs30C_WT %>% 
               dplyr::filter(
                 log2FC < -1, padj < 0.05),
             size = geom_point_size, colour = colour_negative) +
  scale_y_continuous("-log10(p)",
                     limits = c(0,50), expand = c(0,0),
                     oob=scales::squish) +
  scale_x_continuous("log2 fold-change\n← down in YPD, 37°C       up in YPD, 37°C →",
                     breaks = -6:6,
                     limits = limits_x, expand = c(0.01,0.01),
                     oob = scales::squish) +
  coord_cartesian(clip = "off") + 
  ggtitle("YPD, 37°C vs YPD, 30°C<br>*FLC1*") +
  theme(plot.title = element_markdown(colour = "black", size = 18, face = "plain"),
        axis.text.x = element_text(size = 14, colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black"),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.title.y = element_text(size = 14, colour = "black")) 

volcano_37Cvs30C_WT

# Volcano 37C vs 30C in delta
volcano_37Cvs30C_delta <- 
 ggplot(data = deseq_df_37Cvs30C_delta,
         aes(x = log2FC, y = -log10(padj)) ) +
  geom_point(size = geom_point_size - 0.1, colour = colour_neutral) +
  geom_point(data = deseq_df_37Cvs30C_delta %>% 
               dplyr::filter(
                 log2FC  > 1, padj < 0.05),
             size = geom_point_size, colour = colour_positive) +
  geom_point(data = deseq_df_37Cvs30C_delta %>% 
               dplyr::filter(
                 log2FC < -1, padj < 0.05),
             size = geom_point_size, colour = colour_negative) +
  scale_y_continuous("-log10(p)",
                     limits = c(0,50), expand = c(0,0),
                     oob=scales::squish) +
  scale_x_continuous("log2 fold-change\n← down in YPD, 37°C       up in YPD, 37°C →",
                     breaks = -6:6,
                     limits = limits_x, expand = c(0.01,0.01),
                     oob = scales::squish) +
  coord_cartesian(clip = "off") + 
  ggtitle("YPD, 37°C vs YPD, 30°C<br>*flc1*∆") +
  theme(plot.title = element_markdown(colour = "black", size = 18, face = "plain"),
        axis.text.x = element_text(size = 14, colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black"),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.title.y = element_text(size = 14, colour = "black")) 

volcano_37Cvs30C_delta


# Volcano EGTA vs YPD at 37C in WT 
volcano_EGTAvsYPD_37C_WT <- 
 ggplot(data = deseq_df_EGTAvsYPD_37C_WT,
         aes(x = log2FC, y = -log10(padj)) ) +
  geom_point(size = geom_point_size - 0.1, colour = colour_neutral) +
  geom_point(data = deseq_df_EGTAvsYPD_37C_WT %>% 
               dplyr::filter(
                 log2FC  > 1, padj < 0.05),
             size = geom_point_size, colour = colour_positive) +
  geom_point(data = deseq_df_EGTAvsYPD_37C_WT %>% 
               dplyr::filter(
                 log2FC < -1, padj < 0.05),
             size = geom_point_size, colour = colour_negative) +
  scale_y_continuous("-log10(p)",
                     limits = c(0,50), expand = c(0,0),
                     oob=scales::squish) +
  scale_x_continuous("log2 fold-change\n← down in YPD + EGTA       up in YPD + EGTA →",
                     breaks = -6:6,
                     limits = limits_x, expand = c(0.01,0.01),
                     oob = scales::squish) +
  coord_cartesian(clip = "off") + 
  ggtitle("YPD + EGTA, 37°C vs YPD, 37°C<br>*FLC1*") +
  theme(plot.title = element_markdown(colour = "black", size = 18, face = "plain"),
        axis.text.x = element_text(size = 14, colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black"),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.title.y = element_text(size = 14, colour = "black")) 

volcano_EGTAvsYPD_37C_WT

# Volcano EGTA vs YPD at 37C in delta
volcano_EGTAvsYPD_37C_delta <- 
 ggplot(data = deseq_df_EGTAvsYPD_37C_delta,
         aes(x = log2FC, y = -log10(padj)) ) +
  geom_point(size = geom_point_size - 0.1, colour = colour_neutral) +
  geom_point(data = deseq_df_EGTAvsYPD_37C_delta %>% 
               dplyr::filter(
                 log2FC  > 1, padj < 0.05),
             size = geom_point_size, colour = colour_positive) +
  geom_point(data = deseq_df_EGTAvsYPD_37C_delta %>% 
               dplyr::filter(
                 log2FC < -1, padj < 0.05),
             size = geom_point_size, colour = colour_negative) +
  scale_y_continuous("-log10(p)",
                     limits = c(0,50), expand = c(0,0),
                     oob=scales::squish) +
  scale_x_continuous("log2 fold-change\n← down in YPD + EGTA       up in YPD + EGTA →",
                     breaks = -6:6,
                     limits = limits_x, expand = c(0.01,0.01),
                     oob = scales::squish) +
  coord_cartesian(clip = "off") + 
  ggtitle("YPD + EGTA, 37°C vs YPD, 37°C<br>*flc1*∆") +
  theme(plot.title = element_markdown(colour = "black", size = 18, face = "plain"),
        axis.text.x = element_text(size = 14, colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black"),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.title.y = element_text(size = 14, colour = "black")) 

volcano_EGTAvsYPD_37C_delta
```


```{r}
# Open GO analysis of down in YPD 37C vs YPD 30C in WT 
MF_37Cvs30C_WT_down <- read_tsv(
  here("results/GOanalysis/GO_37Cvs30C_WT/GO_37Cvs30C_WT_YPD_down2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "37Cvs30C_WT",
    GO_type = "MF",
    DEG = "down"
  )

BP_37Cvs30C_WT_down <- read_tsv(
  here("results/GOanalysis/GO_37Cvs30C_WT/GO_37Cvs30C_WT_YPD_down2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "37Cvs30C_WT",
    GO_type = "BP",
    DEG = "down"
  )

CC_37Cvs30C_WT_down <- read_tsv(
  here("results/GOanalysis/GO_37Cvs30C_WT/GO_37Cvs30C_WT_YPD_down2x_FDR5_cellular_component_slim.tsv")
) %>% 
  mutate(
    Comparison = "37Cvs30C_WT",
    GO_type = "CC",
    DEG = "down"
  )

GO_slim_37Cvs30C_WT_down <- bind_rows(
  MF_37Cvs30C_WT_down,
  BP_37Cvs30C_WT_down,
  CC_37Cvs30C_WT_down) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)

# Open GO analysis of up in YPD 37C vs YPD 30C in WT 
MF_37Cvs30C_WT_up <- read_tsv(
  here("results/GOanalysis/GO_37Cvs30C_WT/GO_37Cvs30C_WT_YPD_up2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "37Cvs30C_WT",
    GO_type = "MF",
    DEG = "up"
  )

BP_37Cvs30C_WT_up <- read_tsv(
  here("results/GOanalysis/GO_37Cvs30C_WT/GO_37Cvs30C_WT_YPD_up2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "37Cvs30C_WT",
    GO_type = "BP",
    DEG = "up"
  )

#CC_37Cvs30C_WT_up <- read_tsv(
 # here("results/GOanalysis/GO_37Cvs30C_WT/GO_37Cvs30C_WT_YPD_up2x_FDR5_cellular_component_slim.tsv")
#) %>% 
 # mutate(
  #  Comparison = "37Cvs30C_WT",
   # GO_type = "CC",
   # DEG = "up"
  #)

GO_slim_37Cvs30C_WT_up <- bind_rows(
  MF_37Cvs30C_WT_up,
  BP_37Cvs30C_WT_up
  ) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)

# Open GO analysis of down in YPD 37C vs YPD 30C in delta 
MF_37Cvs30C_delta_down <- read_tsv(
  here("results/GOanalysis/GO_37Cvs30C_delta/GO_37Cvs30C_delta_YPD_down2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "37Cvs30C_delta",
    GO_type = "MF",
    DEG = "down"
  )

BP_37Cvs30C_delta_down <- read_tsv(
  here("results/GOanalysis/GO_37Cvs30C_delta/GO_37Cvs30C_delta_YPD_down2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "37Cvs30C_delta",
    GO_type = "BP",
    DEG = "down"
  )

CC_37Cvs30C_delta_down <- read_tsv(
  here("results/GOanalysis/GO_37Cvs30C_delta/GO_37Cvs30C_delta_YPD_down2x_FDR5_cellular_component_slim.tsv")
) %>% 
  mutate(
    Comparison = "37Cvs30C_delta",
    GO_type = "CC",
    DEG = "down"
  )

GO_slim_37Cvs30C_delta_down <- bind_rows(
  MF_37Cvs30C_delta_down,
  BP_37Cvs30C_delta_down,
  CC_37Cvs30C_delta_down) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)

# Open GO analysis of up in YPD 37C vs YPD 30C in delta 
MF_37Cvs30C_delta_up <- read_tsv(
  here("results/GOanalysis/GO_37Cvs30C_delta/GO_37Cvs30C_delta_YPD_up2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "37Cvs30C_delta",
    GO_type = "MF",
    DEG = "up"
  )

BP_37Cvs30C_delta_up <- read_tsv(
  here("results/GOanalysis/GO_37Cvs30C_delta/GO_37Cvs30C_delta_YPD_up2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "37Cvs30C_delta",
    GO_type = "BP",
    DEG = "up"
  )

#CC_37Cvs30C_delta_up <- read_tsv(
 # here("results/GOanalysis/GO_37Cvs30C_delta/GO_37Cvs30C_delta_YPD_up2x_FDR5_cellular_component_slim.tsv")
#) %>% 
 # mutate(
  #  Comparison = "37Cvs30C_delta",
   # GO_type = "CC",
    #DEG = "up"
  #)

GO_slim_37Cvs30C_delta_up <- bind_rows(
  MF_37Cvs30C_delta_up,
  BP_37Cvs30C_delta_up) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)


#Combine into a single dataframe!
GOslim_37Cvs30C <- bind_rows(GO_slim_37Cvs30C_delta_up, GO_slim_37Cvs30C_delta_down, GO_slim_37Cvs30C_WT_up, GO_slim_37Cvs30C_WT_down) %>% 
  mutate(comparison = factor(comparison, 
                             levels = c("37Cvs30C_WT", "37Cvs30C_delta"),
                             labels = c("YPD, 37°C vs YPD, 30°C in FLC1", "YPD, 37°C vs YPD, 30°C in flc1∆")),
         deg = factor(deg, levels = c("down", "up"), labels = c("up in YPD, 30°C", "up in YPD, 37°C"))) %>% 
  ungroup() %>%
  group_by(go_type, comparison) %>% 
  arrange(desc(fold_enrichment), .by_group = T) %>% 
  filter(!name %in% c("cellular component", "molecular function", "biological process", "organelle", "intracellular anatomical structure")) 

GOslim_37Cvs30C$name <- factor(GOslim_37Cvs30C$name, levels = rev(unique(GOslim_37Cvs30C$name)))

```

```{r}
GOslim_up_37 <- ggplot(
  data= GOslim_37Cvs30C %>% 
    filter(deg == 'up in YPD, 37°C') %>%
    mutate(comparison = factor(comparison, 
                              levels = c("YPD, 37°C vs YPD, 30°C in FLC1", "YPD, 37°C vs YPD, 30°C in flc1∆"),
                              labels = c("*FLC1*", "*flc1*∆")),
            go_type = factor(go_type, 
                            levels = c("Biological process", "Molecular function", "Cellular component"),
                            labels = c("Biological\nprocess", "Molecular  \nfunction", "Cellular\ncomponent"))
           ),
  aes(y = name, x = comparison, size = result_count, colour = fold_enrichment)
) + 
  geom_point() +
  scale_size_continuous(limits = c(1,20), breaks = c(5, 10, 15, 20)) + 
  scale_colour_gradient2(low = "grey95", high = "blue3", limits =  c(1,30), breaks = c(5, 10, 15, 20, 25, 30)) + 
  geom_point(shape = 1, colour = "black") + 
  labs(size="Number of\nGenes", colour = "Fold\nEnrichment") + 
  guides(colour = guide_legend(order=2),
         size = guide_legend(order=1, override.aes=list(shape = 1))) + 
  facet_grid2(go_type~., scales = 'free_y', space = "free") +
  theme(plot.title = element_markdown(face = "plain", size = 18),
        strip.text.y = element_text(size = 14, colour = "black", angle = 0, hjust = 0),
        panel.background = element_rect(fill = "grey98"),
        strip.background = element_rect(fill = "grey90"),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.text.x = element_markdown(size = 14, colour = "black"),
        axis.title = element_blank(),
        legend.title = element_text(size = 16, colour = "black"),
        legend.text = element_text(size = 14, colour = "black"),
        legend.justification = "top",
        legend.spacing.y = unit(0.3, 'cm'),
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm')) + 
  ggtitle('Up in YPD, 37°C')
  
GOslim_up_37 

GOslim_up_30 <- ggplot(
  data= GOslim_37Cvs30C %>% 
    filter(deg == 'up in YPD, 30°C') %>%
    mutate(comparison = factor(comparison, 
                              levels = c("YPD, 37°C vs YPD, 30°C in FLC1", "YPD, 37°C vs YPD, 30°C in flc1∆"),
                              labels = c("*FLC1*", "*flc1*∆")),
            go_type = factor(go_type, 
                            levels = c("Biological process", "Molecular function", "Cellular component"),
                            labels = c("Biological\nprocess", "Molecular  \nfunction", "Cellular\ncomponent"))
           ),
  aes(y = name, x = comparison, size = result_count, colour = fold_enrichment)
) + 
  geom_point() +
  scale_size_continuous(limits = c(1,20), breaks = c(5, 10, 15, 20)) + 
  scale_colour_gradient2(low = "grey95", high = "firebrick", limits =  c(1,30), breaks = c(5, 10, 15, 20, 25, 30)) + 
  geom_point(shape = 1, colour = "black") + 
  labs(size="Number of\nGenes", colour = "Fold\nEnrichment") + 
  guides(colour = guide_legend(order=2),
         size = guide_legend(order=1, override.aes=list(shape = 1))) + 
  facet_grid2(go_type~., scales = 'free_y', space = "free") +
  theme(plot.title = element_markdown(face = "plain", size = 18),
        strip.text.y = element_text(size = 14, colour = "black", angle = 0, hjust = 0),
        panel.background = element_rect(fill = "grey98"),
        strip.background = element_rect(fill = "grey90"),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.text.x = element_markdown(size = 14, colour = "black"),
        axis.title = element_blank(),
        legend.title = element_text(size = 16, colour = "black"),
        legend.text = element_text(size = 14, colour = "black"),
        legend.justification = "top",
        legend.spacing.y = unit(0.3, 'cm'),
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm')) + 
  ggtitle('Down in YPD, 37°C')
  
GOslim_up_30 

```


```{r}
# Open GO analysis of down in EGTA vs YPD at 37C in WT 
MF_EGTAvsYPD_WT_down <- read_tsv(
  here("results/GOanalysis/GO_EGTAvsYPD_WT/GO_EGTAvsYPD_WT_down2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "EGTAvsYPD_WT",
    GO_type = "MF",
    DEG = "down"
  )

BP_EGTAvsYPD_WT_down <- read_tsv(
  here("results/GOanalysis/GO_EGTAvsYPD_WT/GO_EGTAvsYPD_WT_down2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "EGTAvsYPD_WT",
    GO_type = "BP",
    DEG = "down"
  )

#CC_EGTAvsYPD_WT_down <- read_tsv(
 # here("results/GOanalysis/GO_EGTAvsYPD_WT/GO_EGTAvsYPD_WT_down2x_FDR5_cellular_component_slim.tsv")
#) %>% 
 # mutate(
  #  Comparison = "EGTAvsYPD_WT",
   # GO_type = "CC",
    #DEG = "down"
  #)

GO_slim_EGTAvsYPD_WT_down <- bind_rows(
  MF_EGTAvsYPD_WT_down,
  BP_EGTAvsYPD_WT_down) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)

# Open GO analysis of up in EGTA vs YPD at 37C in WT 
MF_EGTAvsYPD_WT_up <- read_tsv(
  here("results/GOanalysis/GO_EGTAvsYPD_WT/GO_EGTAvsYPD_WT_up2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "EGTAvsYPD_WT",
    GO_type = "MF",
    DEG = "up"
  )

BP_EGTAvsYPD_WT_up <- read_tsv(
  here("results/GOanalysis/GO_EGTAvsYPD_WT/GO_EGTAvsYPD_WT_up2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "EGTAvsYPD_WT",
    GO_type = "BP",
    DEG = "up"
  )

CC_EGTAvsYPD_WT_up <- read_tsv(
  here("results/GOanalysis/GO_EGTAvsYPD_WT/GO_EGTAvsYPD_WT_up2x_FDR5_cellular_component_slim.tsv")
) %>% 
  mutate(
    Comparison = "EGTAvsYPD_WT",
    GO_type = "CC",
    DEG = "up"
  )

GO_slim_EGTAvsYPD_WT_up <- bind_rows(
  MF_EGTAvsYPD_WT_up,
  BP_EGTAvsYPD_WT_up
  ) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)

# Open GO analysis of down in EGTA vs YPD at 37C in delta 
MF_EGTAvsYPD_delta_down <- read_tsv(
  here("results/GOanalysis/GO_EGTAvsYPD_delta/GO_EGTAvsYPD_delta_down2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "EGTAvsYPD_delta",
    GO_type = "MF",
    DEG = "down"
  )

BP_EGTAvsYPD_delta_down <- read_tsv(
  here("results/GOanalysis/GO_EGTAvsYPD_delta/GO_EGTAvsYPD_delta_down2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "EGTAvsYPD_delta",
    GO_type = "BP",
    DEG = "down"
  )

CC_EGTAvsYPD_delta_down <- read_tsv(
  here("results/GOanalysis/GO_EGTAvsYPD_delta/GO_EGTAvsYPD_delta_down2x_FDR5_cellular_component_slim.tsv")
) %>% 
  mutate(
    Comparison = "EGTAvsYPD_delta",
    GO_type = "CC",
    DEG = "down"
  )

GO_slim_EGTAvsYPD_delta_down <- bind_rows(
  MF_EGTAvsYPD_delta_down,
  BP_EGTAvsYPD_delta_down,
  CC_EGTAvsYPD_delta_down) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)

# Open GO analysis of up EGTA vs YPD at 37C in delta 
MF_EGTAvsYPD_delta_up <- read_tsv(
  here("results/GOanalysis/GO_EGTAvsYPD_delta/GO_EGTAvsYPD_delta_up2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "EGTAvsYPD_delta",
    GO_type = "MF",
    DEG = "up"
  )

BP_EGTAvsYPD_delta_up <- read_tsv(
  here("results/GOanalysis/GO_EGTAvsYPD_delta/GO_EGTAvsYPD_delta_up2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "EGTAvsYPD_delta",
    GO_type = "BP",
    DEG = "up"
  )

CC_EGTAvsYPD_delta_up <- read_tsv(
  here("results/GOanalysis/GO_EGTAvsYPD_delta/GO_EGTAvsYPD_delta_up2x_FDR5_cellular_component_slim.tsv")
) %>% 
  mutate(
    Comparison = "EGTAvsYPD_delta",
    GO_type = "CC",
    DEG = "up"
  )

GO_slim_EGTAvsYPD_delta_up <- bind_rows(
  MF_EGTAvsYPD_delta_up,
  BP_EGTAvsYPD_delta_up) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)


#Combine into a single dataframe!
GOslim_EGTAvsYPD <- bind_rows(GO_slim_EGTAvsYPD_delta_up, GO_slim_EGTAvsYPD_delta_down, GO_slim_EGTAvsYPD_WT_up, GO_slim_EGTAvsYPD_WT_down) %>% 
  mutate(comparison = factor(comparison, 
                             levels = c("EGTAvsYPD_WT", "EGTAvsYPD_delta"),
                             labels = c("YPD + EGTA, 37°C vs YPD, 37°C in FLC1", "YPD + EGTA, 37°C vs YPD, 37°C in in flc1∆")),
         deg = factor(deg, levels = c("down", "up"), labels = c("up in YPD, 37°C", "up in YPD + EGTA, 37°C"))) %>% 
  ungroup() %>%
  group_by(go_type, comparison) %>% 
  arrange(desc(fold_enrichment), .by_group = T) %>% 
  filter(!name %in% c("cellular component", "molecular function", "biological process", "organelle", "intracellular anatomical structure")) 

GOslim_EGTAvsYPD$name <- factor(GOslim_EGTAvsYPD$name, levels = rev(unique(GOslim_EGTAvsYPD$name)))
```

```{r}
GOslim_up_EGTA <- ggplot(
  data= GOslim_EGTAvsYPD %>% 
    filter(deg == 'up in YPD + EGTA, 37°C') %>%
    mutate(comparison = factor(comparison, 
                              levels = c("YPD + EGTA, 37°C vs YPD, 37°C in FLC1", "YPD + EGTA, 37°C vs YPD, 37°C in in flc1∆"),
                              labels = c("*FLC1*", "*flc1*∆")),
            go_type = factor(go_type, 
                            levels = c("Biological process", "Molecular function", "Cellular component"),
                            labels = c("Biological\nprocess", "Molecular \nfunction", "Cellular\ncomponent"))
           ),
  aes(y = name, x = comparison, size = result_count, colour = fold_enrichment)
) + 
  geom_point() +
  scale_size_continuous(limits = c(1,55), breaks = c(10, 20, 30, 40, 50)) + 
  scale_colour_gradient2(low = "grey95", high = "blue3", limits =  c(1,12), breaks = c(2, 4, 6, 8, 10, 12)) + 
  geom_point(shape = 1, colour = "black") + 
  labs(size="Number of\nGenes", colour = "Fold\nEnrichment") + 
  guides(colour = guide_legend(order=2),
         size = guide_legend(order=1, override.aes=list(shape = 1))) + 
  facet_grid2(go_type~., scales = 'free_y', space = "free") +
  theme(plot.title = element_markdown(face = "plain", size = 18),
        strip.text.y = element_text(size = 14, colour = "black", angle = 0, hjust = 0),
        panel.background = element_rect(fill = "grey98"),
        strip.background = element_rect(fill = "grey90"),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.text.x = element_markdown(size = 14, colour = "black"),
        axis.title = element_blank(),
        legend.title = element_text(size = 16, colour = "black"),
        legend.text = element_text(size = 14, colour = "black"),
        legend.justification = "top",
        legend.spacing.y = unit(0.3, 'cm'),
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm')
    
  ) + 
  ggtitle('Up in YPD + EGTA, 37°C')
  
GOslim_up_EGTA

GOslim_up_YPD <- ggplot(
  data= GOslim_EGTAvsYPD %>% 
    filter(deg == 'up in YPD, 37°C') %>%
    mutate(comparison = factor(comparison, 
                              levels = c("YPD + EGTA, 37°C vs YPD, 37°C in FLC1", "YPD + EGTA, 37°C vs YPD, 37°C in in flc1∆"),
                              labels = c("*FLC1*", "*flc1*∆")),
            go_type = factor(go_type, 
                            levels = c("Biological process", "Molecular function", "Cellular component"),
                            labels = c("Biological\nprocess", "Molecular \nfunction", "Cellular\ncomponent"))
           ),
  aes(y = name, x = comparison, size = result_count, colour = fold_enrichment)
) + 
  geom_point() +
  scale_size_continuous(limits = c(1,55), breaks = c(10, 20, 30, 40, 50)) + 
  scale_colour_gradient2(low = "grey95", high = "firebrick", limits =  c(1,12), breaks = c(2, 4, 6, 8, 10, 12)) + 
  geom_point(shape = 1, colour = "black") + 
  labs(size="Number of\nGenes", colour = "Fold\nEnrichment") + 
  guides(colour = guide_legend(order=2),
         size = guide_legend(order=1, override.aes=list(shape = 1))) + 
  facet_grid2(go_type~., scales = 'free_y', space = "free") +
  theme(plot.title = element_markdown(face = "plain", size = 18),
        strip.text.y = element_text(size = 14, colour = "black", angle = 0, hjust = 0),
        panel.background = element_rect(fill = "grey98"),
        strip.background = element_rect(fill = "grey90"),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.text.x = element_markdown(size = 14, colour = "black"),
        axis.title = element_blank(),
        legend.title = element_text(size = 16, colour = "black"),
        legend.text = element_text(size = 14, colour = "black"),
        legend.justification = "top",
        legend.spacing.y = unit(0.3, 'cm'),
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm')
    
  ) + 
  ggtitle('Down in YPD + EGTA, 37°C')
  
GOslim_up_YPD
```
```{r}

eulerr_list_up <- list( 
  '*FLC1*' = DEGs_up_37Cvs30C_WT %>% .$gene,
  '*flc1*∆' = DEGs_up_37Cvs30C_delta %>% .$gene)

eulerr_list_down <- list( 
  '*FLC1*' = DEGs_down_37Cvs30C_WT %>% .$gene,
  '*flc1*∆' = DEGs_down_37Cvs30C_delta %>% .$gene)

eulerr_plot_up <- plot(
  euler(eulerr_list_up), 
  quantities = list(col = "blue3"), 
  #fills = list(fill = c("grey90", "grey40", "orange"), alpha = 0.4),
  edges = list(col = "blue3"),
  legend = F,
  fontsize = 20, 
  labels = list(labels = c("FLC1", "flc1∆"), font = 3),
  main = expression("Up in YPD, 37°C"))
eulerr_plot_up

eulerr_plot_down <- plot(
  euler(eulerr_list_down), 
  quantities = list(col = "firebrick"), 
  #fills = list(fill = c("grey90", "grey40", "orange"), alpha = 0.4),
  edges = list(col = "firebrick"),
  legend = F,
  fontsize = 20, 
  labels = list(labels = c("FLC1", "flc1∆"), font = 3),
  main = expression("Down in YPD, 37°C"))
eulerr_plot_down

plot_grid(
  eulerr_plot_up,
  eulerr_plot_down,
  ncol = 1,
  scale = 0.8
)


```


```{r}
eulerr_list_up_EGTA <- list( 
  '*FLC1*' = DEGs_up_EGTAvsYPD_37C_WT %>% .$gene,
  '*flc1*∆' = DEGs_up_EGTAvsYPD_37C_delta %>% .$gene)

eulerr_list_down_EGTA <- list( 
  '*FLC1*' = DEGs_down_EGTAvsYPD_37C_WT %>% .$gene,
  '*flc1*∆' = DEGs_down_EGTAvsYPD_37C_delta %>% .$gene)

eulerr_plot_up_EGTA <- plot(
  euler(eulerr_list_up_EGTA), 
  quantities = list(col = "blue3"), 
  #fills = list(fill = c("grey90", "grey40", "orange"), alpha = 0.4),
  edges = list(col = "blue3"),
  legend = F,
  fontsize = 20, 
  labels = list(labels = c("FLC1", "flc1∆"), font = 3),
  main = expression("Up in YPD + EGTA, 37°C"))
eulerr_plot_up

eulerr_plot_down_EGTA <- plot(
  euler(eulerr_list_down_EGTA), 
  quantities = list(col = "firebrick"), 
  #fills = list(fill = c("grey90", "grey40", "orange"), alpha = 0.4),
  edges = list(col = "firebrick"),
  legend = F,
  fontsize = 20, 
  labels = list(labels = c("FLC1", "flc1∆"), font = 3),
  main = expression("Down in YPD + EGTA, 37°C"))
eulerr_plot_down

plot_grid(
  eulerr_plot_up_EGTA,
  eulerr_plot_down_EGTA,
  ncol = 1,
  scale = 0.9
)

```

```{r}
### Barplot Venn alternative 
DEG_counts_temperature <- tibble(
  condition = rep(c("*FLC1* only", "*flc1*∆ only", "*FLC1* & *flc1*∆"), times = 2), 
  direction = rep(c("up in 37°C", "up in 30°C"), each = 3),
  count = c(180, 87, 47, 73, 39, 16)
) %>% 
  mutate(
    condition = factor(condition, levels = c("*FLC1* only", "*flc1*∆ only", "*FLC1* & *flc1*∆")),
    direction = factor(direction, levels = c("up in 37°C", "up in 30°C"), labels = c("up in 37°C", "down in 37°C"))
  ) 

DEG_barplot_temperature <- ggplot(DEG_counts_temperature, aes(x = direction, y = count, fill = condition)) + 
  geom_bar(stat="identity", width = 0.7, colour = "black") + 
  scale_fill_manual(values = c("grey98", "grey65", "grey20")) + 
  ylab("Number of DEGs") + 
  xlab("") +
  scale_y_continuous(limits = c(0,350), expand = c(0.01,0.01)) + 
  theme(axis.text.x = element_markdown(size = 14, colour = rev(c("firebrick", "blue3")), angle = 45, hjust =1),
        legend.title = element_blank(),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.title.y = element_text(size = 14, colour = "black"),
        legend.text = element_markdown(size = 14, colour = "black"),
        legend.key.size = unit(0.5,"line"),
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm'),
        legend.justification = "top") 

DEG_barplot_temperature

DEG_barplot_temperature_horizontal <- ggplot(DEG_counts_temperature, aes(y = direction, x = count, fill = condition)) + 
  geom_bar(stat="identity", width = 0.7, colour = "black") + 
  scale_fill_manual(values = c("grey98", "grey65", "grey20")) + 
  xlab("Number of DEGs") + 
  ylab("") +
  scale_x_continuous(limits = c(0,500), expand = c(0.01,0.01)) + 
  theme(axis.text.y = element_markdown(size = 14, colour = rev(c("firebrick", "blue3")), hjust =1),
        legend.title = element_blank(),
        axis.text.x = element_text(size = 14, colour = "black"),
        axis.title.y = element_text(size = 14, colour = "black"),
        legend.text = element_markdown(size = 14, colour = "black"),
        legend.key.size = unit(0.5,"line"),
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm'),
        legend.justification = "top") 

DEG_barplot_temperature_horizontal


### Barplot Venn alternative 
DEG_counts_EGTA <- tibble(
  condition = rep(c("*FLC1* only", "*flc1*∆ only", "*FLC1* & *flc1*∆"), times = 2), 
  direction = rep(c("up in YPD + EGTA", "down in YPD + EGTA"), each = 3),
  count = c(46, 373, 58, 17, 80, 9)
) %>% 
  mutate(
    condition = factor(condition, levels = c("*FLC1* only", "*flc1*∆ only", "*FLC1* & *flc1*∆")),
    direction = factor(direction, levels = c("up in YPD + EGTA", "down in YPD + EGTA"))
  ) 

DEG_barplot_EGTA <- ggplot(DEG_counts_EGTA, aes(x = direction, y = count, fill = condition)) + 
  geom_bar(stat="identity", width = 0.7, colour = "black") + 
  scale_fill_manual(values = c("grey98", "grey65", "grey20")) + 
  ylab("Number of DEGs") + 
  xlab("") +
  scale_y_continuous(limits = c(0,500), expand = c(0.01,0.01)) + 
  theme(axis.text.x = element_markdown(size = 14, colour = rev(c("firebrick", "blue3")), angle = 45, hjust =1),
        legend.title = element_blank(),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.title.y = element_text(size = 14, colour = "black"),
        legend.text = element_markdown(size = 14, colour = "black"),
        legend.key.size = unit(0.5,"line"),
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm'),
        legend.justification = "top") 

DEG_barplot_EGTA

DEG_barplot_EGTA_horizontal <- ggplot(DEG_counts_EGTA, aes(y = direction, x = count, fill = condition)) + 
  geom_bar(stat="identity", width = 0.7, colour = "black") + 
  scale_fill_manual(values = c("grey98", "grey65", "grey20")) + 
  xlab("Number of DEGs") + 
  ylab("") +
  scale_x_continuous(limits = c(0,500), expand = c(0.01,0.01)) + 
  theme(axis.text.y = element_markdown(size = 14, colour = rev(c("firebrick", "blue3")), hjust =1),
        legend.title = element_blank(),
        axis.text.x = element_text(size = 14, colour = "black"),
        axis.title.y = element_text(size = 14, colour = "black"),
        legend.text = element_markdown(size = 14, colour = "black"),
        legend.key.size = unit(0.5,"line"),
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm'),
        legend.justification = "top") 

DEG_barplot_EGTA_horizontal
```

Multi-panel showing how growth temperature affects the transcriptome of FLC1 cells and flc1 cells

```{r fig.width = 14, fig.height=14}
# Plot Volcano plots 
Volcano_temp_grid <- cowplot::plot_grid(
  volcano_37Cvs30C_WT,
  volcano_37Cvs30C_delta,
  DEG_barplot_temperature,
  axis = 'l',
  align = 'v',
  rel_heights = c(1, 1, 1.5),
  ncol = 1,
  scale = 0.9
)

GOterm_temp_grid <- cowplot::plot_grid(
  GOslim_up_37,
  GOslim_up_30,
  axis = "lr",
  align = "v", 
  ncol =1, 
  rel_heights = c(6, 14),
  scale = 0.9)

plot_grid(
  Volcano_temp_grid,
  GOterm_temp_grid,
  axis = "tb",
  align = "h", 
  rel_widths = c(1,1.5)
)


ggsave(here("results", "figures", "37Cvs30C_multipannel.png"), width = 14, height = 14, dpi = 1000)
```


```{r fig.width = 14, fig.height=14}
# Plot Volcano plots 
Volcano_EGTA_grid <- cowplot::plot_grid(
  volcano_EGTAvsYPD_37C_WT,
  volcano_EGTAvsYPD_37C_delta,
  DEG_barplot_EGTA,
  axis = 'l',
  align = 'v',
  rel_heights = c(1, 1, 1.5),
  ncol = 1,
  scale = 0.9
)

GOterm_EGTA_grid <- cowplot::plot_grid(
  GOslim_up_EGTA,
  GOslim_up_YPD,
  axis = "lr",
  align = "v", 
  ncol =1, 
  rel_heights = c(9, 11),
  scale = 0.9)

plot_grid(
  Volcano_EGTA_grid,
  GOterm_EGTA_grid,
  axis = "tb",
  align = "h", 
  rel_widths = c(1,1.5)
)

ggsave(here("results", "figures", "EGTAvsYPD_multipannel.png"), width = 14, height = 14, dpi = 1000)

```



```{r fig.height=16, fig.width=12}
# Temperature
Volcano_grid_temp <- cowplot::plot_grid(
 volcano_37Cvs30C_WT,
 volcano_37Cvs30C_delta,
 axis = "t",
 align = "h",
 ncol =2, 
 scale = 0.95
)

Venn_grid_temp <- cowplot::plot_grid(
  eulerr_plot_up, 
  eulerr_plot_down,
  NULL,
  ncol = 1,
  scale = 0.9)

GOterm_temp_grid_temp <- cowplot::plot_grid(
  GOslim_up_37,
  GOslim_up_30,
  axis = "lr",
  align = "vh", 
  ncol =1, 
  rel_heights = c(6, 14),
  scale = 0.9)

GOterm_eulerr_grid_temp <- cowplot::plot_grid(
  Venn_grid_temp,
  GOterm_temp_grid_temp,
  ncol = 2,
  rel_widths = c(1, 3),
  scale = 0.9
)

plot_grid(
  Volcano_grid_temp,
  GOterm_eulerr_grid_temp,
  axis = "l",
  align = "h", 
  ncol =1,
  rel_heights = c(1,3),
  rel_widths = c(2,1),
  scale = 0.95
)


```

```{r fig.width = 16, fig.height = 14}
Volcano_grid_temp <- cowplot::plot_grid(
 volcano_37Cvs30C_WT,
 volcano_37Cvs30C_delta,
 eulerr_plot_up, 
 eulerr_plot_down,
 rel_heights = c(1,1,0.7,0.7),
 axis = "l",
 align = "v",
 ncol =1, 
 scale = 0.9
)

GOterm_temp_grid_temp <- cowplot::plot_grid(
  GOslim_up_37,
  GOslim_up_30,
  axis = "lr",
  align = "vh", 
  ncol =1, 
  rel_heights = c(6, 14),
  scale = 0.9)

plot_grid(
  Volcano_grid_temp,
  GOterm_temp_grid_temp,
  ncol = 2,
  rel_widths = c(1, 1.5)
)

ggsave(here("results", "figures", "37Cvs30C_multipannel.png"), width = 14, height = 14, dpi = 1000)
```

```{r fig.width = 16, fig.height = 14}
Volcano_grid_EGTA <- cowplot::plot_grid(
 volcano_EGTAvsYPD_37C_WT,
 volcano_EGTAvsYPD_37C_delta,
 eulerr_plot_up_EGTA, 
 eulerr_plot_down_EGTA,
 rel_heights = c(1,1,0.7,0.7),
 axis = "l",
 align = "v",
 ncol =1, 
 scale = 0.9
)

GOterm_EGTA_grid_EGTA <- cowplot::plot_grid(
  GOslim_up_EGTA,
  GOslim_up_YPD,
  axis = "lr",
  align = "vh", 
  ncol =1, 
  rel_heights = c(9, 11),
  scale = 0.9)

plot_grid(
  Volcano_grid_EGTA,
  GOterm_EGTA_grid_EGTA,
  ncol = 2,
  rel_widths = c(1, 1.5)
)

ggsave(here("results", "figures", "EGTAvsYPD_multipannel.png"), width = 14, height = 14, dpi = 1000)
```

