---
title: "RM_thesis_FLC1vsflc1∆"
author: "Rachael Murray"
date: "2025-04-03"
output: html_document
---
# Summary

In this document I explore differences in the transciptome of KN99alpha cells (FLC1) and flc1∆ cells grown in YPD at 30°C and 37°C. 
Differential gene expression analysis was previously performed (see Rmd/DGE_Wtvsdelta_30C_YPD.rmd and Rmd/DGE_Wtvsdelta_30C_YPD.rmd ) and GO term enrichment analysis was carried out at fungidb.com. Here I make nice figures for my PhD thesis. 

Input files: 
- results/DEGs/deseq_df_WTvsdelta_30C_YPD.txt
- results/DEGs/deseq_df_WTvsdelta_37C_YPD.txt
- results/GOanalysis/GO_WTvsdelta_30C_YPD/
- results/GOanalysis/GO_WTvsdelta_37C_YPD/
- results/gene_info/genes_to_highlight.csv

# Load count data from shared script load_count_data.Rmd

```{r call_load_count_data, child = "load_count_data.Rmd"}
```


# FLC1 vs flc1∆ at 30°C and 37°C 

## Volcano plots

Open results from differential gene expression analysis and create volcano plot of FLC1 vs flc1∆ at 30°C and FLC1 vs flc1∆ at 37°C. Differentially expressed gene cutoff = log2-fold change > 1 and pvalue < 0.05. 

Open results from differential gene expression analysis. 
```{r Open DEseq results from FLC1 vs flc1∆ at 30°C and 37°C}
# FLC1 vs flc1∆ 
deseq_df_WTvsdelta_30C_YPD <- read_tsv(
 file = here::here("results", "DEGs",
                              "deseq_df_WTvsdelta_30C_YPD.txt"))
# Number up in flc1∆ at 30°C 
DEGs_down_30 <- deseq_df_WTvsdelta_30C_YPD %>% 
                     dplyr::filter(log2FC < -1, padj < 0.05) %>% 
                     arrange(log2FC)
# Number up in FLC1 at 30°C %>% 
DEGs_up_30 <- deseq_df_WTvsdelta_30C_YPD %>% 
                     dplyr::filter(log2FC > 1, padj < 0.05) %>% 
                     arrange(log2FC)

deseq_df_WTvsdelta_37C_YPD <- read_tsv(
 file = here::here("results", "DEGs",
                              "deseq_df_WTvsdelta_37C_YPD.txt"))

# Number up in flc1∆ at 37°C 
DEGs_down_37 <- deseq_df_WTvsdelta_37C_YPD %>% 
                     dplyr::filter(log2FC < -1, padj < 0.05) %>% 
                     arrange(log2FC)

# Number up in FLC1 at 37°C %>% 
DEGs_up_37 <- deseq_df_WTvsdelta_37C_YPD %>% 
                     dplyr::filter(log2FC > 1, padj < 0.05) %>% 
                     arrange(log2FC)
```
Make volcano plots for each growth temperature. 
```{r Volcano plots for each comparison, fig.width = 4, fig.height=3}
geom_point_size <- 0.3
limits_x <- c(-6, 6)
colour_positive <- "blue3"
colour_negative <- "firebrick"
colour_neutral <- "grey70"

# Volcano YPD 30°C
volcano_WTvsdelta_30C_YPD <- 
 ggplot(data = deseq_df_WTvsdelta_30C_YPD,
         aes(x = log2FC, y = -log10(padj)) ) +
  geom_point(size = geom_point_size - 0.1, colour = colour_neutral) +
  geom_point(data = deseq_df_WTvsdelta_30C_YPD %>% 
               dplyr::filter(
                 log2FC  > 1, padj < 0.05),
             size = geom_point_size, colour = colour_positive) +
  geom_point(data = deseq_df_WTvsdelta_30C_YPD %>% 
               dplyr::filter(
                 log2FC < -1, padj < 0.05),
             size = geom_point_size, colour = colour_negative) +
  scale_y_continuous("-log10(p)",
                     limits = c(0,50), expand = c(0,0),
                     oob=scales::squish) +
  scale_x_continuous("log2 fold-change\n← up in flc1∆       up in FLC1 →",
                     breaks = -6:6,
                     limits = limits_x, expand = c(0.01,0.01),
                     oob = scales::squish) +
  coord_cartesian(clip = "off") + 
  #ggtitle("FLC1 vs flc1∆ in YPD, 30°C") +
  theme(plot.title = element_markdown(colour = "black", size = 18),
        axis.text.x = element_text(size = 14, colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black"),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.title.y = element_text(size = 14, colour = "black")) 

volcano_WTvsdelta_30C_YPD

# Volcano YPD 37°C
volcano_WTvsdelta_37C_YPD <- 
 ggplot(data = deseq_df_WTvsdelta_37C_YPD,
         aes(x = log2FC, y = -log10(padj)) ) +
  geom_point(size = geom_point_size - 0.1, colour = colour_neutral) +
  geom_point(data = deseq_df_WTvsdelta_37C_YPD %>% 
               dplyr::filter(
                 log2FC  > 1, padj < 0.05),
             size = geom_point_size, colour = colour_positive) +
  geom_point(data = deseq_df_WTvsdelta_37C_YPD %>% 
               dplyr::filter(
                 log2FC < -1, padj < 0.05),
             size = geom_point_size, colour = colour_negative) +
  scale_y_continuous("-log10(p)",
                     limits = c(0,50), expand = c(0,0),
                     oob=scales::squish) +
  scale_x_continuous("log2 fold-change\n← up in flc1∆       up in FLC1 →",
                     breaks = -6:6,
                     limits = limits_x, expand = c(0.01,0.01),
                     oob = scales::squish) +
  coord_cartesian(clip = "off") + 
  #ggtitle("FLC1 vs flc1∆ in YPD, 37°C") +
  theme(plot.title = element_markdown(colour = "black", size = 18),
        axis.text.x = element_text(size = 14, colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black"),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.title.y = element_text(size = 14, colour = "black")) 

volcano_WTvsdelta_37C_YPD
```

## Overlap of differentially expressed genes at 30°C and 37°C in FLC1 vs flc1∆

I wanted to know if the same set of genes are differentially expressed genes in flc1∆ at 30°C and 37°C. I first made some venn diagrams but found that a stacked barplot looks much neater. 

Get genes list of up in flc1∆ at 30°C, up in flc1∆ at 37°C, up in FLC1 at 30°C and up in FLC1 at 37°C. 
```{r fig.width=10, fig.height=7}
# Get list of down genes (up in flc1∆) for each growth condition
DGE_30C_YPD_down <- deseq_df_WTvsdelta_30C_YPD %>% 
                     dplyr::filter(log2FC < -1, padj < 0.05) %>% 
                     .$gene 

DGE_37C_YPD_down <- deseq_df_WTvsdelta_37C_YPD %>% 
                     dplyr::filter(log2FC < -1, padj < 0.05) %>% 
                     .$gene

# Get list of up genes (up in FLC1) for each growth condition
DGE_30C_YPD_up <- deseq_df_WTvsdelta_30C_YPD %>% 
                     dplyr::filter(log2FC > 1, padj < 0.05) %>% 
                     .$gene 

DGE_37C_YPD_up <- deseq_df_WTvsdelta_37C_YPD %>% 
                     dplyr::filter(log2FC > 1, padj < 0.05) %>% 
                     .$gene
```

Plot Venn diagrams
```{r Venn diagram, fig.width=6, fig.height=3}
eulerr_list_down <- list( 
  'YPD, 30°C' = DGE_30C_YPD_down,
  'YPD, 37°C' = DGE_37C_YPD_down)

eulerr_list_up <- list( 
  'YPD, 30°C' = DGE_30C_YPD_up,
  'YPD, 37°C' = DGE_37C_YPD_up)

ggvenn_plot_down <- ggvenn(
  eulerr_list_down,
  show_percentage = F,
  fill_color = c("grey20", "grey60","white"),
  fill_alpha = 0.4,
  set_name_size = 4,
  text_color = "firebrick",
  stroke_color = "firebrick",
  stroke_size = 0.5
) + ggtitle("Up in flc1∆")
ggvenn_plot_down

ggvenn_plot_up <- ggvenn(
  eulerr_list_up,
  show_percentage = F,
  fill_color = c("grey20", "grey60","white"),
  fill_alpha = 0.4,
  #text_size = 2.5,
  set_name_size = 4,
  text_color = "blue3",
  stroke_color = "blue3",
  stroke_size = 0.5
)  + ggtitle("Up in FLC1")
ggvenn_plot_up
```
Stakced barplot 
```{r Barplot, fig.width = 4, fig.height=5}
### Barplot Venn alternative 
DEG_counts <- tibble(
  condition = rep(c("YPD, 30°C only", "YPD, 37°C only", "YPD, 30°C & 37°C"), times = 2), 
  direction = rep(c("up in flc1∆", "up in FLC1"), each = 3),
  count = c(276, 153, 457, 172, 11, 29)
) %>% 
  mutate(
    condition = factor(condition, levels = c("YPD, 30°C only", "YPD, 37°C only", "YPD, 30°C & 37°C")),
    direction = factor(direction, levels = c("up in flc1∆", "up in FLC1"))
  ) 

# Horizonatl bars 
DEG_barplot_30C_37C <- ggplot(DEG_counts, aes(y = direction, x = count, fill = condition)) + 
  geom_bar(stat="identity", colour = "black") + 
  scale_fill_manual(values = c("grey95", "grey65", "grey30")) + 
  xlab("DEGs") + 
  ylab("") +
  scale_x_continuous(limits = c(0,900), expand = c(0.01,0.01)) + 
  theme(axis.text.y = element_markdown(size = 16, colour = c("blue3", "firebrick")),
        legend.title = element_blank(),
        axis.text.x = element_text(size = 12, colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black"),
        legend.text = element_text(size = 14, colour = "black"),
        legend.spacing.y = unit(0.3, 'cm'),
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm'),
        legend.justification = "top")

#DEG_barplot_30C_37C

# Vertical bars
DEG_barplot_30C_37C <- ggplot(DEG_counts, aes(x = direction, y = count, fill = condition)) + 
  geom_bar(stat="identity", width = 0.7, colour = "black") + 
  scale_fill_manual(values = c("grey98", "grey65", "grey30")) + 
  ylab("Number of DEGs") + 
  xlab("") +
  scale_y_continuous(limits = c(0,900), expand = c(0.01,0.01)) + 
  theme(axis.text.x = element_markdown(size = 14, colour = c("firebrick", "blue3"), angle = 45, hjust =1),
        legend.title = element_blank(),
        axis.text.y = element_text(size = 14, colour = "black"),
        axis.title.y = element_text(size = 14, colour = "black"),
        legend.text = element_text(size = 14, colour = "black"),
        legend.key.size = unit(0.5,"line"),
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm'),
        legend.justification = "top") 

DEG_barplot_30C_37C

# Percentages useful for adding to chapter results
# Total up in flc1∆ = 886 
# Both = 51.6%
# 30°C = 31.1%
# 37°C = 17.3%
# Total up in FLC1 = 212
# Both = 29/212 = 13.7%
# 30°C = 81.1% 
# 37°C = 29/212 = 5.2%
```
## GO term analysis
GO term analysis FLC1 vs flc1∆ in YPD, 30°C and FLC1 vs flc1∆ in YPD, 37°C. 

```{r Open GO analysis results}
# Open GO analysis for down regulated gene in WT vs flc1∆ at 30°C
MF_slim_30_down <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_30C_YPD/GO_WTvsdelta_30C_YPD_down2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_30C_YPD",
    GO_type = "MF",
    DEG = "down"
  )

BP_slim_30_down <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_30C_YPD/GO_WTvsdelta_30C_YPD_down2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_30C_YPD",
    GO_type = "BP",
    DEG = "down"
  )

CC_slim_30_down <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_30C_YPD/GO_WTvsdelta_30C_YPD_down2x_FDR5_cellular_component_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_30C_YPD",
    GO_type = "CC",
    DEG = "down"
  )

GO_slim_30_down <- bind_rows(
  MF_slim_30_down,
  BP_slim_30_down,
  CC_slim_30_down) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)

GO_slim_30_down$name <- factor(GO_slim_30_down$name, levels = rev(GO_slim_30_down$name))
  

# Open GO analysis for downregulated gene in WT vs flc1∆ at 37°C
MF_slim_37_down <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_YPD/GO_WTvsdelta_37C_YPD_down2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_YPD",
    GO_type = "MF",
    DEG = "down"
  )

BP_slim_37_down <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_YPD/GO_WTvsdelta_37C_YPD_down2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_YPD",
    GO_type = "BP",
    DEG = "down"
  )

CC_slim_37_down <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_YPD/GO_WTvsdelta_37C_YPD_down2x_FDR5_cellular_component_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_YPD",
    GO_type = "CC",
    DEG = "down"
  )

GO_slim_37_down <- bind_rows(
  MF_slim_37_down,
  BP_slim_37_down,
  CC_slim_37_down) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)

GO_slim_37_down$name <- factor(GO_slim_37_down$name, levels = rev(GO_slim_37_down$name))


# Open GO analysis upregulated gene in WT vs flc1∆ at 30°C
MF_slim_30_up <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_30C_YPD/GO_WTvsdelta_30C_YPD_up2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_30C_YPD",
    GO_type = "MF",
    DEG = "up"
  )

BP_slim_30_up <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_30C_YPD/GO_WTvsdelta_30C_YPD_up2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_30C_YPD",
    GO_type = "BP", ,
    DEG = "up"
  )

CC_slim_30_up <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_30C_YPD/GO_WTvsdelta_30C_YPD_up2x_FDR5_cellular_component_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_30C_YPD",
    GO_type = "CC", ,
    DEG = "up"
  )

GO_slim_30_up <- bind_rows(
  MF_slim_30_up,
  BP_slim_30_up,
  CC_slim_30_up) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)

GO_slim_30_up$name <- factor(GO_slim_30_up$name, levels = rev(GO_slim_30_up$name))
  

# Open GO analysis for upregulated gene in WT vs flc1∆ at 37°C
MF_slim_37_up <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_YPD/GO_WTvsdelta_37C_YPD_up2x_FDR5_molecular_function_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_YPD",
    GO_type = "MF",
    DEG = "up"
  )

BP_slim_37_up <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_YPD/GO_WTvsdelta_37C_YPD_up2x_FDR5_biological_process_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_YPD",
    GO_type = "BP",
    DEG = "up"
  )

CC_slim_37_up <- read_tsv(
  here("results/GOanalysis/GO_WTvsdelta_37C_YPD/GO_WTvsdelta_37C_YPD_up2x_FDR5_cellular_component_slim.tsv")
) %>% 
  mutate(
    Comparison = "WTvsdelta_37C_YPD",
    GO_type = "CC",
    DEG = "up"
  )

GO_slim_37_up <- bind_rows(
  MF_slim_37_up,
  BP_slim_37_up,
  CC_slim_37_up) %>%  
  janitor::clean_names() %>% 
  mutate(
    go_type = factor(go_type, levels = c("BP", "MF", "CC"), labels = c("Biological process", "Molecular function", "Cellular component"))
  ) %>% 
  group_by(go_type) %>% 
  arrange(desc(fold_enrichment), .by_group = T)

GO_slim_37_up$name <- factor(GO_slim_37_up$name, levels = rev(GO_slim_37_up$name))

#Combine into a single dataframe!
GOslim <- bind_rows(GO_slim_37_up, GO_slim_30_up, GO_slim_37_down, GO_slim_30_down) %>% 
  mutate(comparison = factor(comparison, 
                             levels = c("WTvsdelta_30C_YPD", "WTvsdelta_37C_YPD"),
                             labels = c("FLC1 vs flc1∆ in YPD, 30°C", "FLC1 vs flc1∆ in YPD, 37°C")),
         deg = factor(deg, levels = c("down", "up"), labels = c("up in flc1∆", "up in FLC1"))) %>% 
  ungroup() %>%
  group_by(go_type, comparison) %>% 
  arrange(desc(fold_enrichment), .by_group = T) %>% 
  filter(!name %in% c("cellular component", "molecular function", "biological process", "organelle", "intracellular anatomical structure")) 

GOslim$name <- factor(GOslim$name, levels = rev(unique(GOslim$name)))
```
Plot GO term results with fold-enrichment on x-axis, GOterms on y-axis. Size of point = number of genes found for each term, linetype indicates GO term category. Facet by temperature. 
```{r Plot GO terms for 30°C and 37°C, fig.width = 10, fig.height=7}
GOslim_WTvsdelat_YPD_30 <- ggplot(
  data = GOslim %>% filter(comparison == "FLC1 vs flc1∆ in YPD, 30°C")) + 
  geom_segment(aes(x = 0, xend = fold_enrichment, y = name, linetype = go_type)) +
  scale_linetype_manual(values = c("solid", "twodash", "dotted")) + 
  ggnewscale::new_scale_color() +  
  geom_point(aes(x = fold_enrichment, y = name, size = result_count, colour = deg)) +
  scale_size_continuous(limits = c(1,80), breaks = c(20, 40, 60, 80)) + 
  scale_colour_manual(values = c("firebrick", "blue3")) + 
    labs(size="Number of Genes", linetype="GO term", colour = "Gene Expression") + 
    xlab("Fold enrichment") + 
    ylab("") + 
    scale_x_continuous(breaks = c(0,2,4,6,8,10,12,14), limits = c(0,14), expand = c(0.01,0.01)) +
    facet_grid2(deg~., scales = "free_y", axes = "all", space = "free") +
    #ggtitle("FLC1 vs flc1∆ in YPD, 30°C") +
    theme(strip.text = element_text(colour = "white"),
          axis.text.y = element_text(size = 14, colour = "black"),
          axis.text.x = element_text(size = 14, colour = "black"),
          axis.title.x = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.justification = "top",
          legend.spacing.y = unit(0.3, 'cm'),
          legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm')
          ) + 
  guides(linetype = guide_legend(order=1),
         colour = guide_legend(order=2),
         size = guide_legend(order=3)) 

GOslim_WTvsdelat_YPD_30

GOslim_WTvsdelat_YPD_37 <- ggplot(
  data = GOslim %>% filter(comparison == "FLC1 vs flc1∆ in YPD, 37°C")) + 
  geom_segment(aes(x = 0, xend = fold_enrichment, y = name, linetype = go_type)) +
  scale_linetype_manual(values = c("solid", "twodash", "dotted")) + 
  ggnewscale::new_scale_color() +  
  geom_point(aes(x = fold_enrichment, y = name, size = result_count, colour = deg)) +
  scale_size_continuous(limits = c(1,80), breaks = c(20, 40, 60, 80)) + 
  scale_colour_manual(values = c("firebrick", "blue3")) + 
    labs(size="Number of Genes", linetype="GO term", colour = "Gene Expression") + 
    xlab("Fold enrichment") + 
    ylab("") + 
    scale_x_continuous(breaks = c(0,2,4,6,8,10,12,14), limits = c(0,14), expand = c(0.01,0.01)) +
    facet_grid2(deg~., scales = "free_y", axes = "all", space = "free") +
    #ggtitle("FLC1 vs flc1∆ in YPD, 37°C") +
    theme(strip.text = element_text(colour = "white"),
          axis.text.y = element_text(size = 14, colour = "black"),
          axis.text.x = element_text(size = 14, colour = "black"),
          axis.title.x = element_text(size = 14, colour = "black"),
          legend.title = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 14, colour = "black"),
          legend.justification = "top",
          legend.spacing.y = unit(0.3, 'cm'),
          legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = 'cm')
          ) + 
  guides(linetype = guide_legend(order=1),
         colour = guide_legend(order=2),
         size = guide_legend(order=3))

GOslim_WTvsdelat_YPD_37
```
## Muti-panel figure DGE in FLC1vsflc1∆ at 30°C and 37°C

```{r Mulitpannel FLC1vsflc1∆ fig.width=16, fig.height=12}
Volcano <- cowplot::plot_grid(
  volcano_WTvsdelta_30C_YPD + theme(plot.title = element_blank()),
  volcano_WTvsdelta_37C_YPD + theme(plot.title = element_blank()),
  DEG_barplot_30C_37C,
  axis = "l",
  align = "v",
  ncol =1, 
  rel_heights = c(1,1,1),
  rel_widths = c(1, 1, 1),
  scale = 0.9
)


DEG_GO <- cowplot::plot_grid(
  GOslim_WTvsdelat_YPD_30,
  GOslim_WTvsdelat_YPD_37 + theme(legend.position = "none"), 
  axis = "lr",
  align = "v", 
  ncol =1, 
  rel_heights = c(23,15),
  scale = 0.9)

plot_grid(
  Volcano,
  DEG_GO,
  rel_widths = c(1,2)
)

#ggsave(here("results", "figures", "WTvsDelta_30C_37C_pannel_no_titles.png"), width = 16, height = 12, dpi = 1200)
```

# Additional figures for FLC1 vs flc1∆


```{r genes associated with GO temrs}
#I used these tables to looks for genes associated with GO terms of interest (transmembrane transport, carbohydrate processing) to highlight in the text of my write-up. 

# H99 genome info 
gene_info_df <- 
  here::here("input", "annotation",
             "H99_geneinfo_FungiDB59.txt") %>%
  readr::read_tsv(comment = "#")  %>% 
  dplyr::rename(gene = GeneID, geneName = GeneName)


TM_transport_30 <- GO_slim_30_down %>% 
  ungroup() %>%
  filter(name == "transmembrane transport") %>% 
  select(result_gene_list) %>% 
  tidyr::separate_rows(result_gene_list, sep = ",") %>% 
  dplyr::rename(gene = result_gene_list) %>% 
  left_join(DEGs_down_30) %>% 
  arrange(log2FC) %>% 
  left_join(gene_info_df)
TM_transport_30

TM_transport_37 <- GO_slim_37_down %>% 
  ungroup() %>%
  filter(name == "transmembrane transport") %>% 
  select(result_gene_list) %>% 
  tidyr::separate_rows(result_gene_list, sep = ",") %>% 
  dplyr::rename(gene = result_gene_list) %>% 
  left_join(DEGs_down_37) %>% 
  arrange(log2FC) %>% 
  left_join(gene_info_df)
TM_transport_37

Carb_process_30 <- GO_slim_30_down %>% 
  ungroup() %>%
  filter(name == "carbohydrate metabolic process") %>% 
  select(result_gene_list) %>% 
  tidyr::separate_rows(result_gene_list, sep = ",") %>% 
  dplyr::rename(gene = result_gene_list) %>% 
  left_join(DEGs_down_30) %>% 
  arrange(log2FC) %>% 
  left_join(gene_info_df)
Carb_process_30

Carb_process_37 <- GO_slim_37_down %>% 
  ungroup() %>%
  filter(name == "carbohydrate metabolic process") %>% 
  select(result_gene_list) %>% 
  tidyr::separate_rows(result_gene_list, sep = ",") %>% 
  dplyr::rename(gene = result_gene_list) %>% 
  left_join(DEGs_down_37) %>% 
  arrange(log2FC) %>% 
  left_join(gene_info_df)
Carb_process_37
```

## Plot mean TPM for genes of interest

```{r count_functions}
# Function to calculate transcipts per million
calculate_cpm <- function(x, round_digits = NA) {
    cpm <- x / sum(x, na.rm=TRUE) * 1e6
    if( !is.na(round_digits) ) {
        cpm <- round(cpm, digits = round_digits)
    }
    cpm
}
```


```{r load_counts}
# Load sample sheet and transcipt counts 
sample_sheet <- 
  here::here("input", "experiment","CryptoRNAseqFLC1SampleKey.csv") %>%
  read_csv() %>%
  select(!Strain) %>%
  mutate(FLC1 = factor(FLC1, 
                       levels = c("WT", "delta"),
                       labels = c("*FLC1*", "*flc1*∆")),
         Condition = paste(Media, "_", Temperature, sep = ""),
         BioRep = as.character(BioRep)
  ) %>% 
  mutate(Condition = factor(Condition, 
                            levels = c("YPD_30", "YPD_37", "YPD_CFW_37", "YPD_EGTA_37", "YPD_CFW_EGTA_37"),
                            labels = c("YPD\n30°C", "YPD\n37°C", "YPD+\nCFW\n37°C", "YPD+\nEGTA\n37°C", "YPD+\nCFW\n+EGTA\n37°C"))
  )


counts_bygene_wide <- 
  here::here("results","counts", "counts.txt") %>%
  read_tsv(comment = "#")
counts_bygene_wide
```


```{r make_tidy_counts, dependson = "load_counts"}
# Calculate TPM and combine with sample sheet 
counts_bygene_long <- counts_bygene_wide %>%
    select(-Chr,-Start,-End,-Strand,-Length) %>%
    pivot_longer(ends_with(".bam"), names_to = "SampleID", values_to = "Count") %>%
    mutate(SampleID = as.numeric(word(SampleID, 2, sep = "_"))) %>%
    group_by(SampleID) %>%
    mutate(TPM = calculate_cpm(Count, round_digits = 2)) %>%
    ungroup() %>%
    left_join(sample_sheet, by = "SampleID") %>%
    select(SampleID, SampleName, FLC1, Condition, Media, Temperature, BioRep, Geneid, Count, TPM) %>%
    arrange(Geneid, FLC1, Condition, Media, Temperature, BioRep)

counts_bygene_long
```

```{r Get mean TPM of bioreps}
mean_df <- counts_bygene_long %>% 
  group_by(Condition, FLC1, Geneid) %>% 
  summarise(mean= mean(TPM),
            sd = sd(TPM),
            se = sd(TPM)/sqrt(n())) %>% 
  dplyr::rename("gene" = "Geneid") %>% 
  left_join(gene_info_df) %>% 
  mutate(label = case_when(is.na(geneName) ~ gene,
                           .default = paste(gene, geneName))
         ) 

```
## Cell wall synthesis and remodelling genes

I curated a list of genes involved in cell wall chitin and glucan synthesis/remodelling. Here I plot the mean transcripts per million calculated from 3 biological replicated of FLC1 and flc1∆ samples at 30°C and 37°C. Although not all genes were classified as differentially express, almost all cell wall genes are more highly exressed in flc1∆ samples that FLC1 at both growth temperatures. 

Error bars are standard deviation of the mean. 
* = differentially expressed at 30°C and 37°C
† = differentially expressed at 30°C

```{r Cell wall genes TPM plots, fig.width = 9, fig.height=10}
### Open select groups of genes 
geneGroups <- read_csv(
  here("results/gene_info/genes_to_highlight.csv")
) %>% 
  mutate(plot_title = case_when(is.na(geneName) ~ paste(geneNameLong, gene, sep = "\n"),
                                .default = paste(geneNameLong, geneName, sep = "\n"))
  ) 

### Plot a grid for genes involved in chitin sythesis and remodelling

Chitin_genes <- geneGroups %>% 
  left_join(mean_df) %>% 
  filter(geneGroupSpecific %in% c("chitin_synthase", "chitin_deacetylase", "chitinase"))

chitin_levels <- c("Chitin synthase 1\nCHS1", "Chitin synthase 2\nCHS2", "Chitin synthase 3\nCHS3", "Chitin synthase 4\nCHS4",   
  "Chitin synthase 5\nCHS5", "Chitin synthase 6\nCHS6", "Chitin synthase 7\nCHS7", "Chitin synthase 8\nCHS8",   
  "Chitin deacetylase 1\nCDA1", "Chitin deacetylase 2\nMP98", "Chitin deacetylase 3\nCDA3", "Polysaccharide deacetylase\nFPD1",
  "Chitinase\nCHI2", "Chitinase\nCHI21", "Chitinase\nCHI22", "Endochitinase\nCHI4")

chitin_labels <- c("Chitin synthase 1<br>CHS1", "Chitin synthase 2<br>CHS2", "Chitin synthase 3<br>CHS3", "Chitin synthase 4<br>CHS4<sup>*</sup>",
  "Chitin synthase 5<br>CHS5<sup>*</sup>", "Chitin synthase 6<br>CHS6", "Chitin synthase 7<br>CHS7<sup>*</sup>", "Chitin synthase 8<br>CHS8<sup>†</sup>",   
  "Chitin deacetylase 1<br>CDA1<sup>†</sup>", "Chitin deacetylase 2<br>MP98<sup>†</sup>", "Chitin deacetylase 3<br>CDA3", "Polysaccharide deacetylase<br>FPD1",
  "Chitinase<br>CHI2<sup>*</sup>", "Chitinase<br>CHI21", "Chitinase<br>CHI22", "Endochitinase<br>CHI4")

Chitin_genes$plot_title <- factor(Chitin_genes$plot_title,levels=chitin_levels, labels=chitin_levels)

ggplot(data = filter(
  Chitin_genes, 
  Condition %in% c("YPD\n30°C", "YPD\n37°C", "YPD+\nEGTA\n37°C")), 
  aes(x = Condition, y = mean, fill = FLC1)) + 
    geom_bar(position=position_dodge(), stat="identity", colour = "black", width = 0.7, linewidth = 0.3) +
    geom_errorbar(aes(ymin=mean - se, ymax=mean + se),
                  width=.2,# Width of the error bars
                  linewidth = 0.3,
                  position=position_dodge(.7)) +
  facet_wrap(~plot_title, scales = "free", ncol= 4, drop = F) + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.01))) + 
  #scale_y_log10()+
  theme(
    legend.justification = "left",
    legend.position = "bottom"
  ) + 
  xlab("") + 
  ylab("Transcripts per Million") + 
  theme(strip.text.x = element_text(hjust = 0, size = 8),
        axis.text = element_text(hjust = 0.5, size = 9),
        axis.title.y = element_text(size=12, margin = margin(t = 0, r = 10, b = 0, l = 0)),
        legend.text = element_markdown(size = 12),
        legend.title = element_blank())

ggsave(here("results/figures/Chitin_genes_TPM.png"), width = 9, height = 8, dpi = 1000)

deseq_df_WTvsdelta_30C_YPD %>% 
  filter(gene %in% Chitin_genes$gene) %>% 
  arrange(log2FC)
deseq_df_WTvsdelta_37C_YPD %>% 
  filter(gene %in% Chitin_genes$gene) %>% 
  arrange(log2FC)

### Plot a grid for genes involved in glucan sythesis and remodelling

Beta_glucan_genes <- geneGroups %>% 
  left_join(mean_df) %>% 
  filter(geneGroupSpecific %in% c("beta_1_3_glucan", "beta_1_6_glucan", "alpha_1_3_glucan"))

beta_glucan_levels <- c("1,3-beta-glucan synthase component\nFKS1","Glucan 1,3-beta-glucosidase\nCNAG_00373","Glucan 1,3-beta-glucosidase\nCNAG_05138", "Glucan 1,3-beta-glucosidase\nEXG104","Glucan 1,3-beta-glucosidase\nCNAG_06336", "Exo-beta-1,3-glucanase\nCNAG_05803","Endo-1,3(4)-beta-glucanase\nEBG1","1,3-beta-glucanosyltransferase\nCNAG_06501", "Alpha-1,3-glucan synthase\nAGS1","Glucan endo-1,3-alpha-glucosidase\nCNAG_02850","Glucan endo-1,3-alpha-glucosidase\nCNAG_07736"," ", "Glucosyltransferase\nKRE5","Glucosidase\nKRE6","Putative glucosidase\nSKN1","Glucosidase\nKRE61","Glucosidase\nKRE62","Beta-glucan synthesis-associated protein\nKRE63","Glucosidase\nKRE64")

beta_glucan_labels <- c("1,3-beta-glucan synthase component\nFKS1","Glucan 1,3-beta-glucosidase\nCNAG_00373","Glucan 1,3-beta-glucosidase\nCNAG_05138", "Glucan 1,3-beta-glucosidase\nEXG104","Glucan 1,3-beta-glucosidase\nCNAG_06336", "Exo-beta-1,3-glucanase\nCNAG_05803","Endo-1,3(4)-beta-glucanase\nEBG1","1,3-beta-glucanosyltransferase\nCNAG_06501", "Alpha-1,3-glucan synthase\nAGS1","Glucan endo-1,3-alpha-glucosidase\nCNAG_02850","Glucan endo-1,3-alpha-glucosidase\nCNAG_07736"," ", "Glucosyltransferase\nKRE5","Glucosidase\nKRE6","Putative glucosidase\nSKN1","Glucosidase\nKRE61","Glucosidase\nKRE62","Beta-glucan synthesis-associated\nKRE63","Glucosidase\nKRE64")

Beta_glucan_genes$plot_title <- factor(Beta_glucan_genes$plot_title,levels=beta_glucan_levels,labels=beta_glucan_labels)

ggplot(data = filter(
  Beta_glucan_genes, 
  Condition %in% c("YPD\n30°C", "YPD\n37°C", "YPD+\nEGTA\n37°C")), 
  aes(x = Condition, y = mean, fill = FLC1)) + 
    geom_bar(position=position_dodge(), stat="identity", colour = "black", width = 0.7, linewidth = 0.3) +
    geom_errorbar(aes(ymin=mean - se, ymax=mean + se),
                  width=.2,# Width of the error bars
                  linewidth = 0.3,
                  position=position_dodge(.7)) +
  facet_wrap(~plot_title, scales = "free", ncol= 4, drop = F) + 
  theme(
    legend.justification = "left",
    legend.position = "bottom"
  ) + 
  xlab("") + 
  ylab("Transcripts per Million") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.01))) + 
  theme(strip.text.x = element_text(hjust = 0, size = 8),
        axis.text = element_text(hjust = 0.5, size = 9),
        axis.title.y = element_text(size=12, margin = margin(t = 0, r = 10, b = 0, l = 0)),
        legend.text = element_markdown(size = 12),
        legend.title = element_blank())

ggsave(here("results/figures/Beta_glucan_genes_TPM.png"), width = 9, height = 10, dpi = 1000)

deseq_df_WTvsdelta_30C_YPD %>% 
  filter(gene %in% Beta_glucan_genes$gene) %>% 
  arrange(log2FC)
deseq_df_WTvsdelta_37C_YPD %>% 
  filter(gene %in% Beta_glucan_genes$gene) %>% 
  arrange(log2FC)

```


```{r fig.width =10, fig.height=4}
### Plot a grid putative inostitol transporters
Inositol_genes <- geneGroups %>% 
  left_join(mean_df) %>% 
  filter(geneGroupSpecific == "inositol_transporter")

inositol_levels <- c("myo-inositol transporter\nITR1", "myo-inositol transporter\nITR1A", "myo-inositol transporter\nITR2", "myo-inositol transporter\nITR3", "myo-inositol transporter\nITR3A", "myo-inositol transporter\nITR3B", "myo-inositol transporter\nITR3C", "Polyol transporter protein 1\nPTP1", "myo-inositol transporter\nITR5", "myo-inositol transporter\nITR6")  

inositol_labels <- c("Myo-inositol transporter\nITR1", "Myo-inositol transporter\nITR1A", "Myo-inositol transporter\nITR2", "Myo-inositol transporter\nITR3", "Myo-inositol transporter\nITR3A", "Myo-inositol transporter\nITR3B", "Myo-inositol transporter\nITR3C", "Polyol transporter protein 1\nPTP1/ITR4", "Myo-inositol transporter\nITR5", "Myo-inositol transporter\nITR6") 

Inositol_genes$plot_title <- factor(Inositol_genes$plot_title,levels=inositol_levels,labels=inositol_labels)

ggplot(data = filter(
  Inositol_genes, 
  Condition %in% c("YPD\n30°C", "YPD\n37°C", "YPD+\nEGTA\n37°C")), 
  aes(x = Condition, y = mean, fill = FLC1)) + 
    geom_bar(position=position_dodge(), stat="identity", colour = "black", width = 0.7, linewidth = 0.3) +
    geom_errorbar(aes(ymin=mean - se, ymax=mean + se),
                  width=.2,# Width of the error bars
                  linewidth = 0.3,
                  position=position_dodge(.7)) +
  facet_wrap(~plot_title, scales = "free", ncol= 5, drop = F) + 
  theme(
    legend.justification = "left",
    legend.position = "bottom"
  ) + 
  xlab("") + 
  ylab("Transcripts per Million") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.01))) + 
  theme(strip.text.x = element_text(hjust = 0, size = 8),
        axis.text = element_text(hjust = 0.5, size = 9),
        axis.title.y = element_text(size=12, margin = margin(t = 0, r = 10, b = 0, l = 0)),
        legend.text = element_markdown(size = 12),
        legend.title = element_blank())

ggsave(here("results/figures/Inositol_transporter_genes_TPM.png"), width = 10, height = 4.5, dpi = 1000)

deseq_df_WTvsdelta_30C_YPD %>% 
  filter(gene %in% Inositol_genes$gene) %>% 
  arrange(log2FC)
deseq_df_WTvsdelta_37C_YPD %>% 
  filter(gene %in% Inositol_genes$gene) %>% 
  arrange(log2FC)
```

```{r fig.width=10,fig.height=6}
### Plot a grid putative inostitol transporters
TF_genes <- geneGroups %>% 
  left_join(mean_df) %>% 
  filter(geneGroup == "transcription_factor") %>% 
  arrange(geneGroupSpecific) %>% 
  mutate(plot_title2 = paste(geneGroupSpecific, geneName, sep = "\n"))

TF_levels <- c("BZP family\nBZP3", "BZP family\nBZP4", "BZP family\nCLR3", "C2Z family\nRIM101", "C2Z family\nZAP104", "FZC family\nFZC30", "FZC family\nFZC51", "FZC family\nFZC34", "FZC family\nNA", "FZC family\nYRM103",  "FZC family\nARO8001", "FZC family\nFZC27", "FZC family\nSTB4", "HLH family\nMLN1", "HLH family\nHLH1")

TF_labels <- c("BZP family\nBZP3", "BZP family\nBZP4", "BZP family\nCLR3", "C2Z family\nRIM101", "C2Z family\nZAP104", "FZC family\nFZC30", "FZC family\nFZC51", "FZC family\nFZC34", "FZC family\nFZC9", "FZC family\nYRM103",  "FZC family\nARO8001", "FZC family\nFZC27", "FZC family\nSTB4", "HLH family\nMLN1", "HLH family\nHLH1")

TF_genes$plot_title2 <- factor(TF_genes$plot_title2,levels=TF_levels, labels=TF_labels)

ggplot(data = filter(
  TF_genes, 
  Condition %in% c("YPD\n30°C", "YPD\n37°C", "YPD+\nEGTA\n37°C")), 
  aes(x = Condition, y = mean, fill = FLC1)) + 
    geom_bar(position=position_dodge(), stat="identity", colour = "black", width = 0.7, linewidth = 0.3) +
    geom_errorbar(aes(ymin=mean - se, ymax=mean + se),
                  width=.2,# Width of the error bars
                  linewidth = 0.3,
                  position=position_dodge(.7)) +
  facet_wrap(~~plot_title2, scales = "free", ncol= 5, drop = F) + 
  theme(
    legend.justification = "left",
    legend.position = "bottom"
  ) + 
  xlab("") + 
  ylab("Transcripts per Million") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.01))) + 
  theme(strip.text.x = element_text(hjust = 0, size = 8),
        axis.text = element_text(hjust = 0.5, size = 9),
        axis.title.y = element_text(size=12, margin = margin(t = 0, r = 10, b = 0, l = 0)),
        legend.text = element_markdown(size = 12),
        legend.title = element_blank())


ggsave(here("results/figures/TF_transporter_genes_TPM.png"), width = 10, height = 7, dpi = 1000)
```


```{r fig.width=10,fig.height=6}
Ca2_genes <- geneGroups %>% 
  left_join(mean_df) %>% 
  filter(geneGroup == "calcium_transport") 

Ca2_genes$plot_title <- factor(Ca2_genes$plot_title,levels=unique(Ca2_genes$plot_title))


ggplot(data = filter(
  Ca2_genes, 
  Condition %in% c("YPD\n30°C", "YPD\n37°C", "YPD+\nEGTA\n37°C")), 
  aes(x = Condition, y = mean, fill = FLC1)) + 
    geom_bar(position=position_dodge(), stat="identity", colour = "black", width = 0.7, linewidth = 0.3) +
    geom_errorbar(aes(ymin=mean - se, ymax=mean + se),
                  width=.2,# Width of the error bars
                  linewidth = 0.3,
                  position=position_dodge(.7)) +
  facet_wrap(~~plot_title, scales = "free", ncol= 4, drop = F) + 
  theme(
    legend.justification = "left",
    legend.position = "bottom"
  ) + 
  xlab("") + 
  ylab("Transcripts per Million") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.01))) + 
  theme(strip.text.x = element_text(hjust = 0, size = 8),
        axis.text = element_text(hjust = 0.5, size = 9),
        axis.title.y = element_text(size=12, margin = margin(t = 0, r = 10, b = 0, l = 0)),
        legend.text = element_markdown(size = 12),
        legend.title = element_blank())


ggsave(here("results/figures/Calcium_transporter_genes_TPM.png"), width = 9.5, height = 5, dpi = 1000)
```


```{r fig.width=7,fig.height=3}
Cu2_genes <- geneGroups %>% 
  left_join(mean_df) %>% 
  filter(geneGroup == "copper_transport") 

Cu2_genes$plot_title <- factor(Cu2_genes$plot_title,levels=unique(Cu2_genes$plot_title))

ggplot(data = filter(
  Cu2_genes, 
  Condition %in% c("YPD\n30°C", "YPD\n37°C", "YPD+\nCFW\n37°C")), 
  aes(x = Condition, y = mean, fill = FLC1)) + 
    geom_bar(position=position_dodge(), stat="identity", colour = "black", width = 0.7, linewidth = 0.3) +
    geom_errorbar(aes(ymin=mean - se, ymax=mean + se),
                  width=.2,# Width of the error bars
                  linewidth = 0.3,
                  position=position_dodge(.7)) +
  facet_wrap(~~plot_title, scales = "free", ncol= 4, drop = F) + 
  theme(
    legend.justification = "left",
    legend.position = "bottom"
  ) + 
  xlab("") + 
  ylab("Transcripts per Million") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.01))) + 
  theme(strip.text.x = element_text(hjust = 0, size = 9),
        axis.text = element_text(hjust = 0.5, size = 9),
        axis.title.y = element_text(size=12, margin = margin(t = 0, r = 1, b = 0, l = 0)),
        legend.text = element_markdown(size = 12),
        legend.title = element_blank())

ggsave(here("results/figures/Copper_transporter_genes_TPM.png"), width = 7, height = 3, dpi = 1000)

```


```{r}
FLC_genes <- geneGroups %>% 
  left_join(mean_df) %>% 
  filter(geneGroup == "FLC_proteins") 

FLC_genes$plot_title <- factor(FLC_genes$plot_title,levels=unique(FLC_genes$plot_title))

ggplot(data = filter(
  FLC_genes, 
  Condition %in% c("YPD\n30°C", "YPD\n37°C", "YPD+\nCFW\n37°C", "YPD+\nEGTA\n37°C", "YPD+\nCFW\n+EGTA\n37°C")), 
  aes(x = Condition, y = mean, fill = FLC1)) + 
    geom_bar(position=position_dodge(), stat="identity", colour = "black", width = 0.7, linewidth = 0.3) +
    geom_errorbar(aes(ymin=mean - se, ymax=mean + se),
                  width=.2,# Width of the error bars
                  linewidth = 0.3,
                  position=position_dodge(.7)) +
  facet_wrap(~~plot_title, scales = "free", ncol= 4, drop = F) + 
  theme(
    legend.justification = "left",
    legend.position = "bottom"
  ) + 
  xlab("") + 
  ylab("Transcripts per Million") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.01))) + 
  theme(strip.text.x = element_text(hjust = 0, size = 9),
        axis.text = element_text(hjust = 0.5, size = 9),
        axis.title.y = element_text(size=12, margin = margin(t = 0, r = 1, b = 0, l = 0)),
        legend.text = element_markdown(size = 12),
        legend.title = element_blank())

```









```{r}
Crz1Cna1 <- read_csv(
  here("Chow2017_Crz1_Cna1_targets/Chow2017_DEG_summary_log2foldchange.csv")
) %>% 
  mutate(gene = case_when(nchar(gene) == 4 ~ paste("CNAG_0", gene, sep = ""),
                            nchar(gene) == 3 ~ paste("CNAG_00", gene, sep = ""),
                            nchar(gene) == 2 ~ paste("CNAG_000", gene, sep = ""),
                            nchar(gene) == 1 ~ paste("CNAG_0000", gene, sep = "")))

Cna1 <- Crz1Cna1 %>% 
  filter(cna1 == "yes") %>% 
  left_join(deseq_df_WTvsdelta_37C_YPD) 

Czr1 <- Crz1Cna1 %>% 
  filter(crz1 == "yes") %>% 
  left_join(deseq_df_EGTAvsYPD_37C_delta) %>% 
  arrange(log2_foldchange)
  


X <- deseq_df_WTvsdelta_30C_YPD %>% 
  select(1:4) %>% 
  dplyr::rename('baseMean30' = 'baseMean',
                'log2FC30' = 'log2FC'
                ) %>%
  left_join(deseq_df_EGTAvsYPD_37C_delta)

ggplot(X, aes(x = log2FC, y = log2FC30)) + 
  geom_point() + 
  geom_text_repel(data = X, aes(label = paste(gene, geneName)), size = 2, colour = "red")

ggplot(data = Czr1 , aes(x = log2FC, y = as.numeric(log2_foldchange))) + 
  geom_point() + 
  geom_text_repel(data = Czr1, aes(label = paste(gene, geneName)), size = 2, colour = "red")

deseq_df_WTvsdelta_37C_YPD

library(ggrepel)
```

