---
title: "RM_thesis_TPM_plots"
author: "Rachael Murray"
date: "2025-07-08"
output: html_document
---

In this document I take the 


```{r load_samplesheet}
sample_sheet <- 
  here::here("input", "experiment","CryptoRNAseqFLC1SampleKey.csv") %>%
  read_csv() %>%
  select(!Strain) %>%
  mutate(FLC1 = factor(FLC1, 
                       levels = c("WT", "delta"),
                       labels = c("*FLC1*", "*flc1*∆")),
         Condition = paste(Media, "_", Temperature, sep = ""),
         BioRep = as.character(BioRep)
  ) %>% 
  mutate(Condition = factor(Condition, 
                            levels = c("YPD_30", "YPD_37", "YPD_CFW_37", "YPD_EGTA_37", "YPD_CFW_EGTA_37"),
                            labels = c("YPD\n30°C", "YPD\n37°C", "YPD+\nCFW\n37°C", "YPD+\nEGTA\n37°C", "YPD+\nCFW\n+EGTA\n37°C")),
         BioRep = paste("BioRep", BioRep)
  )


sample_sheet
```

```{r load counts data}
counts_bygene_wide <- 
  here::here("results","counts", "counts.txt") %>%
  read_tsv(comment = "#")

counts_bygene_wide
```

```{r make_tidy_counts, dependson = "load_counts"}
# Calculate TPM and combine with sample sheet 
counts_bygene_long <- counts_bygene_wide %>%
    select(-Chr,-Start,-End,-Strand,-Length) %>%
    pivot_longer(ends_with(".bam"), names_to = "SampleID", values_to = "Count") %>%
    mutate(SampleID = as.numeric(word(SampleID, 2, sep = "_"))) %>%
    group_by(SampleID) %>%
    mutate(TPM = calculate_cpm(Count, round_digits = 2)) %>%
    ungroup() %>%
    left_join(sample_sheet, by = "SampleID") %>%
    select(SampleID, SampleName, FLC1, Condition, Media, Temperature, BioRep, Geneid, Count, TPM) %>%
    arrange(Geneid, FLC1, Condition, Media, Temperature, BioRep)

counts_bygene_long
```

```{r Get mean TPM of bioreps}
mean_df <- counts_bygene_long %>% 
  group_by(Condition, FLC1, Geneid) %>% 
  summarise(mean= mean(TPM),
            sd = sd(TPM),
            se = sd(TPM)/sqrt(n())) %>% 
  dplyr::rename("gene" = "Geneid") %>% 
  left_join(gene_info_df) %>% 
  mutate(label = case_when(is.na(geneName) ~ gene,
                           .default = paste(gene, geneName))
         ) 

mean_df
```



```{r plot_my_gene, dependson = "make_tidy_counts"}
plot_my_genes_TPM <- function(mygenes = "CNAG_06125",
                              mydata = counts_bygene_long,
                              mycondition = c("YPD", "YPD_EGTA", "YPD_CFW"),
                              mytemperature = c(30, 37)){
    ggplot(data = filter(mydata, 
                         Geneid %in% mygenes,
                         Condition %in% mycondition),
                         aes(x = Condition, y = TPM, colour = FLC1)) +
        geom_point(aes(shape = BioRep), size = 1, alpha = 0.7) +
        stat_summary(aes(group=FLC1), fun.min="mean",fun.max="mean",
                     geom="errorbar",linewidth=1,width=0.4) +  
        guides(shape = guide_legend(override.aes = list(size = 1, alpha = 1)),
               color = guide_legend(override.aes = list(linetype = 1, shape = NA))) +
        #scale_y_log10() +
        theme(axis.title.x=element_blank()) + 
        facet_grid(.~Geneid) 
}


plot_my_genes_TPM
```


```{r}
### Open select groups of genes 
geneGroups <- read_csv(
  here("results/gene_info/genes_to_highlight.csv")
) %>% 
  mutate(plot_title = case_when(is.na(geneName) ~ paste(geneNameLong, gene, sep = "\n"),
                                .default = paste(geneNameLong, geneName, sep = "\n"))
  ) %>% 
  dplyr::rename('Geneid' = 'gene')
```

```{r  fig.width = 8, fig.height=9.5}
Chitin_genes <- geneGroups %>% 
  left_join(counts_bygene_long) %>%
  filter(geneGroupSpecific %in% c("chitin_synthase", "chitin_deacetylase", "chitinase"))

chitin_levels <- c("Chitin synthase 1\nCHS1", "Chitin synthase 2\nCHS2", "Chitin synthase 3\nCHS3", "Chitin synthase 4\nCHS4",   
  "Chitin synthase 5\nCHS5", "Chitin synthase 6\nCHS6", "Chitin synthase 7\nCHS7", "Chitin synthase 8\nCHS8",   
  "Chitin deacetylase 1\nCDA1", "Chitin deacetylase 2\nMP98", "Chitin deacetylase 3\nCDA3", "Polysaccharide deacetylase\nFPD1",
  "Chitinase\nCHI2", "Chitinase\nCHI21", "Chitinase\nCHI22", "Endochitinase\nCHI4")

chitin_labels <- c("Chitin synthase 1<br>CHS1", "Chitin synthase 2<br>CHS2", "Chitin synthase 3<br>CHS3", "Chitin synthase 4<br>CHS4<sup>*</sup>",
  "Chitin synthase 5<br>CHS5<sup>*</sup>", "Chitin synthase 6<br>CHS6", "Chitin synthase 7<br>CHS7<sup>*</sup>", "Chitin synthase 8<br>CHS8<sup>†</sup>",   
  "Chitin deacetylase 1<br>CDA1<sup>†</sup>", "Chitin deacetylase 2<br>MP98<sup>†</sup>", "Chitin deacetylase 3<br>CDA3", "Polysaccharide deacetylase<br>FPD1",
  "Chitinase<br>CHI2<sup>*</sup>", "Chitinase<br>CHI21", "Chitinase<br>CHI22", "Endochitinase<br>CHI4")

Chitin_genes$plot_title <- factor(Chitin_genes$plot_title,levels=chitin_levels, labels=chitin_levels)

ggplot(data = filter(Chitin_genes, Condition %in% c("YPD\n30°C", "YPD\n37°C", "YPD+\nEGTA\n37°C")), 
       aes(x = Condition, y = TPM, colour = FLC1)) +
        geom_point(aes(shape = BioRep), size = 1.2, alpha = 1) +
        stat_summary(aes(group=FLC1), fun.min="mean",fun.max="mean",
                    geom="errorbar",linewidth=0.5,width=0.6) +  
        guides(shape = guide_legend(override.aes = list(size = 1, alpha = 1)),
               color = guide_legend(override.aes = list(linetype = 1, shape = NA, size = 2, linewidth = 1.5))) +
  facet_wrap(~plot_title, scales = "free", ncol= 4, drop = F) + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.01)), limits = c(0, NA)) + 
  #scale_y_log10()+
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.box="vertical",
    legend.justification = "left",
    legend.box.just = "left"
  ) + 
  xlab("") + 
  ylab("Transcripts per Million") + 
  theme(strip.text.x = element_text(hjust = 0, size = 8),
        axis.text = element_text(hjust = 0.5, size = 9),
        axis.title.y = element_text(size=12, margin = margin(t = 0, r = 10, b = 0, l = 0)),
        legend.text = element_markdown(size = 12),
        legend.title = element_blank())



#ggsave(here("results/figures/Chitin_genes_TPM_geom_point.png"), width = 8, height = 9.5, dpi = 1000)
```


```{r fig.width = 8, fig.height=11.5}
Beta_glucan_genes <- geneGroups %>% 
  left_join(counts_bygene_long) %>% 
  filter(geneGroupSpecific %in% c("beta_1_3_glucan", "beta_1_6_glucan", "alpha_1_3_glucan"))

beta_glucan_levels <- c("1,3-beta-glucan synthase component\nFKS1","Glucan 1,3-beta-glucosidase\nCNAG_00373","Glucan 1,3-beta-glucosidase\nCNAG_05138", "Glucan 1,3-beta-glucosidase\nEXG104","Glucan 1,3-beta-glucosidase\nCNAG_06336", "Exo-beta-1,3-glucanase\nCNAG_05803","Endo-1,3(4)-beta-glucanase\nEBG1","1,3-beta-glucanosyltransferase\nCNAG_06501", "Alpha-1,3-glucan synthase\nAGS1","Glucan endo-1,3-alpha-glucosidase\nCNAG_02850","Glucan endo-1,3-alpha-glucosidase\nCNAG_07736"," ", "Glucosyltransferase\nKRE5","Glucosidase\nKRE6","Putative glucosidase\nSKN1","Glucosidase\nKRE61","Glucosidase\nKRE62","Beta-glucan synthesis-associated protein\nKRE63","Glucosidase\nKRE64")

beta_glucan_labels <- c("1,3-beta-glucan synthase component\nFKS1","Glucan 1,3-beta-glucosidase\nCNAG_00373","Glucan 1,3-beta-glucosidase\nCNAG_05138", "Glucan 1,3-beta-glucosidase\nEXG104","Glucan 1,3-beta-glucosidase\nBGL2", "Exo-beta-1,3-glucanase\nCNAG_05803","Endo-1,3(4)-beta-glucanase\nEBG1","1,3-beta-glucanosyltransferase\nGAS1", "Alpha-1,3-glucan synthase\nAGS1","Glucan endo-1,3-alpha-glucosidase\nCNAG_02850","Glucan endo-1,3-alpha-glucosidase\nCNAG_07736"," ", "Glucosyltransferase\nKRE5","Glucosidase\nKRE6","Putative glucosidase\nSKN1","Glucosidase\nKRE61","Glucosidase\nKRE62","Beta-glucan synthesis-associated\nKRE63","Glucosidase\nKRE64")

Beta_glucan_genes$plot_title <- factor(Beta_glucan_genes$plot_title,levels=beta_glucan_levels,labels=beta_glucan_labels)



ggplot(data = filter(Beta_glucan_genes, Condition %in% c("YPD\n30°C", "YPD\n37°C", "YPD+\nEGTA\n37°C")), 
       aes(x = Condition, y = TPM, colour = FLC1)) +
        geom_point(aes(shape = BioRep), size = 1.2, alpha = 1) +
        stat_summary(aes(group=FLC1), fun.min="mean",fun.max="mean",
                    geom="errorbar",linewidth=0.5,width=0.6) +  
        guides(shape = guide_legend(override.aes = list(size = 1, alpha = 1)),
               color = guide_legend(override.aes = list(linetype = 1, shape = NA, size = 2, linewidth = 1.5))) +
  facet_wrap(~plot_title, scales = "free", ncol= 4, drop = F) + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.01)), limits = c(0, NA)) + 
  #scale_y_log10()+
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.box="vertical",
    legend.justification = "left",
    legend.box.just = "left"
  ) + 
  xlab("") + 
  ylab("Transcripts per Million") + 
  theme(strip.text.x = element_text(hjust = 0, size = 8),
        axis.text = element_text(hjust = 0.5, size = 9),
        axis.title.y = element_text(size=12, margin = margin(t = 0, r = 10, b = 0, l = 0)),
        legend.text = element_markdown(size = 12),
        legend.title = element_blank())


ggsave(here("results/figures/Glucan_genes_TPM_geom_point.png"), width = 8, height = 11, dpi = 1000)
```


```{r fig.width=11, fig.height=8}
my_genes <- geneGroups %>% 
  filter(Geneid %in% c(
    "CNAG_03326",
    "CNAG_05581",
    "CNAG_00546",
    "CNAG_05818",
    "CNAG_02217",
    "CNAG_07499",
    "CNAG_05799",
    "CNAG_01230",
    "CNAG_03412",
    "CNAG_02351",
    "CNAG_06508",
    "CNAG_02225",
    "CNAG_06336",
    "CNAG_02860",
    "CNAG_06501")
  ) %>% 
  left_join(counts_bygene_long)



gene_levels <- c("Chitin synthase 2\nCHS2", "Chitin synthase 3\nCHS3", "Chitin synthase 4\nCHS4", "Chitin synthase 5\nCHS5", "Chitin synthase 7\nCHS7", "Chitin synthase 8\nCHS8",  "Chitin deacetylase 1\nCDA1", "Chitin deacetylase 2\nMP98","Chitinase\nCHI2", "Endochitinase\nCHI4", "1,3-beta-glucan synthase component\nFKS1", "Glucan 1,3-beta-glucosidase\nEXG104", "Glucan 1,3-beta-glucosidase\nCNAG_06336", "Endo-1,3(4)-beta-glucanase\nEBG1","1,3-beta-glucanosyltransferase\nCNAG_06501")

gene_labels <- c("Chitin synthase 2\nCHS2", "Chitin synthase 3\nCHS3", "Chitin synthase 4\nCHS4", "Chitin synthase 5\nCHS5", "Chitin synthase 7\nCHS7", "Chitin synthase 8\nCHS8",  "Chitin deacetylase 1\nCDA1", "Chitin deacetylase 2\nMP98","Chitinase\nCHI2", "Endochitinase\nCHI4", "1,3-beta-glucan synthase\nFKS1", "Glucan 1,3-beta-glucosidase\nEXG104", "Glucan 1,3-beta-glucosidase\nBGL2", "Endo-1,3(4)-beta-glucanase\nEBG1","1,3-beta-glucanosyltransferase\nGAS1")


my_genes$plot_title <- factor(my_genes$plot_title,levels=gene_levels, labels=gene_labels)

ggplot(data = filter(my_genes, Condition %in% c("YPD\n30°C", "YPD\n37°C", "YPD+\nEGTA\n37°C")), 
       aes(x = Condition, y = TPM, colour = FLC1)) +
        geom_point(aes(shape = BioRep), size = 1.8, alpha = 1) +
        stat_summary(aes(group=FLC1), fun.min="mean",fun.max="mean",
                    geom="errorbar",linewidth=0.6,width=0.7) +  
        guides(shape = guide_legend(override.aes = list(size = 1, alpha = 1)),
               color = guide_legend(override.aes = list(linetype = 1, shape = NA, size = 2, linewidth = 1.5))) +
  facet_wrap(~plot_title, scales = "free", ncol= 5, drop = F) + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)), limits = c(0, NA)) + 
  scale_colour_manual(values = c("#E69F00", "#56B4E9")) + 
  #scale_y_log10()+
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.box="vertical",
    legend.justification = "left",
    legend.box.just = "left"
  ) + 
  xlab("") + 
  ylab("Transcripts per Million") + 
  theme(strip.text.x = element_text(hjust = 0, size = 9),
        axis.text = element_text(hjust = 0.5, size = 9),
        axis.title.y = element_text(size=12, margin = margin(t = 0, r = 10, b = 0, l = 0)),
        legend.text = element_markdown(size = 12),
        legend.title = element_blank())



ggsave(here("results/figures/Cell_wall_genes_TPM_geom_point_poster.png"), width = 11, height = 8, dpi = 1000)
```



```{r fig.width=9, fig.height=6}
my_genes_short <- geneGroups %>% 
  filter(Geneid %in% c(
    #"CNAG_00546",
    "CNAG_05818",
    "CNAG_02217",
    "CNAG_07499",
    "CNAG_06508",
    "CNAG_02225",
    "CNAG_06336"
    #"CNAG_02860"
    )
  ) %>% 
  left_join(counts_bygene_long)



gene_levels_short <- c("Chitin synthase 5\nCHS5", "Chitin synthase 7\nCHS7", "Chitin synthase 8\nCHS8", "1,3-beta-glucan synthase component\nFKS1", "Glucan 1,3-beta-glucosidase\nEXG104", "Glucan 1,3-beta-glucosidase\nCNAG_06336")

gene_labels_short <- c("Chitin synthase 5\nCHS5", "Chitin synthase 7\nCHS7", "Chitin synthase 8\nCHS8", "1,3-beta-glucan synthase\nFKS1", "Glucan 1,3-beta-glucosidase\nEXG104", "Glucan 1,3-beta-glucosidase\nBGL2")


my_genes_short$plot_title <- factor(my_genes_short$plot_title,levels=gene_levels_short, labels=gene_labels_short)

ggplot(data = filter(my_genes_short, Condition %in% c("YPD\n30°C", "YPD\n37°C", "YPD+\nEGTA\n37°C")), 
       aes(x = Condition, y = TPM, colour = FLC1)) +
        geom_point(aes(shape = BioRep), size = 1.4, alpha = 1) +
        stat_summary(aes(group=FLC1), fun.min="mean",fun.max="mean",
                    geom="errorbar",linewidth=0.6,width=0.7) +  
        guides(shape = guide_legend(override.aes = list(size = 1, alpha = 1)),
               color = guide_legend(override.aes = list(linetype = 1, shape = NA, size = 2, linewidth = 1.5))) +
  facet_wrap(~plot_title, scales = "free", ncol= 4, drop = F) + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, NA)) + 
  scale_colour_manual(values = rev(c("#00A9FF", "#E68613"))) + 
  #scale_y_log10()+
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.box="vertical",
    legend.justification = "left",
    legend.box.just = "left"
  ) + 
  xlab("") + 
  ylab("Transcripts per Million") + 
  theme(strip.text.x = element_text(hjust = 0, size = 9),
        axis.text = element_text(hjust = 0.5, size = 9),
        axis.title.y = element_text(size=12, margin = margin(t = 0, r = 10, b = 0, l = 0)),
        legend.text = element_markdown(size = 12),
        legend.title = element_blank())



ggsave(here("results/figures/Cell_wall_genes_TPM_geom_point_poster.png"), width = 9, height = 6, dpi = 1000)
```


```{r fig.width=12, fig.height=5}
my_calcium_genes_short <- geneGroups %>% 
  filter(geneGroup == "calcium_transport") %>% 
  left_join(counts_bygene_long) %>% 
  filter(!Geneid %in% c("CNAG_00025", "CNAG_03751"))

calcium_gene_levels_short <- unique(my_calcium_genes_short$plot_title)

calcium_gene_labels_short <- c("Calcium/proton exchanger\nCNAG_07538", "Calcium/proton exchanger\nCNAG_03805", "Calcium P-type ATPase\nPMC1", "Mechanosensitive channel\nIRK6", "CSC1-like cation channel\nCNAG_06060",    "Calcium-channel protein\nCCH1")

my_calcium_genes_short$plot_title <- factor(my_calcium_genes_short$plot_title,levels=calcium_gene_levels_short, labels=calcium_gene_labels_short)

poster_genes <- bind_rows(my_genes_short, my_calcium_genes_short)

ggplot(data = filter(poster_genes, Condition %in% c("YPD\n30°C", "YPD\n37°C", "YPD+\nEGTA\n37°C")), 
       aes(x = Condition, y = TPM, colour = FLC1)) +
        geom_point(aes(shape = BioRep), size = 1.4, alpha = 1) +
        stat_summary(aes(group=FLC1), fun.min="mean",fun.max="mean",
                    geom="errorbar",linewidth=0.6,width=0.7) +  
        guides(shape = guide_legend(override.aes = list(size = 1, alpha = 1)),
               color = guide_legend(override.aes = list(linetype = 1, shape = NA, size = 2, linewidth = 1.5))) +
  facet_wrap(~plot_title, scales = "free", ncol= 6, drop = F) + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, NA)) + 
  #scale_colour_manual(values = rev(c("#00A9FF", "#E68613"))) + 
  #scale_y_log10()+
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.box="vertical",
    legend.justification = "left",
    legend.box.just = "left"
  ) + 
  xlab("") + 
  ylab("Transcripts per Million") + 
  theme(strip.text.x = element_text(hjust = 0, size = 9),
        axis.text = element_text(hjust = 0.5, size = 9),
        axis.title.y = element_text(size=12, margin = margin(t = 0, r = 10, b = 0, l = 0)),
        legend.text = element_markdown(size = 12),
        legend.title = element_blank())

ggsave(here("results/figures/BMS_poster_TPM_geom_point_col4.png"), width = 12, height = 5, dpi = 1000)
```






## Cell wall synthesis and remodelling genes

I curated a list of genes involved in cell wall chitin and glucan synthesis/remodelling. Here I plot the mean transcripts per million calculated from 3 biological replicated of FLC1 and flc1∆ samples at 30°C and 37°C. Although not all genes were classified as differentially express, almost all cell wall genes are more highly exressed in flc1∆ samples that FLC1 at both growth temperatures. 