---
title: "Overlap in DGE in Flc1∆ vs WT and the Chow et al. 2017 Crz1 and Cna1 DGE dataset"
author: "Rachael Murray"
date: "2024-12-19"
output: html_document
---

# Summary 

Find out how similar the transcriptomes are of crz1∆, cna1∆ and flc1∆ cells relative to WT at 37°C by comparing differential expressed genes in our dataset (flc1∆) with Chow et al. 2017 (crz1∆ and cna1∆). 

## Load packages
```{r load R packages, include = F}
library(here)
library(dplyr)
library(tidyr)
library(janitor)
library(ggplot2)
library(readr)
library(eulerr)
library(ggvenn)
library(stringr)
library(cowplot)
library(gridExtra)
```

## Load differential expressed gene summary from Chow et al. 2017 

```{r load DGE chow2017}
DGE_chow2017 <- read_csv(
  here::here("Chow2017_Crz1_Cna1_targets", "Chow2017_DEG_summary.csv")
) %>% 
  mutate(GeneID = case_when(nchar(GeneID) == 4 ~ paste("CNAG_0", GeneID, sep = ""),
                            nchar(GeneID) == 3 ~ paste("CNAG_00", GeneID, sep = ""),
                            nchar(GeneID) == 2 ~ paste("CNAG_000", GeneID, sep = ""),
                            nchar(GeneID) == 1 ~ paste("CNAG_0000", GeneID, sep = ""))) %>% 
  mutate(Dataset = factor(Dataset, levels = 
           c("cna1_crz1_thermal_stress", 
             "crz1_thermal_stress", 
             "cna1_thermal_stress", 
             "cna1_crz1_no_stress",
             "cna1_no_stress", 
             "crz1_no_stress")))
```
## Load differentially expressed gene summary for Flc1∆ vs WT at 37°C

```{r load DGE list WTvsdelta 37C}

deglist_WTvsdelta_37C_up <- read_delim(
  here::here("results", "DEGs", "deglist_WTvsdelta_37C_YPD_up2x_FDR5.txt"), 
  delim = " ", 
  col_names = "GeneID"
)

deglist_WTvsdelta_37C_down <- read_delim(
  here::here("results", "DEGs", "deglist_WTvsdelta_37C_YPD_down2x_FDR5.txt"), 
  delim = " ", 
  col_names = "GeneID"
)

deglist_WTvsdelta_37C <- bind_rows(
  deglist_WTvsdelta_37C_up, deglist_WTvsdelta_37C_down
)
```
## Plot Venn of DGEs, Flc1∆ cells vs WT at 37°C only
```{r}
# Turn into lists of genes
FLC1_DGE <- deglist_WTvsdelta_37C$GeneID
CRZ1_CNA1_DGE <- unique(DGE_chow2017$GeneID)  

# Add to a list object with names to display on plot
euler_list <- list( 
  'DGE Flc1' = FLC1_DGE, 
  'DGE Crz1 and/or Cna1' = CRZ1_CNA1_DGE)

# Create the euler plot
eulerr_plot <- plot(
  euler(euler_list), 
  quantities = TRUE, 
  legend = TRUE,
  fills = c("#ECCD97", "#9975DB", "grey90"),
  fontsize = 20
  ) 

# Draw the plot with a title 
A <- grid.arrange(grobs = list(eulerr_plot), top = "Differential expressed genes in Flc1∆ cells at 37°C vs WT\n and crz1∆ and cna1∆ cells vs WT with or without thermal stress")


A
```
## Genes in common
```{r}
DGE_chow2017 %>% 
  filter(GeneID %in% deglist_WTvsdelta_37C$GeneID)
```

## Load differentially expressed gene summary for Flc1∆ vs WT at 30°C

```{r load DGE list WTvsdelta 30C}

deglist_WTvsdelta_30C_up <- read_delim(
  here::here("results", "DEGs", "deglist_WTvsdelta_30C_YPD_up2x_FDR5.txt"), 
  delim = " ", 
  col_names = "GeneID"
)

deglist_WTvsdelta_30C_down <- read_delim(
  here::here("results", "DEGs", "deglist_WTvsdelta_30C_YPD_down2x_FDR5.txt"), 
  delim = " ", 
  col_names = "GeneID"
)

deglist_WTvsdelta_30C <- bind_rows(
  deglist_WTvsdelta_30C_up, deglist_WTvsdelta_30C_down
)
```


## Plot Venn of DGEs, Flc1∆ cells vs WT at 37°C only
```{r}
# Turn into lists of genes
FLC1_DGE_37 <- deglist_WTvsdelta_37C$GeneID
FLC1_DGE_30 <- deglist_WTvsdelta_30C$GeneID
CRZ1_CNA1_DGE <- unique(DGE_chow2017$GeneID)  

# Add to a list object with names to display on plot
euler_list <- list( 
  'DGE Flc1 37°C' = FLC1_DGE_37, 
  'DGE Flc1 30°C' = FLC1_DGE_30,
  'DGE Crz1 and/or Cna1' = CRZ1_CNA1_DGE)

# Create the euler plot
eulerr_plot <- plot(
  euler(euler_list), 
  quantities = TRUE, 
  legend = TRUE,
  #fills = c("#ECCD97", "#9975DB", "grey90"),
  fontsize = 20
  ) 

# Draw the plot with a title 
A <- grid.arrange(grobs = list(eulerr_plot), top = "Differential expressed genes in Flc1∆ cells at 37°C or 30°C vs WT\n and crz1∆ and cna1∆ cells vs WT with or without thermal stress")


A
```